### 2019-08-31 paper.tex
added template files

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Paper Title*}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

\begin{abstract}
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
\section{Methods}
\section{Results}
\section{Conclusion}
\section{Acknowledgement}
\bibliographystyle{IEEEtran} 
\bibliography{biblio}

\end{document}

### 2019-08-31 paper.tex
added a very rough intro

@@
 \usepackage{xcolor}
+\usepackage[binary-units]{siunitx}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\newcommand{\bigdata}{Big Data }
+
+\newcommand{\valerie}[1]{\color{blue}\textbf{Note from Valerie}:
+      #1 \color{black}}
+
 \begin{document}
 
-\title{Paper Title*}
+\title{Performance benefits of all-memory processing for large neuroimaging data}
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\author{\IEEEauthorblockN{Val\'erie Hayot-Sasson$^1$, Shawn T Brown$^2$ and 
+    Tristan Glatard$^1$
+  }\\
+  \IEEEauthorblockA{
+    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
+    $^2$Montreal Neurological Institute, McGill University, Montreal, Canada
   }
-
+}
 \maketitle
@@
 \section{Introduction}
-\section{Methods}
+Data transfer rates are a significant source of bottleneck in \bigdata
+applications. As a result, many \bigdata frameworks have opted to reduce data
+transfers using strategies such as data locality, in-memory computing and lazy
+evaluation. Furthermore, kernel optimizations, such as the Page Cache, also
+attempt to minimize the amount of data transfers that may arise during the
+lifetime of an application through cache hits. It also lessens the burden of
+data transfers by asynchronously flushing to desired storage device.
+
+Adaptations made for the efficient processing of \bigdata applications are not 
+limited to software. Infrastructure technology has also improved to increase the
+bandwidth of slower storage devices, to a point at which cost-effective Solid 
+State Drives (SSD) can emulate RAM-like bandwidths and exhibit a significantly 
+larger capacity that standard DRAM. Moreover, such SSDs provide persistent
+storage enabling all data to be stored in a RAM-like non-volatile filesystem.
+\valerie{Perhaps introduce optane here/do related-work search}
+
+Neuroimaging open-data initiatives have led to extensively large repositories of
+publicly available data. Such initiatives include the BigBrain~\cite{BigBrain}, 
+a one-of-a-kind $\SI{603}{\gibi\byte}$\valerie{verify this w/ Claude}
+histological image of a 76 year-old~\valerie{from memory, so better double-check} healthy human brain
+at 20$\mu$m resolution; the UK Biobank~\cite{ukbiobank}, a repository expected to
+contain approximately $\SI{0.2}{\peta\byte}$ of data (including various magnetic
+resonance (MR) imaging modalities) from 500,000 individuals living in the UK;
+the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
+1200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
+and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
+initiative which aggregates MR data from various laboratories around the world, 
+of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
+data in total.
+
+Due to storage limitations, only subsets of the neuroimaging repositories 
+can be processed in a typical research laboratory. Moreover, as these datasets 
+are extremely large and are only increasing in size, they are typically stored 
+in higher-capacity, slower storage devices. In such conditions, large-scale 
+studies in neuroimaging remain limited to labs with access to adequate 
+infrastructure and are limited by data transfer times.
+
+In this paper, we aim to:
+\begin{itemize}
+        \item Quantify the added value of Intel Optane persistent memory on 
+            processing large neuroimaging data using representative pipelines; and
+        \item Determine when a given Intel Optane storage configuration (Memory 
+            and App Direct mode) is preferable.
+\end{itemize}
+
+\section{Materials and Methods}
+\subsection{Infrastructure}
+\subsection{Storage configuration}
+\subsection{Applications}
+\subsection{Performance model}
 \section{Results}
+\subsection{Memory-mode}
+\subsection{App-Direct mode}
+\section{Discussion}
+\subsection{Memory vs App Direct mode}
+\subsection{Effect of Page Cache}
+\subsection{Added value of persistent memory}
 \section{Conclusion}

### 2019-09-01 experiments/bids-app-example/create_invocations.sh
BIDS app example experiment

@@ -1,0 +1,13 @@
+#!/usr/bin/env bash
+echo "#!/usr/bin/env bash" > bosh-commands.sh  # cleans up previous commands
+echo "mkdir -p logs" >> bosh-commands.sh
+for i in  data/corr/RawDataBIDS/*/sub*
+do
+  dataset=$(echo $i | awk -F '/' '{print $4}')
+  subject=$(echo $i | awk -F '/' '{print $5}' | awk -F '-' '{print $2}')
+  exec_name="${dataset}_${subject}"
+  invocation_file="invocations/${exec_name}.json"
+  sed s,DATASET,data/corr/RawDataBIDS/${dataset},g invocations/template.json | sed s,PARTICIPANT,${subject},g | sed s,OUTPUT_DIR,outputs/${dataset}_${subject},g > ${invocation_file}
+  echo "(echo +++++++++++++++++++++++ ; echo -n 'START DATE: date +%s' ; time bosh exec launch bids-example/boutiques/bids-app-example.json ${invocation_file}) &>logs/${exec_name}.log" >> bosh-commands.sh
+  echo $i $dataset $subject 
+done

### 2019-09-01 experiments/bids-app-example/create_invocations.sh
Fine tuning

@@ -1,13 +1,35 @@
 #!/usr/bin/env bash
-echo "#!/usr/bin/env bash" > bosh-commands.sh  # cleans up previous commands
-echo "mkdir -p logs" >> bosh-commands.sh
+
+set -e
+
+if [ $# != 1 ]
+then
+  echo "usage: $0 <experiment_name>"
+  exit 1
+fi
+
+exp_name=$1
+test ${exp_name} || (echo "Directory ${exp_name} already exists, please remove it first." ; exit 1)
+mkdir -p ${exp_name}/{logs,invocations}
+bosh_file=${exp_name}/bosh_commands.sh
+
+echo "#!/usr/bin/env bash" > ${bosh_file}  # cleans up previous commands
+
+echo "mkdir -p ${exp_name}/logs" >> ${bosh_file}
+
 for i in  data/corr/RawDataBIDS/*/sub*
 do
   dataset=$(echo $i | awk -F '/' '{print $4}')
   subject=$(echo $i | awk -F '/' '{print $5}' | awk -F '-' '{print $2}')
   exec_name="${dataset}_${subject}"
-  invocation_file="invocations/${exec_name}.json"
-  sed s,DATASET,data/corr/RawDataBIDS/${dataset},g invocations/template.json | sed s,PARTICIPANT,${subject},g | sed s,OUTPUT_DIR,outputs/${dataset}_${subject},g > ${invocation_file}
-  echo "(echo +++++++++++++++++++++++ ; echo -n 'START DATE: date +%s' ; time bosh exec launch bids-example/boutiques/bids-app-example.json ${invocation_file}) &>logs/${exec_name}.log" >> bosh-commands.sh
-  echo $i $dataset $subject 
+  invocation_file="${exp_name}/invocations/${exec_name}.json"
+  sed s,DATASET,data/corr/RawDataBIDS/${dataset},g invocation_template.json \
+     | sed s,PARTICIPANT,${subject},g | sed s,OUTPUT_DIR,${exp_name}/outputs/${dataset}_${subject},g\
+      > ${invocation_file}
+  echo "(echo -n START DATE: ; date +%s ;\
+         time bosh exec launch bids-example/boutiques/bids-app-example.json\
+         ${invocation_file} ; echo -n END DATE: ; date +%s ) &>${exp_name}/logs/${exec_name}.log" >> ${bosh_file}
+  echo -n .
 done
+echo
+chmod 755 ${bosh_file}

### 2019-09-01 code/spark_inc.py
fixed to incrementation code

@@ -1,0 +1,338 @@
+#!/usr/bin/env python3
+
+from pyspark import SparkContext, SparkConf
+from io import BytesIO
+from time import sleep, time
+from glob import glob
+import os
+import uuid
+import shutil
+import numpy as np
+import nibabel as nib
+import argparse
+import subprocess
+
+
+def write_bench(name, start_time, end_time, node, filename, executor,
+                benchmark_dir=None, benchmark_file=None):
+    """Write benchmarks to file.
+
+    Keyword arguments:
+    name -- name of task benchmarked
+    start_time -- task execution start time (in seconds)
+    end_time -- task execution end time (in seconds)
+    node -- hostname ID of node which executed task
+    filename -- filename of image task processed.
+    executor -- process ID
+    benchmark_dir -- directory to save benchmarks to. Not required if
+                     file provided
+    benchmark_file -- file to append benchmarks to.
+
+    Returns: filename of benchmark file
+    """
+
+    if not benchmark_file:
+        assert benchmark_dir, 'benchmark_dir parameter has not been defined.'
+
+        try:
+            os.makedirs(benchmark_dir)
+        except Exception as e:
+            pass
+
+        benchmark_file = os.path.join(
+                benchmark_dir,
+                "bench-{}.txt".format(str(uuid.uuid1()))
+                )
+
+    with open(benchmark_file, 'a+') as f:
+        f.write('{0},{1},{2},{3},{4},{5}\n'.format(name, start_time, end_time,
+                                                   node, filename, executor))
+
+    return benchmark_file
+
+
+def read_img(filename, data, benchmark, start, bench_dir=None):
+    """Convert image in-memory binary data into Numpy array.
+    For in-memory Spark only.
+
+    Keyword arguments:
+    filename -- filename for which data belongs to
+    data -- binary image data
+    benchmark -- boolean value that specifies whether to benchmark the task
+    start -- start time of the driver (in seconds)
+    benchmark_dir -- benchmark directory required if benchmarking is requested.
+                     (default None)
+
+    Returns: Tuple containing image data and metadata
+    """
+    start_time = time() - start
+
+    print('Reading image')
+    # load binary data into Nibabel
+    fh = nib.FileHolder(fileobj=BytesIO(data))
+    im = nib.Nifti1Image.from_file_map({'header': fh, 'image': fh})
+
+    data = np.asanyarray(im.dataobj)
+
+    end_time = time() - start
+
+    bn = os.path.basename(filename)
[…]

### 2019-09-01 code/spark_inc.py
fixed in-memory increment

@@ -13,7 +13,7 @@ import argparse
 import subprocess
 
 
-def write_bench(name, start_time, end_time, node, filename, executor,
+def write_bench(name, start_time, end_time, filename, executor,
                 benchmark_dir=None, benchmark_file=None):
     """Write benchmarks to file.
 
@@ -45,8 +45,8 @@ def write_bench(name, start_time, end_time, node, filename, executor,
                 )
 
     with open(benchmark_file, 'a+') as f:
-        f.write('{0},{1},{2},{3},{4},{5}\n'.format(name, start_time, end_time,
-                                                   node, filename, executor))
+        f.write('{0},{1},{2},{3},{4}\n'.format(name, start_time, end_time,
+                                                   filename, executor))
 
     return benchmark_file
 
@@ -80,12 +80,10 @@ def read_img(filename, data, benchmark, start, bench_dir=None):
 
     bench_file = None
     if benchmark:
-        bench_file = write_bench('load_img', start_time, end_time,
-                                 socket.gethostname(), bn,
+        bench_file = write_bench('load_img', start_time, end_time, bn,
                                  os.getpid(), benchmark_dir=bench_dir)
 
-    return (socket.gethostname(), filename, im,
-            bench_file)
+    return (os.getpid(), filename, im, bench_file)
 
 
 def increment_data(idx, filename, im, delay, benchmark, start,
@@ -94,10 +92,9 @@ def increment_data(idx, filename, im, delay, benchmark, start,
     """Increment image data by 1
 
     Keyword arguments:
-    idx -- image index within the RDD (Or node hostname if in-memory)
+    idx -- image index within the RDD (Or pid if in-memory)
     filename -- image filename
-    data -- image data (can be None w/ CLI-based processing)
-    metadata -- image metadata (can be None w/ CLI-based processing)
+    im -- Nibabel image (can be None w/ CLI-based processing)
     delay -- task duration (or sleep time) in seconds
     benchmark -- boolean representing whether task should be benchmarked
     start -- start time of driver
@@ -126,7 +123,7 @@ def increment_data(idx, filename, im, delay, benchmark, start,
     if not cli:
         print("Incrementing data in memory")
         data = np.asanyarray(im.dataobj) + 1
-        im = nib.Nifti1Image(data, im.header, im.affine)
+        im = nib.Nifti1Image(data, im.affine, im.header)
         sleep(delay)
     else:
[…]

### 2019-09-01 code/spark_inc.py
fixed write to disk

@@ -87,8 +87,7 @@ def read_img(filename, data, benchmark, start, bench_dir=None):
 
 
 def increment_data(idx, filename, im, delay, benchmark, start,
-                   iteration=0, work_dir=None, bench_file=None,
-                   cli=False):
+                   iteration=0, work_dir=None, bench_file=None):
     """Increment image data by 1
 
     Keyword arguments:
@@ -99,15 +98,15 @@ def increment_data(idx, filename, im, delay, benchmark, start,
     benchmark -- boolean representing whether task should be benchmarked
     start -- start time of driver
     iteration -- the current iteration value (default 0)
-    work_dir -- work directory to save intermediate results to with CLI
+    work_dir -- work directory to save intermediate results to
                (default None)
     bench_file -- the filepath to save benchmarks to
-    cli -- boolean denoting whether data should be incremented with CLI
 
     Returns: Tuple containing incremented data and associated metadata
     """
     start_time = time() - start
 
+    print('****', work_dir)
     if benchmark and os.path.isdir(bench_file):
         bench_dir = bench_file
         try:
@@ -120,13 +119,12 @@ def increment_data(idx, filename, im, delay, benchmark, start,
                         "bench-{}.txt".format(str(uuid.uuid1()))
                      )
 
-    if not cli:
+    if not work_dir:
         print("Incrementing data in memory")
         data = np.asanyarray(im.dataobj) + 1
         im = nib.Nifti1Image(data, im.affine, im.header)
         sleep(delay)
     else:
-        work_dir = output_dir if work_dir is None else work_dir
         it_dir = "iteration-{}".format(iteration)
 
         work_dir = os.path.join(work_dir, it_dir)
@@ -138,20 +136,20 @@ def increment_data(idx, filename, im, delay, benchmark, start,
 
         fn = filename[5:] if 'file:' in filename else os.path.abspath(filename)
 
-        if benchmark:
-            p = subprocess.Popen(['increment.py', fn,
-                                  work_dir, '--benchmark_file', bench_file,
-                                  '--start', str(start),
-                                  '--delay', str(delay)])
-        else:
-            p = subprocess.Popen(['increment.py', fn,
-                                  work_dir, '--delay', str(delay)])
+        im = nib.load(fn)
+        data = np.asanyarray(im.dataobj) + 1
+        im = nib.Nifti1Image(data, im.affine, im.header)
 
[…]

### 2019-09-01 code/experiment_launch.py
added experiment launch script

@@ -1,0 +1,102 @@
+#!/usr/bin/env python3
+
+import json
+from random import shuffle
+from os import path as op
+import subprocess
+import glob
+
+base_dir = '/home/user/vhayots' 
+block_dir = '1000_blocks'
+input_dataset = op.join(base_dir, block_dir)
+isilon = base_dir
+optane = '/run/user/61218'
+cmd_template = 'spark-submit --master local[{0}] --driver-memory {1} spark_inc.py {2} {3} {4} --benchmark'
+im_size_b = 646461552 
+
+with open('conditions.json', 'r') as f:
+    conditions = json.load(f)
+
+shuffle(conditions)
+
+def drop_caches():
+    p = subprocess.Popen('echo 3 | sudo tee /proc/sys/vm/drop_caches',
+                         shell=True)
+    print(p.communicate())
+
+
+def empty_dir(regex):
+    files = glob.glob(regex)
+    for f in files:
+            os.remove(f)
+
+def delete_output_files(work_dir_flag, mem_dir, mem_dir_out, isilon_dir_out):
+    if isilon_dir_out is not None:
+        empty_dir(op.join(isilon_dir_out, '*.nii'))
+        empty_dir(op.join(work_dir_flag.split(' ')[1], '*.nii'))
+
+    else:
+        empty_dir(op.join(mem_dir, '*.nii'))
+        empty_dir(op.join(mem_dir_out, '*.nii'))
+
+
+for i in range(5):
+    for c in conditions:
+
+        parallel_writes = c['ncpus'] * im_size_b
+        work_dir_flag = None
+        mem_dir_out = None 
+        mem_dir = None
+        isilon_dir_out = None
+
+        if c['storage'] == 'optane':
+            mem_dir = op.join(optane, block_dir) 
+            mem_dir_out = op.join(optane, '{0}-{1}'.format(c['id'], i))
+            p = subprocess.Popen(['cp', input_dataset, mem_dir],
+                                 stdout=subprocess.PIPE, stderr=subprocess.PIPE)
+            print(p.communicate())
+
+            drop_caches()
+
+            cmd = cmd_template.format(c['ncpus'],
+                                      str((parallel_writes * 0.05 + parallel_writes) / 1024**2) + 'G',
+                                      mem_dir,
+                                      mem_dir_out,
+                                      c['iterations'])
+
+        else:
+            isilon_dir_out = op.join(isilon, '{0}-{1}'.format(c['id'], i))
+            drop_caches()
[…]

### 2019-09-01 code/experiment_launch.py
minor code fixes

@@ -2,11 +2,11 @@
 
 import json
 from random import shuffle
-from os import path as op
+from os import path as op, remove
 import subprocess
 import glob
 
-base_dir = '/home/user/vhayots' 
+base_dir = '/home/users/vhayots' 
 block_dir = '1000_blocks'
 input_dataset = op.join(base_dir, block_dir)
 isilon = base_dir
@@ -14,7 +14,7 @@ optane = '/run/user/61218'
 cmd_template = 'spark-submit --master local[{0}] --driver-memory {1} spark_inc.py {2} {3} {4} --benchmark'
 im_size_b = 646461552 
 
-with open('conditions.json', 'r') as f:
+with open('test_condition.json', 'r') as f:
     conditions = json.load(f)
 
 shuffle(conditions)
@@ -28,7 +28,7 @@ def drop_caches():
 def empty_dir(regex):
     files = glob.glob(regex)
     for f in files:
-            os.remove(f)
+        remove(f)
 
 def delete_output_files(work_dir_flag, mem_dir, mem_dir_out, isilon_dir_out):
     if isilon_dir_out is not None:
@@ -48,18 +48,21 @@ for i in range(5):
         mem_dir_out = None 
         mem_dir = None
         isilon_dir_out = None
+        mem_size = str(int((parallel_writes * 0.05 + parallel_writes) / 1024**3)) + 'G'
+        print(mem_size)
 
         if c['storage'] == 'optane':
             mem_dir = op.join(optane, block_dir) 
             mem_dir_out = op.join(optane, '{0}-{1}'.format(c['id'], i))
-            p = subprocess.Popen(['cp', input_dataset, mem_dir],
+            p = subprocess.Popen(['cp', '-r', input_dataset, mem_dir],
                                  stdout=subprocess.PIPE, stderr=subprocess.PIPE)
             print(p.communicate())
 
             drop_caches()
 
+
             cmd = cmd_template.format(c['ncpus'],
-                                      str((parallel_writes * 0.05 + parallel_writes) / 1024**2) + 'G',
+                                      mem_size,
                                       mem_dir,
                                       mem_dir_out,
                                       c['iterations'])
@@ -68,8 +71,8 @@ for i in range(5):
             isilon_dir_out = op.join(isilon, '{0}-{1}'.format(c['id'], i))
             drop_caches()
             cmd = cmd_template.format(c['ncpus'],
-                                      parallel_writes * 0.05 + parallel_writes,
-                                      block_dir,
[…]

### 2019-09-01 code/spark_inc.py

@@ -144,7 +144,7 @@ def increment_data(idx, filename, im, delay, benchmark, start,
         fn = os.path.basename(fn)
 
         if '-' in fn:
-            fn = fn.split('-')[1]
+            fn = fn.split('-')[-1]
 
         out_fn = 'inc{0}-{1}'.format(iteration, fn)
         out_fp = os.path.join(work_dir, out_fn)
@@ -185,8 +185,8 @@ def save_incremented(idx, filename, im, benchmark, start,
 
     start_time = time() - start
 
-    bn = os.path.basename(filename).replace('inc-', '')
-    out_fn = os.path.join(output_dir, 'inc{0}-{1}'.format(iterations, bn))
+    bn = os.path.basename(filename) #.replace('inc-', '')
+    out_fn = os.path.join(output_dir, bn) #'inc{0}-{1}'.format(iterations, bn))
 
     if not wd:
         nib.save(im, out_fn)
@@ -271,7 +271,7 @@ def main():
                                                        bench_file=x[3]))
     else:
         # get all filenames
-        files = glob(os.path.join(args.bb_dir, '*'))
+        files = glob(os.path.join(args.bb_dir, 'bigbrain*'))
         fidx = [i for i in range(0, len(files))]
 
         # Create an RDD of filenames

### 2019-09-01 code/experiment_launch.py
updated condition file reference

@@ -14,7 +14,7 @@ optane = '/run/user/61218'
 cmd_template = 'spark-submit --master local[{0}] --driver-memory {1} spark_inc.py {2} {3} {4} --benchmark'
 im_size_b = 646461552 
 
-with open('test_condition.json', 'r') as f:
+with open('conditions.json', 'r') as f:
     conditions = json.load(f)
 
 shuffle(conditions)

### 2019-09-02 experiments/bids-app-example/analyze_results.sh
Added exit code

@@ -21,6 +21,7 @@ first=true
 i=0
 for log in $(ls -tra ${exp_name}/logs/*.log)
 do
+  echo -n .
   if [ "${first}" = false ]
   then
     echo "," >> ${task_file}
@@ -33,6 +34,9 @@ do
   start=$(date -d @${start})  # convert from unix timestamp to date
   end=$(grep 'END DATE' ${log} | awk -F ':' '{print $2}')
   end=$(date -d @${end})  # convert from unix timestamp to date
-  echo "  { \"Task\": \"${task_number}\", \"Start\": \"${start}\", \"Finish\": \"${end}\" } " >> ${task_file}
+  code=$(grep -A 1 -i "exit code" ${log} | tail -1 | sed -r "s/\x1B\[([0-9]{1,2}(;[0-9]{1,2})?)?[mGK]//g")
+  echo "  { \"Task\": \"${task_number}\", \"Start\": \"${start}\", \
+            \"Finish\": \"${end}\", \"Name\": \"${job_name}\", \
+            \"Exit code\": \"${code}\" } " >> ${task_file}
 done
 echo '] }' >> ${task_file}

### 2019-09-02 code/experiment_launch.py
incrementation experiment bug fixed

@@ -5,9 +5,10 @@ from random import shuffle
 from os import path as op, remove
 import subprocess
 import glob
+import sys
 
 base_dir = '/home/users/vhayots' 
-block_dir = '1000_blocks'
+block_dir = sys.argv[1] #'1000_blocks'
 input_dataset = op.join(base_dir, block_dir)
 isilon = base_dir
 optane = '/run/user/61218'
@@ -40,7 +41,7 @@ def delete_output_files(work_dir_flag, mem_dir, mem_dir_out, isilon_dir_out):
         empty_dir(op.join(mem_dir_out, '*.nii'))
 
 
-for i in range(5):
+for i in range(int(sys.argv[2])):
     for c in conditions:
 
         parallel_writes = c['ncpus'] * im_size_b
@@ -54,7 +55,7 @@ for i in range(5):
         if c['storage'] == 'optane':
             mem_dir = op.join(optane, block_dir) 
             mem_dir_out = op.join(optane, '{0}-{1}'.format(c['id'], i))
-            p = subprocess.Popen(['cp', '-r', input_dataset, mem_dir],
+            p = subprocess.Popen(['cp', '-r', input_dataset, optane],
                                  stdout=subprocess.PIPE, stderr=subprocess.PIPE)
             print(p.communicate())
 

### 2019-09-02 code/spark_inc.py

@@ -288,10 +288,10 @@ def main():
                           preservesPartitioning=True)
 
         for i in range(1, args.iterations):
-            imRDD = imRDD.map(lambda x: increment_data(x[0], x[1], None, None,
+            imRDD = imRDD.map(lambda x: increment_data(x[0], x[1], None,
                                                        delay, args.benchmark,
                                                        start,
-                                                       x[5],
+                                                       x[4],
                                                        work_dir,
                                                        benchmark_dir),
                               preservesPartitioning=True)

### 2019-09-02 experiments/bigbrain-incrementation/spark_inc.py
removed working dir option for spark increment

@@ -87,7 +87,7 @@ def read_img(filename, data, benchmark, start, bench_dir=None):
 
 
 def increment_data(idx, filename, im, delay, benchmark, start,
-                   iteration=0, work_dir=None, bench_file=None):
+                   iteration=0, bench_file=None):
     """Increment image data by 1
 
     Keyword arguments:
@@ -98,15 +98,12 @@ def increment_data(idx, filename, im, delay, benchmark, start,
     benchmark -- boolean representing whether task should be benchmarked
     start -- start time of driver
     iteration -- the current iteration value (default 0)
-    work_dir -- work directory to save intermediate results to
-               (default None)
     bench_file -- the filepath to save benchmarks to
 
     Returns: Tuple containing incremented data and associated metadata
     """
     start_time = time() - start
 
-    print('****', work_dir)
     if benchmark and os.path.isdir(bench_file):
         bench_dir = bench_file
         try:
@@ -119,39 +116,10 @@ def increment_data(idx, filename, im, delay, benchmark, start,
                         "bench-{}.txt".format(str(uuid.uuid1()))
                      )
 
-    if not work_dir:
-        print("Incrementing data in memory")
-        data = np.asanyarray(im.dataobj) + 1
-        im = nib.Nifti1Image(data, im.affine, im.header)
-        sleep(delay)
-    else:
-        it_dir = "iteration-{}".format(iteration)
-
-        work_dir = os.path.join(work_dir, it_dir)
-
-        try:
-            os.makedirs(work_dir)
-        except Exception as e:
-            pass
-
-        fn = filename[5:] if 'file:' in filename else os.path.abspath(filename)
-
-        im = nib.load(fn)
-        data = np.asanyarray(im.dataobj) + 1
-        im = nib.Nifti1Image(data, im.affine, im.header)
-
-
-        fn = os.path.basename(fn)
-
-        if '-' in fn:
-            fn = fn.split('-')[-1]
-
-        out_fn = 'inc{0}-{1}'.format(iteration, fn)
-        out_fp = os.path.join(work_dir, out_fn)
-
-        nib.save(im, out_fp)
-
-        filename = out_fp
+    print("Incrementing data in memory")
+    data = np.asanyarray(im.dataobj) + 1
+    im = nib.Nifti1Image(data, im.affine, im.header)
+    sleep(delay)
 
     end_time = time() - start
 
@@ -166,7 +134,7 @@ def increment_data(idx, filename, im, delay, benchmark, start,
 
 
[…]

### 2019-09-02 experiments/bigbrain-incrementation/experiment_launch.py
modified experiment launch script

@@ -45,40 +45,28 @@ for i in range(int(sys.argv[2])):
     for c in conditions:
 
         parallel_writes = c['ncpus'] * im_size_b
-        work_dir_flag = None
-        mem_dir_out = None 
-        mem_dir = None
-        isilon_dir_out = None
+        in_dir = input_dataset
+        out_dir = None
         mem_size = str(int((parallel_writes * 0.05 + parallel_writes) / 1024**3)) + 'G'
         print(mem_size)
 
         if c['storage'] == 'optane':
-            mem_dir = op.join(optane, block_dir) 
-            mem_dir_out = op.join(optane, '{0}-{1}'.format(c['id'], i))
+            in_dir = op.join(optane, block_dir) 
+            out_dir = op.join(optane, '{0}-{1}'.format(c['id'], i))
+
+            # Copy dataset to memory
             p = subprocess.Popen(['cp', '-r', input_dataset, optane],
                                  stdout=subprocess.PIPE, stderr=subprocess.PIPE)
             print(p.communicate())
-
-            drop_caches()
-
-
-            cmd = cmd_template.format(c['ncpus'],
-                                      mem_size,
-                                      mem_dir,
-                                      mem_dir_out,
-                                      c['iterations'])
-
         else:
-            isilon_dir_out = op.join(isilon, '{0}-{1}'.format(c['id'], i))
-            drop_caches()
-            cmd = cmd_template.format(c['ncpus'],
-                                      mem_size,
-                                      input_dataset,
-                                      isilon_dir_out,
-                                      c['iterations'])
-            work_dir_flag = '--work_dir {}'.format(op.join(isilon, 'work_dir'))
-            cmd = '{0} {1}'.format(cmd, work_dir_flag)
+            out_dir = op.join(isilon, '{0}-{1}'.format(c['id'], i))
 
+        drop_caches()
+        cmd = cmd_template.format(c['ncpus'],
+                                  mem_size,
+                                  in_dir,
+                                  out_dir,
+                                  c['iterations'])
 
         if c['delay'] != 0 :
             delay_flag = '--delay {}'.format(c['delay'])

### 2019-09-02 experiments/bigbrain-incrementation/experiment_launch.py
bug fixes

@@ -22,7 +22,7 @@ shuffle(conditions)
 
 def drop_caches():
     p = subprocess.Popen('echo 3 | sudo tee /proc/sys/vm/drop_caches',
-                         shell=True)
+                         shell=True, stdin=subprocess.PIPE, stdout=subprocess.PIPE)
     print(p.communicate())
 
 
@@ -31,20 +31,18 @@ def empty_dir(regex):
     for f in files:
         remove(f)
 
-def delete_output_files(work_dir_flag, mem_dir, mem_dir_out, isilon_dir_out):
-    if isilon_dir_out is not None:
-        empty_dir(op.join(isilon_dir_out, '*.nii'))
-        empty_dir(op.join(work_dir_flag.split(' ')[1], '*.nii'))
-
-    else:
+def delete_output_files(mem_dir, out_dir):
+    if mem_dir is not None:
         empty_dir(op.join(mem_dir, '*.nii'))
-        empty_dir(op.join(mem_dir_out, '*.nii'))
+
+    empty_dir(op.join(out_dir, '*.nii'))
 
 
 for i in range(int(sys.argv[2])):
     for c in conditions:
 
         parallel_writes = c['ncpus'] * im_size_b
+        mem_in_dir = None
         in_dir = input_dataset
         out_dir = None
         mem_size = str(int((parallel_writes * 0.05 + parallel_writes) / 1024**3)) + 'G'
@@ -52,6 +50,7 @@ for i in range(int(sys.argv[2])):
 
         if c['storage'] == 'optane':
             in_dir = op.join(optane, block_dir) 
+            mem_in_dir = in_dir
             out_dir = op.join(optane, '{0}-{1}'.format(c['id'], i))
 
             # Copy dataset to memory
@@ -78,7 +77,7 @@ for i in range(int(sys.argv[2])):
         p = subprocess.Popen(cmd, stdin=subprocess.PIPE, stderr=subprocess.PIPE)
         print(p.communicate())
 
-        delete_output_files(work_dir_flag, mem_dir, mem_dir_out, isilon_dir_out)
+        delete_output_files(mem_in_dir, out_dir)
 
 
 

### 2019-09-02 experiments/bigbrain-incrementation/experiment_launch.py
fixes to experiment launch script

@@ -6,13 +6,14 @@ from os import path as op, remove
 import subprocess
 import glob
 import sys
+from shutil import copytree
 
 base_dir = '/home/users/vhayots' 
 block_dir = sys.argv[1] #'1000_blocks'
 input_dataset = op.join(base_dir, block_dir)
 isilon = base_dir
 optane = '/run/user/61218'
-cmd_template = 'spark-submit --master local[{0}] --driver-memory {1} spark_inc.py {2} {3} {4} --benchmark'
+cmd_template = 'spark-submit --master local[{0}] --driver-memory {1} --conf spark.network.timeout=10000001 --conf spark.executor.heartbeatInterval=10000000 spark_inc.py {2} {3} {4} --benchmark'
 im_size_b = 646461552 
 
 with open('conditions.json', 'r') as f:
@@ -54,9 +55,8 @@ for i in range(int(sys.argv[2])):
             out_dir = op.join(optane, '{0}-{1}'.format(c['id'], i))
 
             # Copy dataset to memory
-            p = subprocess.Popen(['cp', '-r', input_dataset, optane],
-                                 stdout=subprocess.PIPE, stderr=subprocess.PIPE)
-            print(p.communicate())
+            in_dir = copytree(input_dataset, in_dir)
+            print(in_dir, 'created')
         else:
             out_dir = op.join(isilon, '{0}-{1}'.format(c['id'], i))
 

### 2019-09-02 experiments/bigbrain-incrementation/experiment_launch.py
changed in-mem experiment mountpoint

@@ -12,7 +12,7 @@ base_dir = '/home/users/vhayots'
 block_dir = sys.argv[1] #'1000_blocks'
 input_dataset = op.join(base_dir, block_dir)
 isilon = base_dir
-optane = '/run/user/61218'
+optane = '/nvme-disk1'
 cmd_template = 'spark-submit --master local[{0}] --driver-memory {1} --conf spark.network.timeout=10000001 --conf spark.executor.heartbeatInterval=10000000 spark_inc.py {2} {3} {4} --benchmark'
 im_size_b = 646461552 
 

### 2019-09-03 experiments/bigbrain-incrementation/increment.py
added incrementation cli

@@ -1,0 +1,94 @@
+#!/usr/bin/env python
+
+import argparse
+import nibabel as nib
+from os import path as op, makedirs as md
+import time
+import subprocess
+import socket
+try:
+    from threading import get_ident
+except Exception as e:
+    from thread import get_ident
+
+
+def increment(fn, outdir, delay, benchmark_file, start_time):
+    """Increment image data by 1
+
+    Keyword arguments:
+    fn -- filename containing image to increment
+    outdir -- output directory to save data to
+    delay -- task duration (sleep time) in seconds
+    benchmark_file -- file to save benchmarks to
+    start_time -- application driver start time
+    """
+    print('Incrementing image: ', fn)
+
+    start = time.time() - start_time
+    im = nib.load(fn)
+    end = time.time() - start_time
+    print("read time", end)
+
+    if benchmark_file is not None:
+        write_bench(benchmark_file, "read_file", start, end,
+                    socket.gethostname(), op.basename(fn), get_ident())
+
+    inc_data = im.get_data() + 1
+
+    im = nib.Nifti1Image(inc_data, affine=im.affine, header=im.header)
+
+    out_fn = ('inc-{}'.format(op.basename(fn))
+              if 'inc' not in op.basename(fn)
+              else op.basename(fn))
+
+    out_fn = op.join(outdir, out_fn)
+
+    start = time.time() - start_time
+    nib.save(im, out_fn)
+    end = time.time() - start_time
+    print("write time", time.time() - start)
+
+    if benchmark_file is not None:
+        write_bench(benchmark_file, "write_file", start, end,
+                    socket.gethostname(), op.basename(out_fn), get_ident())
+
+    time.sleep(delay)
+    print('Saved image to: ', out_fn)
+
+
+def write_bench(benchmark_file, name, start_time, end_time, node, filename,
+                executor):
+
+    with open(benchmark_file, 'a+') as f:
+        f.write('{0} {1} {2} {3} {4} {5}\n'.format(name, start_time, end_time,
+                                                   node, filename, executor))
+
+
+def main():
+
+    print('Incrementation CLI started')
+    parser = argparse.ArgumentParser(description="BigBrain incrementation")
+    parser.add_argument('filename', type=str,
+                        help=('the file to be incremented'))
+    parser.add_argument('output_dir', type=str,
+                        help='the output directory')
+    parser.add_argument('--benchmark_file', type=str, default=None,
+                        help='the file to write benchmarks to')
[…]

### 2019-09-14 experiments/bids-app-example/launch-disk.sh
Added disk reps

@@ -1,19 +1,37 @@
 #!/usr/bin/env bash
 
 date
-echo DISK-REP1
+echo DISK-REP2
 echo 3 | sudo tee /proc/sys/vm/drop_caches                                       
-parallel --jobs 96 < disk-1/bosh_commands.sh                                
+parallel --jobs 96 < disk-2/bosh_commands.sh                                
 
 
 sleep 100  # cool down a little
 
-mv disk-1 ~/tristan/paper-memory-storage/experiments/bids-app-example/
+mv disk-2 ~/tristan/paper-memory-storage/experiments/bids-app-example/
     
 date
-echo DISK-REP4
+echo DISK-REP5
 echo 3 | sudo tee /proc/sys/vm/drop_caches                                       
-parallel --jobs 25 < disk-4/bosh_commands.sh  
+parallel --jobs 25 < disk-5/bosh_commands.sh  
 
-mv disk-4 ~/tristan/paper-memory-storage/experiments/bids-app-example/
+mv disk-5 ~/tristan/paper-memory-storage/experiments/bids-app-example/
     
+sleep 100  # cool down a little
+
+date
+echo DISK-REP3
+echo 3 | sudo tee /proc/sys/vm/drop_caches                                       
+parallel --jobs 96 < disk-3/bosh_commands.sh                                
+
+
+sleep 100  # cool down a little
+
+mv disk-3 ~/tristan/paper-memory-storage/experiments/bids-app-example/
+    
+date
+echo DISK-REP6
+echo 3 | sudo tee /proc/sys/vm/drop_caches                                       
+parallel --jobs 25 < disk-6/bosh_commands.sh  
+
+mv disk-6 ~/tristan/paper-memory-storage/experiments/bids-app-example/

### 2019-09-14 experiments/bids-app-example/pmem-3/bosh_commands.sh
added pmem analysis

@@ -1,0 +1,687 @@
+#!/usr/bin/env bash
+mkdir -p pmem-3/logs
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025457.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025457.log
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025458.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025458.log
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025459.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025459.log
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025460.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025460.log
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025461.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025461.log
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025462.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025462.log
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025463.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025463.log
+(echo -n START DATE: ; date +%s ;         time bosh exec launch bids-example/boutiques/bids-app-example.json         pmem-3/invocations/IACAS_0025464.json ; echo -n END DATE: ; date +%s ) &>pmem-3/logs/IACAS_0025464.log
[…]

### 2019-09-14 experiments/bigbrain-incrementation/figures/makespan.py
some cleanup and fixes to barchart script

@@ -1,0 +1,67 @@
+#!/usr/bin/env python3
+
+import sys
+import pandas as pd
+import matplotlib.pyplot as plt
+from glob import glob
+from os import path as op
+import numpy as np
+
+
+def makespan(df, task="task_duration"):
+
+    labels = ['tmpfs', 'Optane', 'local disk', 'Isilon']
+    ind = np.arange(len(labels))
+    ind_mem = np.delete(ind, 1)
+    dft = df[df["Task"] == "task_duration"].groupby(["filename"])
+
+    df_end = dft.End.max()
+    df_start = dft.Start.min()
+    df_td = pd.concat([df_start, df_end], axis=1)
+    df_td['makespan'] = df_td.End - df_td.Start
+    df_td['disk'] = df_td.index.map(lambda x: x.split('_')[0].split('-')[-1])
+    
+    s_mean = df_td.rename(columns={'makespan': 'mean'}).set_index('disk').groupby(level=0).mean()
+    s_std = df_td.rename(columns={'makespan': 'std'}).set_index('disk').groupby(level=0).std()
+
+    df_res = pd.concat([s_mean['mean'], s_std['std']], axis=1)
+
+
+    df_res_mem = df_res[df_res.index.map(lambda x: x in ['tmpfs', 'local', 'isilon'])]
+    df_res_mem = df_res_mem.sort_index(ascending=False)
+
+    df_res_ad = df_res[df_res.index.map(lambda x: 'AD' in x)]
+    df_res_ad = df_res_ad.sort_index(ascending=False)
+    df_res_ad.rename(index={'tmpfsAD':'tmpfs', 'optaneAD': 'optane', 'localAD': 'local', 'isilonAD':'isilon'}, inplace=True)
+
+    print(df_res_mem)
+    print(df_res_ad)
+
+    width = 0.35
+
+    fig, ax = plt.subplots()
+    pmem_read = ax.bar(ind_mem - width/2, df_res_mem['mean'], width, alpha=0.4, color='b', yerr=df_res_mem['std'], label="Memory mode")
+    pad_read = ax.bar(ind + width/2, df_res_ad['mean'], width, alpha=0.4, color='orange', yerr=df_res_ad['std'], label="App Direct mode")
+
+    #plt.ylim(0, 13000)
+    ax.set_ylabel('Mean makespan (s)')
+    ax.set_xlabel('Storage type')
+    ax.set_xticks(ind)
+    ax.set_xticklabels(labels)
[…]

### 2019-09-14 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -3,45 +3,89 @@
 import sys
 import pandas as pd
 import matplotlib.pyplot as plt
+import matplotlib.patches as mpatch
 from glob import glob
 from os import path as op
 import numpy as np
 
 
+def add_disk_col(df):
+    df['disk'] = df.index.map(lambda x: x.split('_')[0].split('-')[-1])
+
+
 def stacked_bar(df):
 
-    gk = df.groupby(["filename"]).groups.keys()
+    labels = ['tmpfs', 'Optane', 'local disk', 'Isilon']
+    ind = np.arange(len(labels))
+    ind_mem = np.delete(ind, 1)
+
+    dfr = df[df["Task"] == "read_file"].rename(columns={'Duration': 'read_time'}).groupby(["filename"])["read_time"].sum()
+    dfi = df[df["Task"] == "increment_file"].rename(columns={'Duration': 'increment_time'}).groupby(["filename"])["increment_time"].sum()
+    dfw = df[df["Task"] == "write_file"].rename(columns={'Duration': 'write_time'}).groupby(["filename"])["write_time"].sum()
+
+    df_sum = pd.concat([dfr, dfi, dfw], axis=1)
     
-    get_disk = lambda x: x.split('_')[0].split('-')[-1]
+    add_disk_col(df_sum)
 
-    ind = [get_disk(g) for g in gk]
+    df_mean = df_sum.set_index('disk').groupby('disk').mean()
+    df_std = df_sum.set_index('disk').groupby('disk').std()
+    print('\n\nMean:')
+    print(df_mean)
+    print('\n\nSTD:')
+    print(df_std)
 
-    dfr = df[df["Task"] == "read_file"].groupby(["filename"])["Duration"].sum()
-    dfi = df[df["Task"] == "increment_file"].groupby(["filename"])["Duration"].sum()
-    dfw = df[df["Task"] == "write_file"].groupby(["filename"])["Duration"].sum()
+    def prep_mem_df(df):
+        mem_labels = ['tmpfs', 'local', 'isilon']
+        df = df[df.index.map(lambda x: x in mem_labels)]
+        return df.sort_index(ascending=False)
 
-    width = 0.4
-    print(dfr)
-    print(dfi)
-    print(dfw)
+    df_mean_mem = prep_mem_df(df_mean)
+    df_std_mem = prep_mem_df(df_std)
 
[…]

### 2019-09-16 report/report.tex
added report and fixed figure generation scripts

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\algblock{Input}{EndInput}
\algnotext{EndInput}
\algblock{Output}{EndOutput}
\algnotext{EndOutput}
\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[binary-units]{siunitx}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand{\bigdata}{Big Data }
\newcommand{\bigbrain}{BigBrain }

\newcommand{\valerie}[1]{\color{blue}\textbf{Note from Valerie}:
      #1 \color{black}}

\begin{document}

\title{Performance benefits of Intel Optane DC storage for the processing of large neuroimaging data}

\author{\IEEEauthorblockN{Val\'erie Hayot-Sasson$^1$, Shawn T Brown$^2$ and 
    Tristan Glatard$^1$
  }\\
  \IEEEauthorblockA{
    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
    $^2$Montreal Neurological Institute, McGill University, Montreal, Canada
  }
}
\maketitle

\begin{abstract}
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
Neuroimaging open-data initiatives have led to extensively large repositories of
publicly available data. Such initiatives include the \bigbrain~\cite{BigBrain}, 
a one-of-a-kind $\SI{603}{\gibi\byte}$\valerie{verify this w/ Claude}
histological image of a 76 year-old~\valerie{from memory, so better double-check} healthy human brain
at 20$\mu$m resolution; the UK Biobank~\cite{ukbiobank}, a repository expected to
contain approximately $\SI{0.2}{\peta\byte}$ of data (including various magnetic
resonance (MR) imaging modalities) from 500,000 individuals living in the UK;
the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
1200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
initiative which aggregates MR data from various laboratories around the world, 
of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
data in total.

Due to storage limitations, only subsets of the neuroimaging repositories 
can be processed in a typical research laboratory. Moreover, as these datasets 
are extremely large and are only increasing in size, they are typically stored 
in higher-capacity, slower storage devices. In such conditions, large-scale 
studies in neuroimaging remain limited to labs with access to adequate 
infrastructure and are limited by data transfer times.

Furthermore, intermediary data is often required during results analysis. This limits
any performance benefits that can arise from volatile in-memory computing as intermediary
results need to be materialized onto persistent storage. As neuroimaging datasets continue
to increase in size, this movement of data will significantly increase the processing time
of neuroimaging data.

In this paper, we aim to:
\begin{itemize}
        \item Quantify the added value of Intel Optane persistent memory on 
            processing large neuroimaging data using representative pipelines; and
        \item Determine when a given Intel Optane storage configuration (Memory 
            and App Direct mode) is preferable.
\end{itemize}

\section{Materials and Methods}
The application pipelines, benchmarks, performance data, and analysis scripts used 
to implement the methods described hereafter are all available at 
\url{https://github.com/valhayot/paper-memory-storage} for 
further inspection and reproducibility. Links to the processing engines 
and processed data are provided in the text.\valerie{perhaps remove this for double blind}

\subsection{Infrastructure}

The server used in our experiments contain $\SI{3}{\tera\byte}$ of Intel Optane SSD storage, 700~GB of
DRAM..

\begin{itemize}
        \item amount of RAM

\end{itemize}

\subsection{Storage configuration}

\begin{itemize}
    \item Memory Mode
    \item App Direct Mode
\end{itemize}
\subsection{Performance model}

For the purposes of these experiments, we use the inequality introduced in
\cite{paperinmem}. That is,
\valerie{it's probably sufficient to just cite the reference, no?}
\begin{equation}                                                                 
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}                                                                                                                                  
\end{equation} 
where:
\begin{itemize}
    \item $D$ is the total amount of data written by the application
    \item $C$ is the total execution time of the application
    \item $\delta$ is the disk bandwidth
    \item $\gamma$ is the max number of concurrent processes on a node
\end{itemize}

Using this inequality, we anticipate that whenever the application is within the limits,
the storage selection will make little difference as storage devices with writeback enabled
will be able to leverage the memory cache. It is expected that Optane in App Direct mode
will be most useful when this inequality is not satisfied. As Memory mode will inevitably
increase the amount of data that can be cached by the storage, it is expected that the overall
performance of all storage devices, even when the inequality is not satisfied.

\subsection{Applications}
\subsubsection{ \bigbrain Incrementation}

\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}[1]
    \Input
        \Desc{$x$}{a sleep delay in seconds}
        \Desc{$n$}{a number of iterations}
        \Desc{$C$}{a set of image chunks}
        \Desc{$fs$}{filesystem to write to (tmpfs, Optane, local disk, Isilon)}
    \EndInput
    \ForEach{$chunk \in C$}
        \State read $chunk$ from $fs$
        \For{$i \in [1, n]$}
            \State $chunk\gets chunk+1$
            \State sleep $x$
        \EndFor
        \State save $chunk$ to $fs$
    \EndFor
    \end{algorithmic}
\end{algorithm}

Due to the size and uniqueness of the \bigbrain, standardized processing pipelines
have yet to be developed for it. In order to quantify the effects of data location
on such a dataset, we have implemented a naive synthetic application that takes 
the image, split into blocks, and increments all the voxels within each block, in parallel (Figure~\ref{alg:incrementation}).
This application
enables us to read and write to different storage devices and ensure that written
data could not have been previously cache
in-memory, by ensuring that read and written data are not the same. This application
will be parallelized using both GNU Parallel and Apache Spark 2.4.3 and implemented in Python 3.6. While the GNU Parallel application
is currently limited to 1 incrementation step, the Apache Spark implementation allows us to increase
the number of incrementation steps by a variable amount, thereby increasing the amount of intermediate
data generated.

We plan to execute this pipeline on both the $\SI{76}{\giga\byte}$ 40~$\mu$m 
\bigbrain split into 125 $\SI{617}{\mega\byte}$ blocks and the $\SI{603}{\giga\byte}$
\bigbrain at 20~$\mu$m split into 1000 $\SI{603}{\mega\byte}$ \valerie{verify} blocks.

At this moment, we have executed the application using GNU Parallel for parallelization
using 25 and 96 processes on the 125 40$\mu$m \bigbrain blocks. Data was read and written to either
tmpfs, Optane (App Direct only), local disk and Isilon. In all configurations (25 and 96 processes)
only 1 iteration was performed. At 25 processes, a sleep delay of 120 seconds was added at the
incrementation step, which would emulate tasks of longer duration and ensure that we satisfy Equation~\ref{eq:page-cache-inequality} for all devices.
At 96 processes with 0 sleep delay, we estimate being above the inequality for local disk and Isilon.
All experiments were repeated 5x in each environment (App Direct and Memory mode)

\subsubsection{ BIDS App Example}

The BIDS App Example is a template example for creating a Brain Imaging Data Structure (BIDS)
compliant application. It runs a standard neuroimaging brain extraction application (FSL BET) 
on all the anatomical (T1W) images of a dataset containing numerous subjects. This step is 
referred to as Participant analysis, within the BIDS App Example. An optional
step of the BIDS App example, referred to as Group analysis, computes the average brain
mask size of the entire dataset.

For our experiments, we used the entire CoRR dataset, made available on DataLad and
applied Participant analysis on it. As in the BigBrain Incrementation, the experiment parallelized 
using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory mode
using Isilon, local disk, Optane and tmpfs as storage. As participant analysis operates on only a subset of the dataset,
the inequality is satified for all conditions. Each experiment was repeated 3x.
\section{Results}

\subsection{\bigbrain incrementation}
\subsubsection{25 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/makespan-makespan_1it_25cpu_120delay.pdf}
    \captionsetup{width=\columnwidth}
    \caption{Makespan of \bigbrain Incrementation using 25 processes with a sleep delay of 120s on 125 blocks on the 40~$\mu$m image given different storage configurations. 5 repetitions were performed.}\label{fig:makespan-25cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/1it_25cpu_120delay.pdf}
    \captionsetup{width=\columnwidth}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 processes with a sleep delay of 120s on 125 blocks on the 40~$\mu$m image given different storage configurations. 5 repetitions were performed.}\label{fig:stacked-25cpus}
\end{figure}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1567709282-tmpfsAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{tmpfs}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1567705458-optaneAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{Optane}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1567704324-localAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{local disk}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1567707833-isilonAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{Isilon}\label{fig:gantt25isilon}
\end{subfigure}
\caption{Gantt charts for each storage device (App Direct mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes and a sleep delay of 120s}\label{fig:gantt25}
\end{figure*}

As shown in Figure~\ref{fig:makespan-25cpus}, despite being under the inequality, the
performance still varies according to selected storage device. As expected, Optane performs
slightly better than local disk and significantly better than Isilon. Tmpfs outperforms all
devices, with only slightly better performance than Optane and local disk, in both App Direct mode and Memory mode.
There is no apparent difference between the two modes on the given devices.

When observing the total task breakdown (Figure~\ref{fig:stacked-25cpus}), it is
apparent that the differences between the makespans arise from difference in incrementation
time between all devices. While tmpfs and local disk have comparable write times, Optane
has a more significant write time difference. Whereas, Isilon has more significant read and
write time compared to all other devices, it's incrementation time is comparable to that
of local disk. Further inspection of each storage device's Gantt chart (see Figure~\ref{fig:gantt25})
show that increased duration of tasks stagger the execution of each individual tasks. This, in turn,
results in better performance towards the end of the execution, where tasks are significantly more staggered.
Only tmpfs appears to be able to execute each task efficiently with little staggering.

\subsubsection{96 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/makespan-makespan_1it_96cpu_0delay.pdf}
    \captionsetup{width=\columnwidth}
    \caption{Makespan of \bigbrain Incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations. 5 repetitions were performed.}\label{fig:makespan-96cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/1it_96cpu_0delay.pdf}
    \captionsetup{width=\columnwidth}
    \caption{total read/increment/write breakdown of \bigbrain incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations. 5 repetitions were performed.}\label{fig:stacked-96cpus}
\end{figure}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1568150082-local_1it_96cpus_0delay_npc-3.pdf}
    \caption{Memory mode}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1567714252-localAD_1it_96cpus_0delay_npc-3.pdf}
    \caption{App Direct mode}
\end{subfigure}
\caption{Gantt charts for local disk processing 125 blocks of the 40$\mu$m BigBrain using 96 processes with no delay}\label{fig:gantt96}
\end{figure*}

At 96 processes, it was expected that the application would not satisfy the inequality.
In Figure~\ref{fig:makespan-96cpus} we see that there is now a greater disparity between the
makespans of experiments executed in Memory and App Direct mode, given the same storage. Local disk
in App Direct mode is significantly faster than its Memory mode counterpart. Local disk in App Direct
mode is even faster than Optane, by a factor of 1.7. Tmpfs was also found to perform better in App Direct mode, although
the performance difference is not significant. Only Isilon's performance did not vary between the
two execution modes.

As with 25 process, the main cause of performance variability appears to be the incrementation
time (Figure~\ref{fig:stacked-96cpus}. Local disk running in App Direct Mode has an average
incrementation time that is nearly 15~x faster than in Memory mode. The incrementation time
for local disk is, on average, 6~x faster than on Optane.

Isilon in App Direct generally performed the same as in memory mode, however, the write
was almost 4000~seconds slower, on average, in App Direct mode.

A further look at the Gantt charts between Memory mode and App Direct mode (Figure~\ref{fig:gantt96})
shows that the sole difference between the two modes is the duration of the incrementation step.
While both modes have an increased incrementation time, the incrementation duration in
Memory mode is much more important than that of App Direct mode.

\subsection{BIDS App Example}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bids-app-example/25cores.pdf}
    \caption{25 processes}\label{fig:bm25}
\end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bids-app-example/96cores.pdf}
    \caption{96 processes}\label{fig:bm96}
\end{subfigure}
\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bids-app-example/25cores-sum.pdf}
    \caption{25 processes}\label{fig:bb25}
\end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{../experiments/bids-app-example/96cores-sum.pdf}
    \caption{96 processes}\label{fig:bb96}
\end{subfigure}
\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}\label{fig:bbd}
\end{figure*}

Unlike the BigBrain incrementation at 25 Processes, the makespan of the BIDS App Example
executing on Optane is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
to the BigBrain Incrementation, tmpfs has a slight lead over local disk and Isilon performs the
worst. The same pattern can be found with 96 process (Figure~\ref{fig:bm96}). When comparing
the executions of 25 and 96 processes, the performance improved by a similar factor on all
storage devices.

The CPU and I/O breakdowns (Figure~\ref{fig:bbd}) show that there is a significant increase
in the time spent on I/O at 96 processes when compared to 25 processes. Furthermore, there appears
to also be an increase in the amount of time spent on CPU processing. In both cases,
tmpfs spends the least amount of time on I/O and CPU, followed by local disk, Optane, and finally,
Isilon.
\section{Discussion}
\subsection{Memory vs App Direct mode}

While Memory mode provides a cost-effective solution to increasing the total memory
size, it is slower than App Direct mode for embarrassingly parallel applications such
as \bigbrain incrementation. This is likely due to the Page Cache being composed of a
combination of both DRAM and Optane in Memory mode. Since the data produced by the application should be
able to fit within DRAM, writing to a slower bandwidth device such as Optane inevitably
deteriorates performance. As App Direct separates usage of DRAM and Optane, devices only utilize DRAM in this mode.
The only device that performed better in Memory mode was 
Isilon. This may be explained by a larger Page Cache in Memory mode, enabling Isilon to
store more data in memory prior to initiating synchronous writes to Isilon.

I/O-less task duration also seem to be affected by the execution mode, with those tasks
also performing significantly better in App Direct. A potential explanation for this may be 
that all memory accesses to Optane are slower due to the devices bandwidth, meaning that
tasks that do not perform I/O, but access data loaded into memory are slower in Memory mode.
This can be seen in Figure~\ref{fig:gantt96} where for local disk, incrementation in App Direct
mode is significantly shorter. Only when the memory bus is overwhelmed does incrementation take
longer for App Direct, however, it is nothing compared to the average duration of the incrementation
tasks in Memory mode.

Once again, Isilon is the only device which does not experience an increase in duration for
I/O-less tasks. In fact, the incrementation duration is less than that of local disk in Memory mode
and Optane in App Direct mode. It may be that since even the reads are staggered
(Figure~\ref{fig:gantt25isilon}), the memory bus is less strained by all the parallel processes,
therefore incurring less of an overall delay for all memory accesses.

\subsection{Effect of Page Cache}

As long as all the written data does not exceed the dirty ratio, there is little 
benefit brought by the use of Optane. This is the case with both
our applications writing less than $\SI{1}{\gibi\byte}$ to storage on a system with 
$\SI{768}{\gibi\byte}$ of DRAM and a dirty ratio of 20. However, it is expected 
that as the memory footprints of applications increase to the point where the kernel
would be forced to perform synchronous writes to disk if it were to rely on DRAM alone,
the benefit of using Optane in Memory mode will become more apparent.

However, depending on the neuroimaging application, the amount of data that could be
loaded in parallel would be limited by the amount of parallelism possible on the system.
For instance, typical fMRI tasks process many small images (100s of MB in size) in parallel.
While the datasets themselves can reach PB in size, the amount of data that can be processed in
parallel can be $\SI{10}{\giga\byte}$ if each take processes $\SI{100}{\mega\byte}$ and the 
maximum amount of parallelism is 100.

On the other hand, Memory mode can be very useful for applications operating on
ultra-high resolution images, like the \bigbrain. For instance, visualization applications
many need to load significant amounts of data in memory. Given the size of these datasets, 
extending the amount of available memory using Optane could be quite beneficial and allow
for faster visualization of larger datasets.

\subsection{Added value of persistent memory}

The added value of Optane a persistent storage device depend on the amount of parallelism
of the application. For instance, Optane performs better than local disk at 25 Processes,
but worse than local disk at 96 processes. This can be explained by the fact that it performs
I/O on the memory bus. While this may reduce overall latency to Optane, the bandwidth of 
writing to Optane is still significantly less than DRAM. Despite Optane having a similar bandwidth
to local disk, local disk uses DRAM as cache whereas Optane does not. This fact is 
particularly evident with the amount of data that is written in parallel at 96 processes. With 25 processes
and at least a 2 min
delay between each write, the load on the memory bus is low enough to be efficiently handled.

fMRI application will generally attempt to leverage as much parallelism as possible. As this
is the case, it is not believed that App Direct mode would bring a significant performance advantage
unless the server's CPUs are limited, in which case there is less concurrent accesses on the device.
The same would apply to ultra-high resolution application unless they are sequential, as may be the case with
visualization.
\section{Conclusion}

The benefits of Optane for neuroimaging applications are limited to large memory
accesses with limited parallelism, as would be the case for visualization. This is
the case for both Memory mode and App Direct mode.

While Memory mode leads to deterioration in performance of non-Optane storage devices,
they act as expected in App Direct mode. It is currently unknown by how much performance
of a non-Optane storage device would be affected by a large memory access.

To be able to fully determine the benefits of Optane for neuroimaging application,
it is necessary to process larger datasets. We plan to execute incrementation pipeline on the 
20~$\mu$m BigBrain using Apache Spark. 
\section{Acknowledgement}
\bibliographystyle{IEEEtran} 
\bibliography{biblio}

\end{document}

### 2019-09-16 experiments/bids-app-example/plot.py

@@ -2,30 +2,57 @@
 
 import os, json
 from matplotlib import pyplot as plt 
+from statistics import mean, stdev
 
-conditions_96_jobs = ["in-mem-1", "nvme-1", "pmem-1", "disk-1", "isilon-1"]
-conditions_25_jobs = ["in-mem-4", "nvme-4", "pmem-4", "disk-4", "isilon-4"]
-condition_ticks = ["tmpfs", "optane", "optaneAD", "disk", "isilon"]
+conditions_96_jobs = ["in-mem-1", "pmem-1", "disk-1", "isilon-1"]
+conditions_25_jobs = ["in-mem-4", "pmem-4", "disk-4", "isilon-4"]
+condition_ticks = ["tmpfs", "Optane", "disk", "Isilon"]
+width = 0.35
 
 def get_makespans(conditions):
-    makespans = []
+    mean_makespans = []
+    std_makespans = []
+
+    def get_reps(n_cpus):
+        start = '1'
+        ids = range(1, 4)
+        run_mksp = []
+
+        if n_cpus == 25:
+            start = '4'
+            ids = range(4, 7)
+
+        for i in ids:
+            with open(os.path.join(cond.replace(start, str(i)),"makespan_seconds")) as f:
+                m = int(f.read())
+            run_mksp.append(m)
+
+        return run_mksp
+
     for cond in conditions:
-        with open(os.path.join(cond,"makespan_seconds")) as f:
-            m = int(f.read())
-        makespans.append(m)
-    return makespans 
+        run_mksp = []
+        if '1' in cond:
+            run_mksp = get_reps(96) 
+        elif '4' in cond:
+            run_mksp = get_reps(25)
 
+        mean_makespans.append(mean(run_mksp))
+        std_makespans.append(stdev(run_mksp))
 
+    return mean_makespans, std_makespans 
+
+mksp_96 = get_makespans(conditions_96_jobs)
 plt.ylabel("Makespan (s)")
 plt.ylim((0,5000))
-plt.bar([1, 2, 3, 4, 5], get_makespans(conditions_96_jobs), tick_label=condition_ticks)
+plt.bar([1, 2, 3, 4], mksp_96[0], tick_label=condition_ticks, width=width, yerr=mksp_96[1])
 plt.savefig("96cores.pdf")
 
 plt.clf()
 
+mksp_25 = get_makespans(conditions_25_jobs)
[…]

### 2019-09-16 scripts/gantt.py

@@ -11,7 +11,7 @@ import time
 import argparse
 
 
-def gantt_increment(df, data_file):
+def gantt_increment(df, data_file, ncpus):
     color = {
         "read_file": "turquoise",
         "increment_file": "crimson",
@@ -28,9 +28,21 @@ def gantt_increment(df, data_file):
             data = r[1][["Start", "Duration"]]
             ax.broken_barh(data.values, (i - 0.4, 0.8), color=color[r[0]])
 
-    # ax.set_yticks(range(len(labels)))
-    # ax.set_yticklabels(labels)
+    r_read = mpatch.Patch(facecolor='turquoise', label='Read')
+    r_inc = mpatch.Patch(facecolor='crimson', label='Increment')
+    r_write = mpatch.Patch(facecolor='purple', label='Write')
+    ax.legend(handles=[r_read, r_inc, r_write], loc='upper center',
+              bbox_to_anchor=(0.5, 1.07), ncol=3, fancybox=True,
+              fontsize='x-small', framealpha=1)
+    ax.set_yticks(range(len(labels)))
+    ax.set_yticklabels([])
     ax.set_xlabel("time [s]")
+
+    if ncpus == 25:
+        ax.set_xlim(0, 1200)
+    else:
+        ax.set_xlim(0, 140)
+    ax.tick_params(axis='both', which='major', labelsize=8)
     plt.tight_layout()
     plt.savefig("gantt-{}.pdf".format(op.basename(data_file)))
 
@@ -76,7 +88,7 @@ def main():
     data_file = None
 
     if args.incrementation_tf is not None:
-        data_file = args["incrementation_tf"]
+        data_file = args.incrementation_tf
         df = pd.read_csv(
             data_file,
             delim_whitespace=True,
@@ -85,7 +97,7 @@ def main():
         df["Duration"] = df["End"] - df["Start"]
         df["Start"] = df["Start"] - df["Start"].min()
 
-        gantt_increment(df, data_file)
+        gantt_increment(df, data_file, args.cpus)
 
     else:
         data_file = args.bids_tf

### 2019-09-17 report/report.tex
fixed report figures, added figure generation script, created README for report

@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/makespan-makespan_1it_25cpu_120delay.pdf}
+    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/makespan-25cpus-120delay-1it.pdf}
     \captionsetup{width=\columnwidth}
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/1it_25cpu_120delay.pdf}
+    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/stacked-25cpus-120delay-1it.pdf}
     \captionsetup{width=\columnwidth}
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/makespan-makespan_1it_96cpu_0delay.pdf}
+    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/makespan-96cpus-0delay-1it.pdf}
     \captionsetup{width=\columnwidth}
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/1it_96cpu_0delay.pdf}
+    \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/stacked-96cpus-0delay-1it.pdf}
     \captionsetup{width=\columnwidth}

### 2019-09-17 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -77,7 +77,7 @@ def stacked_bar(df):
     ax.set_xticklabels(labels)
     ax.legend(handles=[r_mem, r_ad, r_read, r_inc, r_write])
 
-    plt.savefig('{}.pdf'.format(sys.argv[2]))
+    plt.savefig('stacked-{}.pdf'.format(sys.argv[2]))
     
 
 all_files = glob(op.abspath(sys.argv[1]))

### 2019-09-17 gen_figures.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+curr_dir=`pwd`
+cd ./experiments/bids-app-example
+./plot.py
+cd $curr_dir/experiments/bigbrain-incrementation/figures
+./makespan.py '../results/*25cpus*' 25cpus-120delay-1it
+./stacked_bar.py '../results/*25cpus*' 25cpus-120delay-1it
+./makespan.py '../results/*96cpus*' 96cpus-0delay-1it
+./stacked_bar.py '../results/*96cpus*' 96cpus-0delay-1it
+cd $curr_dir
+./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567704324-localAD_1it_25cpus_120delay_withpc-2.out --cpus 25
+./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567705458-optaneAD_1it_25cpus_120delay_withpc-2.out --cpus 25
+./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567707833-isilonAD_1it_25cpus_120delay_withpc-2.out --cpus 25
+./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567709282-tmpfsAD_1it_25cpus_120delay_withpc-2.out --cpus 25
+./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567714252-localAD_1it_96cpus_0delay_npc-3.out --cpus 96
+./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1568150082-local_1it_96cpus_0delay_npc-3.out --cpus 96
+mv ./gantt-*.pdf ./experiments/bigbrain-incrementation/figures
+

### 2019-09-17 scripts/gantt.py

@@ -44,7 +44,7 @@ def gantt_increment(df, data_file, ncpus):
         ax.set_xlim(0, 140)
     ax.tick_params(axis='both', which='major', labelsize=8)
     plt.tight_layout()
-    plt.savefig("gantt-{}.pdf".format(op.basename(data_file)))
+    plt.savefig("gantt-{}.pdf".format(op.basename(data_file).strip('.out')))
 
 
 def gantt_bids(df, data_file):

### 2019-09-17 report/report.tex
added biblio, abstract and some information on optane

@@
 \begin{abstract}
+    Open-access neuroimaging datasets have reached up to $\SI{}{\peta\byte}$ of data, and
+    continue to grow. The ability to leverage the entirety of these datasets are
+    limited to a restricted number of labs with both the time and infrastructure
+    to process the data. Whereas BigData engines have significantly reduced
+    application performance penalties with respect to data movement, their applied
+    strategies are not necessarily practical within neuroimaging workflows where
+    intermediary made need to be materialized to shared storage for post-processing
+    analysis. In this paper we evaluate the performance advantage brought by Intel
+    Optane DC persistent memory storage for the processing of large neuroimaging 
+    datasets using the two available configurations modes: Memory Mode and App Direct Mode.
+    We employ a synthetic algorithm on the $\SI{76}{\gibi\byte}$ \bigbrain
+    and a real neuroimaging application on the CoRR dataset using 25 and 96 parallel
+    processes in both cases. It was found that performance of all available storage
+    devices (e.g. tmpfs, local disk and Isilon) is somewhat deteriorated by memory mode
+    when compared to App Direct Mode. In contrast, Optane in App Direct Mode was
+    found to be slower than local disk, when the load on the bus was high due to sheer
+    amount of concurrent processes and short task durations. It is believed that
+    Optane would likely be useful in the accessing of typically larger-than-memory
+    data in a limited-parallelism manner, such as with visualization applications.
+    Further experiments with larger datasets remain to be performed.
+
 \end{abstract}
-
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
 
@@
 publicly available data. Such initiatives include the \bigbrain~\cite{BigBrain}, 
-a one-of-a-kind $\SI{603}{\gibi\byte}$\valerie{verify this w/ Claude}
-histological image of a 76 year-old~\valerie{from memory, so better double-check} healthy human brain
+a one-of-a-kind $\SI{603}{\gibi\byte}$
+histological image of a 65 year-old healthy human brain
 at 20$\mu$m resolution; the UK Biobank~\cite{ukbiobank}, a repository expected to
@@
 
+Whereas popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
+incorporated software solutions to limit data transfers (e.g. in-memory computing,
+data locality, and lazy evaluation), hardware has also adapted to the growing datasets.
+One such improvement is the concept of placing flash storage directly on the Dual
+In-line Memory Module (DIMM), thereby reducing the latency of accessing data on 
+flash storage devices. While the latency of these devices is improved, the bandwidth
+remains the same. However, as noted by \cite{nvdimms}, a severe performance degradation
+can be experienced by having memory traffic and I/O placed on the same bus.
+
+Intel Optane DC Persistent Memory storagei~\cite{optanebrief} is a recently introduced high-performance
+storage technology that also sits on the DIMMs to reduce latency to the device.
+Unlike DRAM DIMMs, it can accommodate twice the volume ($\SI{512}{\giga\byte}$) of
+storage enabling it to be a more cost effective alternative to DRAM storage.
+There are two principle configuration modes, namely Memory Mode and App Direct Mode,
+that enable the storage to either be accessed as an extension of available memory 
+or as a non-volatile memory storage device.
+
 In this paper, we aim to:
 \begin{itemize}
-        \item Quantify the added value of Intel Optane persistent memory on 
+        \item Quantify the added value of Intel Optane DC Persistent Memory on 
             processing large neuroimaging data using representative pipelines; and
         \item Determine when a given Intel Optane storage configuration (Memory 
-            and App Direct mode) is preferable.
+            and App Direct Mode) is preferable.
 \end{itemize}
@@
 \url{https://github.com/valhayot/paper-memory-storage} for 
-further inspection and reproducibility. Links to the processing engines 
-and processed data are provided in the text.\valerie{perhaps remove this for double blind}
+further inspection and reproducibility.
 
@@
 
-The server used in our experiments contain $\SI{3}{\tera\byte}$ of Intel Optane SSD storage, 700~GB of
-DRAM..
-
-\begin{itemize}
-        \item amount of RAM
-
-\end{itemize}
+The server used in our experiments contain $\SI{3}{\tera\byte}$ of Intel Optane Persistent memory, 768GB of DRAM,
+Isilon, local disk(SSD).
+CPU: Cascade lake
+OS: Red Hat 7.1
+\valerie{include storage bandwidth table here}
 
@@
 
-\begin{itemize}
-    \item Memory Mode
-    \item App Direct Mode
-\end{itemize}
+\subsubsection{Memory Mode}
+
+Memory Mode allows Intel Optane DC persistent memory to extend the system's available
+memory. In this mode, Optane uses DRAM as a cache and Optane can be accessed as
+volatile addressable main memory. By extending main memory, Memory Mode enables the
+fast access of large volumes of data.
+
+\subsubsection{App Direct Mode}
+
+App Direct Mode enables Optane to be accessed as a high-performance storage device.
+Unlike Memory Mode, the OS is able to differentiate between main memory and Optane
+storage accesses, treating them as two distinct entities. As Optane is located on
+the DIMMs, it exhibits lower latencies than traditional SSDs, while being able
+to be accessed like an SSD.
+
 \subsection{Performance model}
@@
 \cite{paperinmem}. That is,
-\valerie{it's probably sufficient to just cite the reference, no?}
 \begin{equation}                                                                 
@@
 the storage selection will make little difference as storage devices with writeback enabled
-will be able to leverage the memory cache. It is expected that Optane in App Direct mode
-will be most useful when this inequality is not satisfied. As Memory mode will inevitably
+will be able to leverage the memory cache. It is expected that Optane in App Direct Mode
+will be most useful when this inequality is not satisfied. As Memory Mode will inevitably
 increase the amount of data that can be cached by the storage, it is expected that the overall
@@
 
-We plan to execute this pipeline on both the $\SI{76}{\giga\byte}$ 40~$\mu$m 
-\bigbrain split into 125 $\SI{617}{\mega\byte}$ blocks and the $\SI{603}{\giga\byte}$
-\bigbrain at 20~$\mu$m split into 1000 $\SI{603}{\mega\byte}$ \valerie{verify} blocks.
+We plan to execute this pipeline on both the $\SI{76}{\gibi\byte}$ 40~$\mu$m 
+\bigbrain split into 125 $\SI{617}{\mega\byte}$ blocks and the $\SI{603}{\gibi\byte}$
+\bigbrain at 20~$\mu$m split into 1000 $\SI{617}{\mebi\byte}$ blocks.
 
@@
 At 96 processes with 0 sleep delay, we estimate being above the inequality for local disk and Isilon.
-All experiments were repeated 5x in each environment (App Direct and Memory mode)
+All experiments were repeated 5x in each environment (App Direct and Memory Mode)
 
@@
 applied Participant analysis on it. As in the BigBrain Incrementation, the experiment parallelized 
-using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory mode
+using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode
 using Isilon, local disk, Optane and tmpfs as storage. As participant analysis operates on only a subset of the dataset,
@@
 \end{subfigure}
-\caption{Gantt charts for each storage device (App Direct mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes and a sleep delay of 120s}\label{fig:gantt25}
+\caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes and a sleep delay of 120s}\label{fig:gantt25}
 \end{figure*}
@@
 slightly better than local disk and significantly better than Isilon. Tmpfs outperforms all
-devices, with only slightly better performance than Optane and local disk, in both App Direct mode and Memory mode.
+devices, with only slightly better performance than Optane and local disk, in both App Direct Mode and Memory Mode.
 There is no apparent difference between the two modes on the given devices.
@@
     \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1568150082-local_1it_96cpus_0delay_npc-3.pdf}
-    \caption{Memory mode}
+    \caption{Memory Mode}
 \end{subfigure}
@@
     \includegraphics[width=\columnwidth]{../experiments/bigbrain-incrementation/figures/gantt-1567714252-localAD_1it_96cpus_0delay_npc-3.pdf}
-    \caption{App Direct mode}
+    \caption{App Direct Mode}
 \end{subfigure}
@@
 In Figure~\ref{fig:makespan-96cpus} we see that there is now a greater disparity between the
-makespans of experiments executed in Memory and App Direct mode, given the same storage. Local disk
-in App Direct mode is significantly faster than its Memory mode counterpart. Local disk in App Direct
-mode is even faster than Optane, by a factor of 1.7. Tmpfs was also found to perform better in App Direct mode, although
+makespans of experiments executed in Memory and App Direct Mode, given the same storage. Local disk
+in App Direct Mode is significantly faster than its Memory Mode counterpart. Local disk in App Direct
+mode is even faster than Optane, by a factor of 1.7. Tmpfs was also found to perform better in App Direct Mode, although
 the performance difference is not significant. Only Isilon's performance did not vary between the
@@
 time (Figure~\ref{fig:stacked-96cpus}. Local disk running in App Direct Mode has an average
-incrementation time that is nearly 15~x faster than in Memory mode. The incrementation time
+incrementation time that is nearly 15~x faster than in Memory Mode. The incrementation time
 for local disk is, on average, 6~x faster than on Optane.
@@
 Isilon in App Direct generally performed the same as in memory mode, however, the write
-was almost 4000~seconds slower, on average, in App Direct mode.
+was almost 4000~seconds slower, on average, in App Direct Mode.
 
-A further look at the Gantt charts between Memory mode and App Direct mode (Figure~\ref{fig:gantt96})
+A further look at the Gantt charts between Memory Mode and App Direct Mode (Figure~\ref{fig:gantt96})
 shows that the sole difference between the two modes is the duration of the incrementation step.
 While both modes have an increased incrementation time, the incrementation duration in
-Memory mode is much more important than that of App Direct mode.
+Memory Mode is much more important than that of App Direct Mode.
 
@@
 \section{Discussion}
-\subsection{Memory vs App Direct mode}
+\subsection{Memory vs App Direct Mode}
 
-While Memory mode provides a cost-effective solution to increasing the total memory
-size, it is slower than App Direct mode for embarrassingly parallel applications such
+While Memory Mode provides a cost-effective solution to increasing the total memory
+size, it is slower than App Direct Mode for embarrassingly parallel applications such
 as \bigbrain incrementation. This is likely due to the Page Cache being composed of a
-combination of both DRAM and Optane in Memory mode. Since the data produced by the application should be
+combination of both DRAM and Optane in Memory Mode. Since the data produced by the application should be
 able to fit within DRAM, writing to a slower bandwidth device such as Optane inevitably
 deteriorates performance. As App Direct separates usage of DRAM and Optane, devices only utilize DRAM in this mode.
-The only device that performed better in Memory mode was 
-Isilon. This may be explained by a larger Page Cache in Memory mode, enabling Isilon to
+The only device that performed better in Memory Mode was 
+Isilon. This may be explained by a larger Page Cache in Memory Mode, enabling Isilon to
 store more data in memory prior to initiating synchronous writes to Isilon.
@@
 that all memory accesses to Optane are slower due to the devices bandwidth, meaning that
-tasks that do not perform I/O, but access data loaded into memory are slower in Memory mode.
+tasks that do not perform I/O, but access data loaded into memory are slower in Memory Mode.
 This can be seen in Figure~\ref{fig:gantt96} where for local disk, incrementation in App Direct
@@
 longer for App Direct, however, it is nothing compared to the average duration of the incrementation
-tasks in Memory mode.
+tasks in Memory Mode.
 
 Once again, Isilon is the only device which does not experience an increase in duration for
-I/O-less tasks. In fact, the incrementation duration is less than that of local disk in Memory mode
-and Optane in App Direct mode. It may be that since even the reads are staggered
+I/O-less tasks. In fact, the incrementation duration is less than that of local disk in Memory Mode
+and Optane in App Direct Mode. It may be that since even the reads are staggered
 (Figure~\ref{fig:gantt25isilon}), the memory bus is less strained by all the parallel processes,
@@
 would be forced to perform synchronous writes to disk if it were to rely on DRAM alone,
-the benefit of using Optane in Memory mode will become more apparent.
+the benefit of using Optane in Memory Mode will become more apparent.
 
@@
 
-On the other hand, Memory mode can be very useful for applications operating on
+On the other hand, Memory Mode can be very useful for applications operating on
 ultra-high resolution images, like the \bigbrain. For instance, visualization applications
@@
 fMRI application will generally attempt to leverage as much parallelism as possible. As this
-is the case, it is not believed that App Direct mode would bring a significant performance advantage
+is the case, it is not believed that App Direct Mode would bring a significant performance advantage
 unless the server's CPUs are limited, in which case there is less concurrent accesses on the device.
@@
 accesses with limited parallelism, as would be the case for visualization. This is
-the case for both Memory mode and App Direct mode.
+the case for both Memory Mode and App Direct Mode.
 
-While Memory mode leads to deterioration in performance of non-Optane storage devices,
-they act as expected in App Direct mode. It is currently unknown by how much performance
+While Memory Mode leads to deterioration in performance of non-Optane storage devices,
+they act as expected in App Direct Mode. It is currently unknown by how much performance
 of a non-Optane storage device would be affected by a large memory access.

### 2019-09-18 report/report.tex

@@
       #1 \color{black}}
+\newcommand{\tristan}[1]{\color{orange}\textbf{Note from Tristan}:
+      #1 \color{black}}
 
@@
 
-\title{Performance benefits of Intel Optane DC storage for the processing of large neuroimaging data}
+\title{Performance benefits of Intel Optane DC storage for the parallel processing of large neuroimaging data}
 
@@
     continue to grow. The ability to leverage the entirety of these datasets are
-    limited to a restricted number of labs with both the time and infrastructure
+    limited to a restricted number of labs with both the capacity and infrastructure
     to process the data. Whereas BigData engines have significantly reduced
@@
     strategies are not necessarily practical within neuroimaging workflows where
-    intermediary made need to be materialized to shared storage for post-processing
+    intermediary results may need to be materialized to shared storage for post-processing
     analysis. In this paper we evaluate the performance advantage brought by Intel
-    Optane DC persistent memory storage for the processing of large neuroimaging 
+    Optane DC \tristan{expan acronym?} persistent memory storage for the processing of large neuroimaging 
     datasets using the two available configurations modes: Memory Mode and App Direct Mode.
     We employ a synthetic algorithm on the $\SI{76}{\gibi\byte}$ \bigbrain
-    and a real neuroimaging application on the CoRR dataset using 25 and 96 parallel
-    processes in both cases. It was found that performance of all available storage
+    and a real neuroimaging application on the CoRR \tristan{expand acronym} dataset using 25 and 96 parallel
+    processes in both cases. Our results show that performance of all available storage
     devices (e.g. tmpfs, local disk and Isilon) is somewhat deteriorated by memory mode
@@
 the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
-1200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
+1,200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
 and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
-initiative which aggregates MR data from various laboratories around the world, 
+initiative which aggregates MR data from various centres around the world, 
 of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
@@
 
-Due to storage limitations, only subsets of the neuroimaging repositories 
+Due to storage limitations, only subsets of such neuroimaging repositories 
 can be processed in a typical research laboratory. Moreover, as these datasets 
 are extremely large and are only increasing in size, they are typically stored 
-in higher-capacity, slower storage devices. In such conditions, large-scale 
+in higher-capacity, slower storage devices \tristan{such as?}. In such conditions, large-scale 
 studies in neuroimaging remain limited to labs with access to adequate 
@@
 remains the same. However, as noted by \cite{nvdimms}, a severe performance degradation
-can be experienced by having memory traffic and I/O placed on the same bus.
+can be experienced by having memory traffic and I/O placed on the same bus \tristan{references are
+missing or the bib file hasn't been updated}.
 
@@
 \end{itemize}
+
+\tristan{awesome intro Valerie, great context, right level of detail, nice 
+parallel between software and hardware solutions for big data, super well informed.}
 
@@
 
+\tristan{I guess tmpfs, Isilon and local disks should be described here.}
+
 \subsection{Performance model}
@@
 For the purposes of these experiments, we use the inequality introduced in
-\cite{paperinmem}. That is,
+\cite{paperinmem} to characterize data-intensive applications. That is,
 \begin{equation}                                                                 
@@
 Using this inequality, we anticipate that whenever the application is within the limits,
-the storage selection will make little difference as storage devices with writeback enabled
+the storage selection will make little difference as storage devices with writeback enabled \tristan{This should be explained 
+before, in methods (infrastructure) and possibly in intro.}
 will be able to leverage the memory cache. It is expected that Optane in App Direct Mode
-will be most useful when this inequality is not satisfied. As Memory Mode will inevitably
+will be most useful when this inequality is not satisfied \tristan{why?}. As Memory Mode will inevitably
 increase the amount of data that can be cached by the storage, it is expected that the overall
-performance of all storage devices, even when the inequality is not satisfied.
+performance of all storage devices \tristan{will what?}, even when the inequality is not satisfied.
 
@@
 on such a dataset, we have implemented a naive synthetic application that takes 
-the image, split into blocks, and increments all the voxels within each block, in parallel (Figure~\ref{alg:incrementation}).
+the image, split into blocks, and increments all the voxels within each block, in parallel (Algorithm~\ref{alg:incrementation}).
 This application
@@
 in-memory, by ensuring that read and written data are not the same. This application
-will be parallelized using both GNU Parallel and Apache Spark 2.4.3 and implemented in Python 3.6. While the GNU Parallel application
+will be parallelized using both GNU Parallel and Apache Spark 2.4.3 and implemented in Python 3.6 \tristan{In a subsequent 
+version you could give more details about implementations, including containerization, etc}. While the GNU Parallel application
 is currently limited to 1 incrementation step, the Apache Spark implementation allows us to increase
@@
 
-\subsubsection{ BIDS App Example}
+\subsubsection{BIDS App Example}
 
@@
 There is no apparent difference between the two modes on the given devices.
+\tristan{You should explain why Optane is not reported in memory mode.}
 
@@
 apparent that the differences between the makespans arise from difference in incrementation
-time between all devices. While tmpfs and local disk have comparable write times, Optane
+time between all devices except Isilon. In addition, while tmpfs and local disk have comparable write times, Optane
 has a more significant write time difference. Whereas, Isilon has more significant read and
@@
 Only tmpfs appears to be able to execute each task efficiently with little staggering.
-
+\tristan{In the methods, you should say how time was measured.}
+\tristan{Is Isilon in writeback? If so, why is write time so large?}
 \subsubsection{96 Processes}
@@
 As with 25 process, the main cause of performance variability appears to be the incrementation
-time (Figure~\ref{fig:stacked-96cpus}. Local disk running in App Direct Mode has an average
+time (Figure~\ref{fig:stacked-96cpus}). Local disk running in App Direct Mode has an average
 incrementation time that is nearly 15~x faster than in Memory Mode. The incrementation time
@@
 \end{figure*}
-
+\tristan{I/O time might include sleep time if the process is throttled, we should explain how I/O and CPU were measured.}
 Unlike the BigBrain incrementation at 25 Processes, the makespan of the BIDS App Example
@@
 The only device that performed better in Memory Mode was 
-Isilon. This may be explained by a larger Page Cache in Memory Mode, enabling Isilon to
+Isilon \tristan{But the difference looks very small}. This may be explained by a larger Page Cache in Memory Mode, enabling Isilon to
 store more data in memory prior to initiating synchronous writes to Isilon.
@@
 longer for App Direct, however, it is nothing compared to the average duration of the incrementation
-tasks in Memory Mode.
+tasks in Memory Mode. \tristan{how about throttling?}
 

### 2019-09-19 report/report.tex
addressed @glatard's comments

@@
     analysis. In this paper we evaluate the performance advantage brought by Intel
-    Optane DC \tristan{expan acronym?} persistent memory storage for the processing of large neuroimaging 
+    Optane DC persistent memory storage for the processing of large neuroimaging 
     datasets using the two available configurations modes: Memory Mode and App Direct Mode.
     We employ a synthetic algorithm on the $\SI{76}{\gibi\byte}$ \bigbrain
-    and a real neuroimaging application on the CoRR \tristan{expand acronym} dataset using 25 and 96 parallel
+    and a real neuroimaging application on the Consortium for Reliability and Reproducibility (CoRR)
+    dataset using 25 and 96 parallel
     processes in both cases. Our results show that performance of all available storage
@@
     found to be slower than local disk, when the load on the bus was high due to sheer
-    amount of concurrent processes and short task durations. It is believed that
+    amount of concurrent processes and short task duration. It is believed that
     Optane would likely be useful in the accessing of typically larger-than-memory
@@
 are extremely large and are only increasing in size, they are typically stored 
-in higher-capacity, slower storage devices \tristan{such as?}. In such conditions, large-scale 
+in higher-capacity, slower storage devices, such as hard disk drives, or external
+parallel file systems, such as Lustre. In such conditions, large-scale 
 studies in neuroimaging remain limited to labs with access to adequate 
-infrastructure and are limited by data transfer times.
+infrastructure.
 
-Furthermore, intermediary data is often required during results analysis. This limits
+Furthermore, intermediary data is often required for post-processing analysis. This limits
 any performance benefits that can arise from volatile in-memory computing as intermediary
 results need to be materialized onto persistent storage. As neuroimaging datasets continue
-to increase in size, this movement of data will significantly increase the processing time
-of neuroimaging data.
+to increase in size, the movement of data will significantly increase the processing time.
 
-Whereas popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
-incorporated software solutions to limit data transfers (e.g. in-memory computing,
+To mitigate the effects of data writes, kernel-based
+strategies,such as the writeback cache, have been developed. The writeback cache
+allows processes to use the memory as a cache for writing. The cache size,
+however,
+is configurable, but nevertheless limited. When writes approach the cache's capacity,
+processes performing writes start to be throttled. After capacity is reached,
+processes can no longer use the cache until all written data on cache is flushed
+to the appropriate storage device.
+
+Whereas Operating Systems and popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
+incorporated software solutions to limit data transfers (e.g. writeback cache, in-memory computing,
 data locality, and lazy evaluation), hardware has also adapted to the growing datasets.
@@
 remains the same. However, as noted by \cite{nvdimms}, a severe performance degradation
-can be experienced by having memory traffic and I/O placed on the same bus \tristan{references are
-missing or the bib file hasn't been updated}.
+can be experienced by having memory traffic and I/O placed on the same bus.
 
-Intel Optane DC Persistent Memory storagei~\cite{optanebrief} is a recently introduced high-performance
+Intel Optane DC Persistent Memory storage~\cite{optanebrief} is a recently introduced high-performance
 storage technology that also sits on the DIMMs to reduce latency to the device.
@@
 
-\tristan{awesome intro Valerie, great context, right level of detail, nice 
-parallel between software and hardware solutions for big data, super well informed.}
-
 \section{Materials and Methods}
@@
 OS: Red Hat 7.1
+writeback enabled for local disk (unsure about Isilon)
 \valerie{include storage bandwidth table here}
@@
 
+As Memory Mode would affect the processing time regardless of the device
+selected, we evaluated Memory Mode on tmpfs, local disk and Isilon. tmpfs
+consisted of $\SI{298}{\giga\byte}$ and was located in \texttt{/run/user/61218}.
+Local disk comprised of a $\SI{240}{\giga\byte}$ Micron 5100 PRO SSD, of which
+$\SI{150}{\giga\byte}$ were mounted on \texttt{/home}. The $\SI{628}{\tera\byte}$
+Isilon was mounted as NFS4 to \texttt{/home/users}
+
+Optane was not evaluated separately as it is indistinct from DRAM memory in
+Memory Mode
 \subsubsection{App Direct Mode}
@@
 Unlike Memory Mode, the OS is able to differentiate between main memory and Optane
-storage accesses, treating them as two distinct entities. As Optane is located on
+storage, treating them as two distinct entities. As Optane is located on
 the DIMMs, it exhibits lower latencies than traditional SSDs, while being able
-to be accessed like an SSD.
+to be accessed like an SSD. Optane does not use the DRAM as cache in App Direct Mode.
 
-\tristan{I guess tmpfs, Isilon and local disks should be described here.}
+The same storage devices were used to evaluate App Direct Mode, with the inclusion
+of Intel Optane DC Persistent Memory, of which $\SI{1.5}{\tera\byte}$ of storage
+were mounted on \texttt{/pvme-1}.
 
@@
 Using this inequality, we anticipate that whenever the application is within the limits,
-the storage selection will make little difference as storage devices with writeback enabled \tristan{This should be explained 
-before, in methods (infrastructure) and possibly in intro.}
+the storage selection will make little difference, as storage devices with writeback enabled 
 will be able to leverage the memory cache. It is expected that Optane in App Direct Mode
-will be most useful when this inequality is not satisfied \tristan{why?}. As Memory Mode will inevitably
+will be most useful when this inequality is not satisfied as storage devices will
+not be able to use the writeback cache. As Memory Mode will inevitably
 increase the amount of data that can be cached by the storage, it is expected that the overall
-performance of all storage devices \tristan{will what?}, even when the inequality is not satisfied.
+performance of all storage devices improve, even when the inequality is not satisfied.
 
@@
 
+Time is measured within the application using Python's \texttt{time} module 
+called before and after any reads, increments or writes.
+
 We plan to execute this pipeline on both the $\SI{76}{\gibi\byte}$ 40~$\mu$m 
@@
 the inequality is satified for all conditions. Each experiment was repeated 3x.
+
+Time in this application was obtained using Linux's \texttt{time(1)} application.
+\texttt{Real} was used to measure CPU time, where \texttt{Sys} was used to measure
+I/O time.
 \section{Results}
@@
 Only tmpfs appears to be able to execute each task efficiently with little staggering.
-\tristan{In the methods, you should say how time was measured.}
-\tristan{Is Isilon in writeback? If so, why is write time so large?}
 \subsubsection{96 Processes}
@@
 \end{figure*}
-\tristan{I/O time might include sleep time if the process is throttled, we should explain how I/O and CPU were measured.}
 Unlike the BigBrain incrementation at 25 Processes, the makespan of the BIDS App Example
@@
 The only device that performed better in Memory Mode was 
-Isilon \tristan{But the difference looks very small}. This may be explained by a larger Page Cache in Memory Mode, enabling Isilon to
-store more data in memory prior to initiating synchronous writes to Isilon.
+Isilon, however the difference is minimal.
+%This may be explained by a larger Page Cache in Memory Mode, enabling Isilon to
+%store more data in memory prior to initiating synchronous writes to Isilon.
 
@@
 also performing significantly better in App Direct. A potential explanation for this may be 
-that all memory accesses to Optane are slower due to the devices bandwidth, meaning that
-tasks that do not perform I/O, but access data loaded into memory are slower in Memory Mode.
+that all memory accesses to Optane are slower due to the device's bandwidth, meaning that
+tasks that do not perform I/O, but access data loaded into memory are slower in Memory Mode. Another explanation might be that the kernel is throttling the processes
+as they are producing significant amounts of output, thereby increasing the
+incrementation time.
 This can be seen in Figure~\ref{fig:gantt96} where for local disk, incrementation in App Direct
@@
 longer for App Direct, however, it is nothing compared to the average duration of the incrementation
-tasks in Memory Mode. \tristan{how about throttling?}
+tasks in Memory Mode.
 

### 2019-09-19 paper.tex
remove report folder, fixed readme

--- 2019-08-31 paper.tex
@@
 \usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\algnewcommand\algorithmicforeach{\textbf{for each}}
+\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
+\algblock{Input}{EndInput}
+\algnotext{EndInput}
+\algblock{Output}{EndOutput}
+\algnotext{EndOutput}
+\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}
 \usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 \usepackage{textcomp}
 \usepackage{xcolor}
+\usepackage{hyperref}
+\usepackage[binary-units]{siunitx}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\newcommand{\bigdata}{Big Data }
+\newcommand{\bigbrain}{BigBrain }
+
+\newcommand{\valerie}[1]{\color{blue}\textbf{Note from Valerie}:
+      #1 \color{black}}
+\newcommand{\tristan}[1]{\color{orange}\textbf{Note from Tristan}:
+      #1 \color{black}}
+
 \begin{document}
 
-\title{Paper Title*}
+\title{Performance benefits of Intel Optane DC storage for the parallel processing of large neuroimaging data}
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\author{\IEEEauthorblockN{Val\'erie Hayot-Sasson$^1$, Shawn T Brown$^2$ and 
+    Tristan Glatard$^1$
+  }\\
+  \IEEEauthorblockA{
+    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
+    $^2$Montreal Neurological Institute, McGill University, Montreal, Canada
   }
-
+}
 \maketitle
@@
 \begin{abstract}
+    Open-access neuroimaging datasets have reached up to $\SI{}{\peta\byte}$ of data, and
+    continue to grow. The ability to leverage the entirety of these datasets are
+    limited to a restricted number of labs with both the capacity and infrastructure
+    to process the data. Whereas BigData engines have significantly reduced
+    application performance penalties with respect to data movement, their applied
+    strategies are not necessarily practical within neuroimaging workflows where
+    intermediary results may need to be materialized to shared storage for post-processing
+    analysis. In this paper we evaluate the performance advantage brought by Intel
+    Optane DC persistent memory storage for the processing of large neuroimaging 
+    datasets using the two available configurations modes: Memory Mode and App Direct Mode.
+    We employ a synthetic algorithm on the $\SI{76}{\gibi\byte}$ \bigbrain
+    and a real neuroimaging application on the Consortium for Reliability and Reproducibility (CoRR)
+    dataset using 25 and 96 parallel
+    processes in both cases. Our results show that performance of all available storage
+    devices (e.g. tmpfs, local disk and Isilon) is somewhat deteriorated by memory mode
+    when compared to App Direct Mode. In contrast, Optane in App Direct Mode was
+    found to be slower than local disk, when the load on the bus was high due to sheer
+    amount of concurrent processes and short task duration. It is believed that
+    Optane would likely be useful in the accessing of typically larger-than-memory
+    data in a limited-parallelism manner, such as with visualization applications.
+    Further experiments with larger datasets remain to be performed.
+
 \end{abstract}
 
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
+\section{Introduction}
+Neuroimaging open-data initiatives have led to extensively large repositories of
+publicly available data. Such initiatives include the \bigbrain~\cite{BigBrain}, 
+a one-of-a-kind $\SI{603}{\gibi\byte}$
+histological image of a 65 year-old healthy human brain
+at 20$\mu$m resolution; the UK Biobank~\cite{ukbiobank}, a repository expected to
+contain approximately $\SI{0.2}{\peta\byte}$ of data (including various magnetic
+resonance (MR) imaging modalities) from 500,000 individuals living in the UK;
+the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
+1,200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
+and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
+initiative which aggregates MR data from various centres around the world, 
+of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
+data in total.
 
-\section{Introduction}
-\section{Methods}
+Due to storage limitations, only subsets of such neuroimaging repositories 
+can be processed in a typical research laboratory. Moreover, as these datasets 
+are extremely large and are only increasing in size, they are typically stored 
+in higher-capacity, slower storage devices, such as hard disk drives, or external
+parallel file systems, such as Lustre. In such conditions, large-scale 
+studies in neuroimaging remain limited to labs with access to adequate 
+infrastructure.
+
+Furthermore, intermediary data is often required for post-processing analysis. This limits
+any performance benefits that can arise from volatile in-memory computing as intermediary
+results need to be materialized onto persistent storage. As neuroimaging datasets continue
+to increase in size, the movement of data will significantly increase the processing time.
+
+To mitigate the effects of data writes, kernel-based
+strategies,such as the writeback cache, have been developed. The writeback cache
+allows processes to use the memory as a cache for writing. The cache size,
+however,
+is configurable, but nevertheless limited. When writes approach the cache's capacity,
+processes performing writes start to be throttled. After capacity is reached,
+processes can no longer use the cache until all written data on cache is flushed
+to the appropriate storage device.
+
+Whereas Operating Systems and popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
+incorporated software solutions to limit data transfers (e.g. writeback cache, in-memory computing,
+data locality, and lazy evaluation), hardware has also adapted to the growing datasets.
+One such improvement is the concept of placing flash storage directly on the Dual
+In-line Memory Module (DIMM), thereby reducing the latency of accessing data on 
+flash storage devices. While the latency of these devices is improved, the bandwidth
+remains the same. However, as noted by \cite{nvdimms}, a severe performance degradation
+can be experienced by having memory traffic and I/O placed on the same bus.
+
+Intel Optane DC Persistent Memory storage~\cite{optanebrief} is a recently introduced high-performance
+storage technology that also sits on the DIMMs to reduce latency to the device.
+Unlike DRAM DIMMs, it can accommodate twice the volume ($\SI{512}{\giga\byte}$) of
+storage enabling it to be a more cost effective alternative to DRAM storage.
+There are two principle configuration modes, namely Memory Mode and App Direct Mode,
+that enable the storage to either be accessed as an extension of available memory 
+or as a non-volatile memory storage device.
+
+In this paper, we aim to:
+\begin{itemize}
+        \item Quantify the added value of Intel Optane DC Persistent Memory on 
+            processing large neuroimaging data using representative pipelines; and
+        \item Determine when a given Intel Optane storage configuration (Memory 
+            and App Direct Mode) is preferable.
+\end{itemize}
+
+\section{Materials and Methods}
+The application pipelines, benchmarks, performance data, and analysis scripts used 
+to implement the methods described hereafter are all available at 
+\url{https://github.com/valhayot/paper-memory-storage} for 
+further inspection and reproducibility.
+
+\subsection{Infrastructure}
+
+The server used in our experiments contain $\SI{3}{\tera\byte}$ of Intel Optane Persistent memory, 768GB of DRAM,
+Isilon, local disk(SSD).
+CPU: Cascade lake
+OS: Red Hat 7.1
+writeback enabled for local disk (unsure about Isilon)
+\valerie{include storage bandwidth table here}
+
+\subsection{Storage configuration}
+
+\subsubsection{Memory Mode}
+
+Memory Mode allows Intel Optane DC persistent memory to extend the system's available
+memory. In this mode, Optane uses DRAM as a cache and Optane can be accessed as
+volatile addressable main memory. By extending main memory, Memory Mode enables the
+fast access of large volumes of data.
+
+As Memory Mode would affect the processing time regardless of the device
+selected, we evaluated Memory Mode on tmpfs, local disk and Isilon. tmpfs
+consisted of $\SI{298}{\giga\byte}$ and was located in \texttt{/run/user/61218}.
+Local disk comprised of a $\SI{240}{\giga\byte}$ Micron 5100 PRO SSD, of which
+$\SI{150}{\giga\byte}$ were mounted on \texttt{/home}. The $\SI{628}{\tera\byte}$
+Isilon was mounted as NFS4 to \texttt{/home/users}
+
+Optane was not evaluated separately as it is indistinct from DRAM memory in
+Memory Mode
+\subsubsection{App Direct Mode}
+
+App Direct Mode enables Optane to be accessed as a high-performance storage device.
+Unlike Memory Mode, the OS is able to differentiate between main memory and Optane
+storage, treating them as two distinct entities. As Optane is located on
+the DIMMs, it exhibits lower latencies than traditional SSDs, while being able
+to be accessed like an SSD. Optane does not use the DRAM as cache in App Direct Mode.
+
+The same storage devices were used to evaluate App Direct Mode, with the inclusion
+of Intel Optane DC Persistent Memory, of which $\SI{1.5}{\tera\byte}$ of storage
+were mounted on \texttt{/pvme-1}.
+
+\subsection{Performance model}
+
+For the purposes of these experiments, we use the inequality introduced in
+\cite{paperinmem} to characterize data-intensive applications. That is,
+\begin{equation}                                                                 
+\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
+\end{equation} 
+where:
+\begin{itemize}
+    \item $D$ is the total amount of data written by the application
+    \item $C$ is the total execution time of the application
+    \item $\delta$ is the disk bandwidth
+    \item $\gamma$ is the max number of concurrent processes on a node
+\end{itemize}
+
+Using this inequality, we anticipate that whenever the application is within the limits,
+the storage selection will make little difference, as storage devices with writeback enabled 
+will be able to leverage the memory cache. It is expected that Optane in App Direct Mode
+will be most useful when this inequality is not satisfied as storage devices will
+not be able to use the writeback cache. As Memory Mode will inevitably
+increase the amount of data that can be cached by the storage, it is expected that the overall
+performance of all storage devices improve, even when the inequality is not satisfied.
+
+\subsection{Applications}
+\subsubsection{ \bigbrain Incrementation}
+
+\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
+    \begin{algorithmic}[1]
+    \Input
+        \Desc{$x$}{a sleep delay in seconds}
+        \Desc{$n$}{a number of iterations}
+        \Desc{$C$}{a set of image chunks}
+        \Desc{$fs$}{filesystem to write to (tmpfs, Optane, local disk, Isilon)}
+    \EndInput
+    \ForEach{$chunk \in C$}
+        \State read $chunk$ from $fs$
+        \For{$i \in [1, n]$}
+            \State $chunk\gets chunk+1$
+            \State sleep $x$
+        \EndFor
+        \State save $chunk$ to $fs$
+    \EndFor
+    \end{algorithmic}
+\end{algorithm}
+
+Due to the size and uniqueness of the \bigbrain, standardized processing pipelines
+have yet to be developed for it. In order to quantify the effects of data location
+on such a dataset, we have implemented a naive synthetic application that takes 
+the image, split into blocks, and increments all the voxels within each block, in parallel (Algorithm~\ref{alg:incrementation}).
+This application
+enables us to read and write to different storage devices and ensure that written
+data could not have been previously cache
+in-memory, by ensuring that read and written data are not the same. This application
+will be parallelized using both GNU Parallel and Apache Spark 2.4.3 and implemented in Python 3.6 \tristan{In a subsequent 
+version you could give more details about implementations, including containerization, etc}. While the GNU Parallel application
+is currently limited to 1 incrementation step, the Apache Spark implementation allows us to increase
+the number of incrementation steps by a variable amount, thereby increasing the amount of intermediate
+data generated.
+
+Time is measured within the application using Python's \texttt{time} module 
+called before and after any reads, increments or writes.
+
+We plan to execute this pipeline on both the $\SI{76}{\gibi\byte}$ 40~$\mu$m 
+\bigbrain split into 125 $\SI{617}{\mega\byte}$ blocks and the $\SI{603}{\gibi\byte}$
+\bigbrain at 20~$\mu$m split into 1000 $\SI{617}{\mebi\byte}$ blocks.
+
+At this moment, we have executed the application using GNU Parallel for parallelization
+using 25 and 96 processes on the 125 40$\mu$m \bigbrain blocks. Data was read and written to either
+tmpfs, Optane (App Direct only), local disk and Isilon. In all configurations (25 and 96 processes)
+only 1 iteration was performed. At 25 processes, a sleep delay of 120 seconds was added at the
+incrementation step, which would emulate tasks of longer duration and ensure that we satisfy Equation~\ref{eq:page-cache-inequality} for all devices.
+At 96 processes with 0 sleep delay, we estimate being above the inequality for local disk and Isilon.
+All experiments were repeated 5x in each environment (App Direct and Memory Mode)
+
+\subsubsection{BIDS App Example}
+
+The BIDS App Example is a template example for creating a Brain Imaging Data Structure (BIDS)
+compliant application. It runs a standard neuroimaging brain extraction application (FSL BET) 
+on all the anatomical (T1W) images of a dataset containing numerous subjects. This step is 
+referred to as Participant analysis, within the BIDS App Example. An optional
+step of the BIDS App example, referred to as Group analysis, computes the average brain
+mask size of the entire dataset.
+
+For our experiments, we used the entire CoRR dataset, made available on DataLad and
+applied Participant analysis on it. As in the BigBrain Incrementation, the experiment parallelized 
+using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode
+using Isilon, local disk, Optane and tmpfs as storage. As participant analysis operates on only a subset of the dataset,
+the inequality is satified for all conditions. Each experiment was repeated 3x.
+
+Time in this application was obtained using Linux's \texttt{time(1)} application.
+\texttt{Real} was used to measure CPU time, where \texttt{Sys} was used to measure
+I/O time.
 \section{Results}
+
+\subsection{\bigbrain incrementation}
+\subsubsection{25 Processes}
+\begin{figure}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-25cpus-120delay-1it.pdf}
+    \captionsetup{width=\columnwidth}
+    \caption{Makespan of \bigbrain Incrementation using 25 processes with a sleep delay of 120s on 125 blocks on the 40~$\mu$m image given different storage configurations.
+5 repetitions were performed.}\label{fig:makespan-25cpus}
+\end{figure}
+\begin{figure}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-25cpus-120delay-1it.pdf}
+    \captionsetup{width=\columnwidth}
+    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 processes with a sleep delay of 120s on 125 blocks on the 40~$\mu$m image given different storage configurations.
+5 repetitions were performed.}\label{fig:stacked-25cpus}
+\end{figure}
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567709282-tmpfsAD_1it_25cpus_120delay_withpc-2.pdf}
+    \caption{tmpfs}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567705458-optaneAD_1it_25cpus_120delay_withpc-2.pdf}
+    \caption{Optane}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567704324-localAD_1it_25cpus_120delay_withpc-2.pdf}
+    \caption{local disk}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567707833-isilonAD_1it_25cpus_120delay_withpc-2.pdf}
+    \caption{Isilon}\label{fig:gantt25isilon}
+\end{subfigure}
+\caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes and a sleep delay of 120s}\label{fig:gantt25}
+\end{figure*}
+
+As shown in Figure~\ref{fig:makespan-25cpus}, despite being under the inequality, the
+performance still varies according to selected storage device. As expected, Optane performs
+slightly better than local disk and significantly better than Isilon. Tmpfs outperforms all
+devices, with only slightly better performance than Optane and local disk, in both App Direct Mode and Memory Mode.
+There is no apparent difference between the two modes on the given devices.
+\tristan{You should explain why Optane is not reported in memory mode.}
+
+When observing the total task breakdown (Figure~\ref{fig:stacked-25cpus}), it is
+apparent that the differences between the makespans arise from difference in incrementation
+time between all devices except Isilon. In addition, while tmpfs and local disk have comparable write times, Optane
+has a more significant write time difference. Whereas, Isilon has more significant read and
+write time compared to all other devices, it's incrementation time is comparable to that
+of local disk. Further inspection of each storage device's Gantt chart (see Figure~\ref{fig:gantt25})
+show that increased duration of tasks stagger the execution of each individual tasks. This, in turn,
+results in better performance towards the end of the execution, where tasks are significantly more staggered.
+Only tmpfs appears to be able to execute each task efficiently with little staggering.
+\subsubsection{96 Processes}
+\begin{figure}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-96cpus-0delay-1it.pdf}
+    \captionsetup{width=\columnwidth}
+    \caption{Makespan of \bigbrain Incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations.
+5 repetitions were performed.}\label{fig:makespan-96cpus}
+\end{figure}
+\begin{figure}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-96cpus-0delay-1it.pdf}
+    \captionsetup{width=\columnwidth}
+    \caption{total read/increment/write breakdown of \bigbrain incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations.
+5 repetitions were performed.}\label{fig:stacked-96cpus}
+\end{figure}
+
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1568150082-local_1it_96cpus_0delay_npc-3.pdf}
+    \caption{Memory Mode}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567714252-localAD_1it_96cpus_0delay_npc-3.pdf}
+    \caption{App Direct Mode}
+\end{subfigure}
+\caption{Gantt charts for local disk processing 125 blocks of the 40$\mu$m BigBrain using 96 processes with no delay}\label{fig:gantt96}
+\end{figure*}
+
+At 96 processes, it was expected that the application would not satisfy the inequality.
+In Figure~\ref{fig:makespan-96cpus} we see that there is now a greater disparity between the
+makespans of experiments executed in Memory and App Direct Mode, given the same storage. Local disk
+in App Direct Mode is significantly faster than its Memory Mode counterpart. Local disk in App Direct
+mode is even faster than Optane, by a factor of 1.7. Tmpfs was also found to perform better in App Direct Mode, although
+the performance difference is not significant. Only Isilon's performance did not vary between the
+two execution modes.
+
+As with 25 process, the main cause of performance variability appears to be the incrementation
+time (Figure~\ref{fig:stacked-96cpus}). Local disk running in App Direct Mode has an average
+incrementation time that is nearly 15~x faster than in Memory Mode. The incrementation time
+for local disk is, on average, 6~x faster than on Optane.
+
+Isilon in App Direct generally performed the same as in memory mode, however, the write
+was almost 4000~seconds slower, on average, in App Direct Mode.
+
+A further look at the Gantt charts between Memory Mode and App Direct Mode (Figure~\ref{fig:gantt96})
+shows that the sole difference between the two modes is the duration of the incrementation step.
+While both modes have an increased incrementation time, the incrementation duration in
+Memory Mode is much more important than that of App Direct Mode.
+
+\subsection{BIDS App Example}
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/25cores.pdf}
+    \caption{25 processes}\label{fig:bm25}
+\end{subfigure}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/96cores.pdf}
+    \caption{96 processes}\label{fig:bm96}
+\end{subfigure}
+\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}
+\end{figure*}
+
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/25cores-sum.pdf}
+    \caption{25 processes}\label{fig:bb25}
+\end{subfigure}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/96cores-sum.pdf}
+    \caption{96 processes}\label{fig:bb96}
+\end{subfigure}
+\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}\label{fig:bbd}
+\end{figure*}
+Unlike the BigBrain incrementation at 25 Processes, the makespan of the BIDS App Example
+executing on Optane is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
+to the BigBrain Incrementation, tmpfs has a slight lead over local disk and Isilon performs the
+worst. The same pattern can be found with 96 process (Figure~\ref{fig:bm96}). When comparing
+the executions of 25 and 96 processes, the performance improved by a similar factor on all
+storage devices.
+
+The CPU and I/O breakdowns (Figure~\ref{fig:bbd}) show that there is a significant increase
+in the time spent on I/O at 96 processes when compared to 25 processes. Furthermore, there appears
+to also be an increase in the amount of time spent on CPU processing. In both cases,
+tmpfs spends the least amount of time on I/O and CPU, followed by local disk, Optane, and finally,
+Isilon.
+\section{Discussion}
+\subsection{Memory vs App Direct Mode}
+
+While Memory Mode provides a cost-effective solution to increasing the total memory
+size, it is slower than App Direct Mode for embarrassingly parallel applications such
+as \bigbrain incrementation. This is likely due to the Page Cache being composed of a
+combination of both DRAM and Optane in Memory Mode. Since the data produced by the application should be
+able to fit within DRAM, writing to a slower bandwidth device such as Optane inevitably
+deteriorates performance. As App Direct separates usage of DRAM and Optane, devices only utilize DRAM in this mode.
+The only device that performed better in Memory Mode was 
+Isilon, however the difference is minimal.
+%This may be explained by a larger Page Cache in Memory Mode, enabling Isilon to
+%store more data in memory prior to initiating synchronous writes to Isilon.
+
+I/O-less task duration also seem to be affected by the execution mode, with those tasks
+also performing significantly better in App Direct. A potential explanation for this may be 
+that all memory accesses to Optane are slower due to the device's bandwidth, meaning that
+tasks that do not perform I/O, but access data loaded into memory are slower in Memory Mode. Another explanation might be that the kernel is throttling the processes
+as they are producing significant amounts of output, thereby increasing the
+incrementation time.
+This can be seen in Figure~\ref{fig:gantt96} where for local disk, incrementation in App Direct
+mode is significantly shorter. Only when the memory bus is overwhelmed does incrementation take
+longer for App Direct, however, it is nothing compared to the average duration of the incrementation
+tasks in Memory Mode.
+
+Once again, Isilon is the only device which does not experience an increase in duration for
+I/O-less tasks. In fact, the incrementation duration is less than that of local disk in Memory Mode
+and Optane in App Direct Mode. It may be that since even the reads are staggered
+(Figure~\ref{fig:gantt25isilon}), the memory bus is less strained by all the parallel processes,
+therefore incurring less of an overall delay for all memory accesses.
+
+\subsection{Effect of Page Cache}
+
+As long as all the written data does not exceed the dirty ratio, there is little 
+benefit brought by the use of Optane. This is the case with both
+our applications writing less than $\SI{1}{\gibi\byte}$ to storage on a system with 
+$\SI{768}{\gibi\byte}$ of DRAM and a dirty ratio of 20. However, it is expected 
+that as the memory footprints of applications increase to the point where the kernel
+would be forced to perform synchronous writes to disk if it were to rely on DRAM alone,
+the benefit of using Optane in Memory Mode will become more apparent.
+
+However, depending on the neuroimaging application, the amount of data that could be
+loaded in parallel would be limited by the amount of parallelism possible on the system.
+For instance, typical fMRI tasks process many small images (100s of MB in size) in parallel.
+While the datasets themselves can reach PB in size, the amount of data that can be processed in
+parallel can be $\SI{10}{\giga\byte}$ if each take processes $\SI{100}{\mega\byte}$ and the 
+maximum amount of parallelism is 100.
+
+On the other hand, Memory Mode can be very useful for applications operating on
+ultra-high resolution images, like the \bigbrain. For instance, visualization applications
+many need to load significant amounts of data in memory. Given the size of these datasets, 
+extending the amount of available memory using Optane could be quite beneficial and allow
+for faster visualization of larger datasets.
+
+\subsection{Added value of persistent memory}
+
+The added value of Optane a persistent storage device depend on the amount of parallelism
+of the application. For instance, Optane performs better than local disk at 25 Processes,
+but worse than local disk at 96 processes. This can be explained by the fact that it performs
+I/O on the memory bus. While this may reduce overall latency to Optane, the bandwidth of 
+writing to Optane is still significantly less than DRAM. Despite Optane having a similar bandwidth
+to local disk, local disk uses DRAM as cache whereas Optane does not. This fact is 
+particularly evident with the amount of data that is written in parallel at 96 processes. With 25 processes
+and at least a 2 min
+delay between each write, the load on the memory bus is low enough to be efficiently handled.
+
+fMRI application will generally attempt to leverage as much parallelism as possible. As this
+is the case, it is not believed that App Direct Mode would bring a significant performance advantage
+unless the server's CPUs are limited, in which case there is less concurrent accesses on the device.
+The same would apply to ultra-high resolution application unless they are sequential, as may be the case with
+visualization.
 \section{Conclusion}
+
+The benefits of Optane for neuroimaging applications are limited to large memory
+accesses with limited parallelism, as would be the case for visualization. This is
+the case for both Memory Mode and App Direct Mode.
+
+While Memory Mode leads to deterioration in performance of non-Optane storage devices,
+they act as expected in App Direct Mode. It is currently unknown by how much performance
+of a non-Optane storage device would be affected by a large memory access.
+
+To be able to fully determine the benefits of Optane for neuroimaging application,
+it is necessary to process larger datasets. We plan to execute incrementation pipeline on the 
+20~$\mu$m BigBrain using Apache Spark. 
 \section{Acknowledgement}

### 2019-10-21 experiments/bigbrain-incrementation/experiment_launch.py
some reorganization of results/new results

@@ -27,9 +27,9 @@ ef_dir = op.join(exp_dir, 'experiment_files')
 results_dir = op.join(exp_dir, 'results')
 cmd_template = 'time $(parallel --jobs {} < {})'
 if 'em' in sys.argv[3]:
-    cmd_template_spark = 'time $(spark-submit --conf spark.network.timeout=10000000 --master local[{}] --driver-memory 700G spark_inc.py {} {} 1 --benchmark --emulate)'
+    cmd_template_spark = 'time $(spark-submit --conf spark.network.timeout=10000000 --master local[{}] --driver-memory 700G spark_inc.py {} {} 1 --delay {} --benchmark --emulate)'
 else:
-    cmd_template_spark = 'time $(spark-submit --conf spark.network.timeout=10000000 --master local[{}] --driver-memory 700G spark_inc.py {} {} 1 --benchmark)'
+    cmd_template_spark = 'time $(spark-submit --conf spark.network.timeout=10000000 --master local[{}] --driver-memory 700G spark_inc.py {} {} 1 --delay {} --benchmark)'
 
 im_size_b = 646461552 
 
@@ -108,7 +108,6 @@ for i in range(int(sys.argv[2])):
             out_dir = op.join(isilon, out_name)
 
         if not spark:
-            c['delay'] = 0
 
             ef_cmd = [op.join(exp_dir, 'generate_ef.sh'), in_dir, out_dir, str(c['delay'])]
             print(ef_cmd)
@@ -122,7 +121,7 @@ for i in range(int(sys.argv[2])):
 
         else:
             drop_caches()
-            cmd = cmd_template_spark.format(c['ncpus'], in_dir, out_dir)
+            cmd = cmd_template_spark.format(c['ncpus'], in_dir, out_dir, c['delay'])
 
         print(cmd)
 

### 2019-10-21 experiments/bigbrain-incrementation/generate_ef.sh

@@ -8,10 +8,23 @@ app="python /home/users/vhayots/paper-memory-storage/experiments/bigbrain-increm
 efout="experiment_files/$bc-ef.sh"
 
 rm $efout
+num_files=$(ls $fs/* | wc -l)
+its=$((1000 / 70))
+rem=$((1000 % 70))
 
-for f in $fs/*
+for ((i=1;i<=its;i++))
+do
+    for f in $fs/*
+    do
+        bn="$(basename -- $f)"
+        echo $bn $cond
+        echo "$app $f $cond --benchmark_file $cond/$bc-$bn.out --delay $delay" >> $efout
+    done
+done
+
+for f in $(ls $fs/* | head -n rem)
 do
     bn="$(basename -- $f)"
     echo $bn $cond
     echo "$app $f $cond --benchmark_file $cond/$bc-$bn.out --delay $delay" >> $efout
-done 
+done

### 2019-10-21 experiments/bigbrain-incrementation/figures/makespan.py
fixes to makespan script

@@ -8,60 +8,259 @@ from os import path as op
 import numpy as np
 
 
-def makespan(df, task="task_duration"):
+def configure_df(df):
+    df_end = df.End.max()
+    df_start = df.Start.min()
 
-    labels = ['tmpfs', 'Optane', 'local disk', 'Isilon']
+    df_data = pd.concat([df_start, df_end], axis=1)
+    df_data["makespan"] = df_data.End - df_data.Start
+    df_data["disk"] = df_data.index.map(
+        lambda x: x.split("_")[0].split("-")[-1] + ("-em" if "_em" in x else "-real")
+    )
+
+    s_mean = (
+        df_data.rename(columns={"makespan": "mean"})
+        .set_index("disk")
+        .groupby(level=0)
+        .mean()
+    )
+    s_std = (
+        df_data.rename(columns={"makespan": "std"})
+        .set_index("disk")
+        .groupby(level=0)
+        .std()
+    )
+
+    df_res = pd.concat([s_mean["mean"], s_std["std"]], axis=1)
+
+    return df_res
+
+
+
+def makespan(df, task="task_duration", spark=False):
+
+    labels = ["tmpfs", "Optane", "local disk", "Isilon"]
+    if not spark:
+        dfp = df[df["Task"] == "task_duration"].groupby(["filename"])
+    else:
+        dfp = df[(~df["filename"].str.contains('spark')) & (df["Task"] == "task_duration") & ~(df["filename"].str.contains('em'))].groupby(["filename"])
+        dfpe = df[(~df["filename"].str.contains('spark')) & (df["Task"] == "task_duration") & (df["filename"].str.contains('em'))].groupby(["filename"])
+        dfs = df[df["filename"].str.contains('spark') & ~(df["filename"].str.contains('em'))].groupby(["filename"])
+        dfse = df[df["filename"].str.contains('spark') & (df["filename"].str.contains('em'))].groupby(["filename"])
+
     ind = np.arange(len(labels))
-    ind_mem = np.delete(ind, 1)
-    dft = df[df["Task"] == "task_duration"].groupby(["filename"])
 
-    df_end = dft.End.max()
-    df_start = dft.Start.min()
-    df_td = pd.concat([df_start, df_end], axis=1)
-    df_td['makespan'] = df_td.End - df_td.Start
-    df_td['disk'] = df_td.index.map(lambda x: x.split('_')[0].split('-')[-1])
-    
[…]

### 2019-10-22 paper.tex
minor fix
updates to text and figure generation fixes

@@
 in-memory, by ensuring that read and written data are not the same. This application
-will be parallelized using both GNU Parallel and Apache Spark 2.4.3 and implemented in Python 3.6 \tristan{In a subsequent 
+was parallelized us in two different ways: using both GNU Parallel and Apache Spark 2.4.3 (PySpark). 
+All task code was implemented in Python 3.6 \tristan{In a subsequent 
 version you could give more details about implementations, including containerization, etc}. While the GNU Parallel application
@@
 
-We plan to execute this pipeline on both the $\SI{76}{\gibi\byte}$ 40~$\mu$m 
+We have executed this pipeline on both the $\SI{76}{\gibi\byte}$ 40~$\mu$m 
 \bigbrain split into 125 $\SI{617}{\mega\byte}$ blocks and the $\SI{603}{\gibi\byte}$
@@
 
-At this moment, we have executed the application using GNU Parallel for parallelization
+For the 40~$\mu$m \bigbrain, we have executed the application using GNU Parallel for parallelization
 using 25 and 96 processes on the 125 40$\mu$m \bigbrain blocks. Data was read and written to either
@@
 At 96 processes with 0 sleep delay, we estimate being above the inequality for local disk and Isilon.
-All experiments were repeated 5x in each environment (App Direct and Memory Mode)
+All experiments were repeated 5x in each environment (App Direct and Memory Mode).
+
+For the 20~$\mu$m \bigbrain, we used the same application, however in this case, it was executed
+using both Apache Spark and GNU Parallel. The configuration was also generally the same, having 
+experiments using both 25 and 96 cpus. As the previous experiments would have already shown how
+page cache availability for the dirty writes can compare with Optane, we did not include
+a task duration delay. Rather, we just measure the performance differences between the various
+storage devices, in addition to comparing how each device scales by modifying the cluster size.
+
+Since storage was limited on tmpfs and local disk and was not large enough to process the entire
+20~$\mu$m \bigbrain, we had to emulate the reading and writing to these devices. We achieved this
+by copying only 70 of the 1000 blocks onto the storage devices, but used 1000 tasks to read, increment and
+write these 70 blocks.
 
@@
 Time in this application was obtained using Linux's \texttt{time(1)} application.
-\texttt{Real} was used to measure CPU time, where \texttt{Sys} was used to measure
+\texttt{Real + Sys} was used to measure CPU time,
+whereas \texttt{User - (Real + Sys)} was used to measure
 I/O time.
@@
 
-\subsection{\bigbrain incrementation}
+\subsection{40~$\mu$m \bigbrain incrementation}
 \subsubsection{25 Processes}
@@
 There is no apparent difference between the two modes on the given devices.
-\tristan{You should explain why Optane is not reported in memory mode.}
 
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-96cpus-0delay-1it.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-96cpus-0delay-1it.pdf}
     \captionsetup{width=\columnwidth}
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-96cpus-0delay-1it.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-96cpus-0delay-1it.pdf}
     \captionsetup{width=\columnwidth}
@@
 
+\subsection{20~$\mu$ \bigbrain incrementation}
+\subsubsection{25 Processes}
+
+Although tmpfs offers the best performance in terms of makespan duration, Optane
+is a close contender. While tmpfs is approximately 5x faster than Optane, Optane can be
+as much as 12x faster than on Isilon (Figure~\ref{fig:20mksp25}). By using the results of 
+our emulation, we can speculate that Optane will also be significantly faster than 
+local disk in a real scenario. The real Optane execution is almost 3x faster than
+the fastest local disk emulation configuration (i.e. Memory mode - Spark). As the
+emulation results have been found to be faster than the real results, potentially as a side-effect
+of increased accessed to cached data, it is expected
+that the true makespan of the application running on local disk would be significantly
+greater.
+
+The read/increment/write breakdown (Figure~\ref{fig:20stackedp25}) shows that tmpfs
+spends the least amount of time executing all the tasks, however, it spends more
+time than Optane incrementing the data. Otherwise, tmpfs is nearly 470x faster with
+reading and nearly 11x faster at writing than Optane. Despite not reaching memory
+speeds, Optane is significantly faster than other persistent storage devices.
+
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_25cpus.pdf}
+    \caption{real}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-em-20bb_25cpus.pdf}
+    \caption{emulated}
+\end{subfigure}
+\caption{Makespan of the incrementation algorithm processing the 20~$\mu$m BigBrain using
+25 processes}\label{fig:20mksp25}
+\end{figure*}
+
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_25cpus.pdf}
+    \caption{real}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-emulated-20bb_25cpus.pdf}
+    \caption{emulated}
+\end{subfigure}
+\caption{Total read/increment/write breakdown of the  incrementation algorithm processing the 20~$\mu$m BigBrain using
+25 processes}\label{fig:20stackedp25}
+\end{figure*}
+
+\subsubsection{96 Processes}
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_96cpus.pdf}
+    \caption{real}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-em-20bb_96cpus.pdf}
+    \caption{emulated}
+\end{subfigure}
+\caption{Makespan of the incrementation algorithm processing the 20~$\mu$m BigBrain using
+96 processes}\label{fig:20mksp96}
+\end{figure*}
+
+\begin{figure*}
+    \begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_96cpus.pdf}
+    \caption{real}
+\end{subfigure}
+\begin{subfigure}{\columnwidth}
+        \centering
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-emulated-20bb_96cpus.pdf}
+    \caption{emulated}
+\end{subfigure}
+\caption{Total read/increment/write breakdown of the  incrementation algorithm processing the 20~$\mu$m BigBrain using 96 processes}\label{fig:20stackedp96}
+\end{figure*}
 \subsection{BIDS App Example}

### 2019-10-22 experiments/bigbrain-incrementation/figures/makespan.py

@@ -17,7 +17,6 @@ def configure_df(df):
     df_data["disk"] = df_data.index.map(
         lambda x: x.split("_")[0].split("-")[-1] + ("-em" if "_em" in x else "-real")
     )
-
     s_mean = (
         df_data.rename(columns={"makespan": "mean"})
         .set_index("disk")
@@ -68,11 +67,15 @@ def makespan(df, task="task_duration", spark=False):
             },
             inplace=True,
         )
-        print(df_pres_mem)
     else:
         df_sres = configure_df(dfs)
         df_peres = configure_df(dfpe)
         df_seres = configure_df(dfse)
+
+        print(df_pres)
+        print(df_sres)
+        print(df_peres)
+        print(df_seres)
 
 
         df_pres_mem = df_pres[
@@ -135,8 +138,8 @@ def makespan(df, task="task_duration", spark=False):
         ind_ad_real = np.delete(ind, [0, 2])
         labels = ["tmpfs", "optane", "local", "isilon"]
 
-        print(ind_mem_em)
-        print(df_pres_mem_em)
+        #print(ind_mem_em)
+        #print(df_pres_mem_em)
         pmem_pem = ax.bar(
             ind_mem_em - 3 * width / 2,
             df_pres_mem_em["mean"],
@@ -147,8 +150,18 @@ def makespan(df, task="task_duration", spark=False):
             label="Memory mode - GNU Parallel",
         )
 
+        '''pad_read_real = ax.bar(
+            ind_ad_real - width / 2,
+            df_res_ad["mean"],
+            width,
+            alpha=0.4,
+            color="orange",
+            yerr=df_res_ad["std"],
+            label="App Direct real",
+        )'''
+
         pmem_sem = ax.bar(
-            ind_mem_em - width / 2,
+            ind_mem_em + width / 2,
             df_sres_mem_em["mean"],
             width,
             alpha=0.4,
@@ -157,15 +170,7 @@ def makespan(df, task="task_duration", spark=False):
             label="Memory mode - Spark",
         )
 
-        '''pad_read_real = ax.bar(
-            ind_ad_real + width / 2,
-            df_res_ad["mean"],
-            width,
-            alpha=0.4,
-            color="orange",
-            yerr=df_res_ad["std"],
-            label="App Direct real",
-        )'''
+
         pad_sem = ax.bar(
             ind + 1.5 * width,
             df_sres_ad_em["mean"],
[…]

### 2019-10-22 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -10,84 +10,586 @@ import numpy as np
 
 
 def add_disk_col(df):
-    df['disk'] = df.index.map(lambda x: x.split('_')[0].split('-')[-1])
+    df["disk"] = df.index.map(
+        lambda x: x.split("_")[0].split("-")[-1] + ("-em" if "_em" in x else "-real")
+    )
 
 
-def stacked_bar(df):
+def get_mean_std(df, r_task, i_task, w_task):
+    dfr = (
+        df[df["Task"] == r_task]
+        .rename(columns={"Duration": "read_time"})
+        .groupby(["filename"])["read_time"]
+        .sum()
+    )
+    dfi = (
+        df[df["Task"] == i_task]
+        .rename(columns={"Duration": "increment_time"})
+        .groupby(["filename"])["increment_time"]
+        .sum()
+    )
+    dfw = (
+        df[df["Task"] == w_task]
+        .rename(columns={"Duration": "write_time"})
+        .groupby(["filename"])["write_time"]
+        .sum()
+    )
 
-    labels = ['tmpfs', 'Optane', 'local disk', 'Isilon']
+    df_sum = pd.concat([dfr, dfi, dfw], axis=1)
+
+    add_disk_col(df_sum)
+
+    df_mean = df_sum.set_index("disk").groupby("disk").mean()
+    df_std = df_sum.set_index("disk").groupby("disk").std()
+    print("\n\nMean:")
+    print(df_mean)
+    print("\n\nSTD:")
+    print(df_std)
+
+    return df_mean, df_std
+
+
+def stacked_bar(df, spark=False):
+
+    labels = ["tmpfs", "Optane", "local disk", "Isilon"]
     ind = np.arange(len(labels))
     ind_mem = np.delete(ind, 1)
 
-    dfr = df[df["Task"] == "read_file"].rename(columns={'Duration': 'read_time'}).groupby(["filename"])["read_time"].sum()
-    dfi = df[df["Task"] == "increment_file"].rename(columns={'Duration': 'increment_time'}).groupby(["filename"])["increment_time"].sum()
-    dfw = df[df["Task"] == "write_file"].rename(columns={'Duration': 'write_time'}).groupby(["filename"])["write_time"].sum()
+    r_task = "read_file"
+    i_task = "increment_file"
+    w_task = "write_file"
 
-    df_sum = pd.concat([dfr, dfi, dfw], axis=1)
-    
-    add_disk_col(df_sum)
[…]

### 2019-10-22 gen_figures.sh

@@ -8,6 +8,10 @@ cd $curr_dir/experiments/bigbrain-incrementation/figures
 ./stacked_bar.py '../results/*25cpus*' 25cpus-120delay-1it
 ./makespan.py '../results/*96cpus*' 96cpus-0delay-1it
 ./stacked_bar.py '../results/*96cpus*' 96cpus-0delay-1it
+./makespan.py '../results/20bb_experiments/*25cpus*' 20bb_25cpus spark
+./stacked_bar.py '../results/20bb_experiments/*25cpus*' 20bb_25cpus spark
+./makespan.py '../results/20bb_experiments/*96cpus*' 20bb_96cpus spark
+./stacked_bar.py '../results/20bb_experiments/*96cpus*' 20bb_96cpus spark
 cd $curr_dir
 ./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567704324-localAD_1it_25cpus_120delay_withpc-2.out --cpus 25
 ./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567705458-optaneAD_1it_25cpus_120delay_withpc-2.out --cpus 25

### 2019-10-22 scripts/gantt.py

@@ -9,19 +9,62 @@ from json import load
 from datetime import datetime
 import time
 import argparse
+import numpy as np
 
 
-def gantt_increment(df, data_file, ncpus):
+def gantt_increment(df, data_file, ncpus, spark=False):
     color = {
         "read_file": "turquoise",
         "increment_file": "crimson",
         "write_file": "purple",
     }
 
+    if spark:
+        color = {
+            "load_img": "turquoise",
+            "increment_data": "crimson",
+            "save_incremented": "purple",
+        }
+
     fig, ax = plt.subplots(figsize=(6, 3))
     labels = []
 
-    df = df[df["Task"] != "task_duration"]
+    df = df[~df["Task"].str.contains("task_duration") & ~df["Task"].str.contains("driver_program")]
+
+
+    program_start = df.Start.min()
+    program_end = df.End.max() - program_start
+    
+    def count_overlaps(df):
+        #df = df[df['Task'].str.contains("increment_file")]
+        df['tmp'] = 1
+        df1 = pd.merge(df,df.reset_index(),on=['tmp'])
+        df = df.drop('tmp', axis=1)
+        df1 = df1[(df1["Start_x"]<=df1["Start_y"])  & (df1["End_x"]> df1["Start_y"])]
+        #df1 = df1[((df1["Start_x"]>=df1["Start_y"]) & (df1["Start_x"] <= df1["End_y"]))]
+        df['overlaps'] = df1.groupby('index').size()
+        return (df)
+
+    '''
+    for idx, rw in df.iterrows():
+        overlaps = 0
+        for index, row in df.iterrows():
+            if not rw.equals(row):
+                start_idx_row = row['Start']
+                end_idx_row = row['Duration'] + start_idx_row
+                start_idx_rw = rw['Start']
+                end_idx_rw = rw['Duration'] + start_idx_rw
+
+                if ((start_idx_rw >= start_idx_row and start_idx_rw <= end_idx_row) or
+                    (end_idx_rw >= start_idx_row and end_idx_rw <= end_idx_row)):
+                    overlaps += 1
+
+        if overlaps > max_overlaps:
+            max_overlaps = overlaps
+    '''
+    print('Avg number of overlaps:', count_overlaps(df)['overlaps'].mean())
+
+    print('Number of threads:', len(df.groupby("ThreadId")))
     for i, task in enumerate(df.groupby("ThreadId")):
         labels.append(task[0])
[…]

### 2019-10-22 paper.tex
Fixed typos

@@
 \usepackage{hyperref}
+\usepackage{xspace}
 \usepackage[binary-units]{siunitx}
@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\newcommand{\bigdata}{Big Data }
-\newcommand{\bigbrain}{BigBrain }
+\newcommand{\bigdata}{Big Data\xspace}
+\newcommand{\bigbrain}{BigBrain\xspace}
 
@@
 has a more significant write time difference. Whereas, Isilon has more significant read and
-write time compared to all other devices, it's incrementation time is comparable to that
+write time compared to all other devices, its incrementation time is comparable to that
 of local disk. Further inspection of each storage device's Gantt chart (see Figure~\ref{fig:gantt25})
-show that increased duration of tasks stagger the execution of each individual tasks. This, in turn,
+shows that increased duration of tasks stagger the execution of each individual tasks. This, in turn,
 results in better performance towards the end of the execution, where tasks are significantly more staggered.
@@
 
-As with 25 process, the main cause of performance variability appears to be the incrementation
+As with 25 processes, the main cause of performance variability appears to be the incrementation
 time (Figure~\ref{fig:stacked-96cpus}). Local disk running in App Direct Mode has an average
@@
 
-The added value of Optane a persistent storage device depend on the amount of parallelism
-of the application. For instance, Optane performs better than local disk at 25 Processes,
+The added value of Optane as a persistent storage device depends on the amount of parallelism
+of the application. For instance, Optane performs better than local disk at 25 processes,
 but worse than local disk at 96 processes. This can be explained by the fact that it performs

### 2019-10-22 paper.tex
normalized colours in figures

--- 2019-10-22 paper.tex
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-25cpus-120delay-1it.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-25cpus-120delay-1it.pdf}
     \captionsetup{width=\columnwidth}
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-25cpus-120delay-1it.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-25cpus-120delay-1it.pdf}
     \captionsetup{width=\columnwidth}
@@
 speeds, Optane is significantly faster than other persistent storage devices.
+For instance, it can be almost up to 10x faster in reads than Isilon and almost 
+9x faster in writes. It is important, however, to note that total read times in App
+Direct mode for the Spark execution on Isilon are almost twice as fast as their 
+equivalent Optane counterpart. On the other hand, this decrease read time is offset
+by the write time which are 9x slower.
 
+Emulated breakdown results also agree that the bulk of the differences in makespans
+can be attributed to write time, with local disk writes being notably slower than
+Optane. In fact, local disk write speeds nearly match that of Isilon. Isilon results
+in a longer makespan due to increase read and incrementation time.
 \begin{figure*}
@@
 \subsubsection{96 Processes}
+
+With nearly a four-fold increase in parallelization, the makespan generally did
+not improve other but a few 100s on Isilon. Overall, the general trend is similar
+than with 25 cpus. As can be seen in Figure~\ref{fig:20mksp96}, tmpfs is the
+fastest storage device. Regardless, Optane remains significantly faster than the
+other storage devices (e.g. Isilon). However, emulated disk results appear to 
+imply that the speedup difference between Optane and local disk will be smaller.
+
 \begin{figure*}

### 2019-10-22 experiments/bids-app-example/plot.py

@@ -44,7 +44,7 @@ def get_makespans(conditions):
 mksp_96 = get_makespans(conditions_96_jobs)
 plt.ylabel("Makespan (s)")
 plt.ylim((0,5000))
-plt.bar([1, 2, 3, 4], mksp_96[0], tick_label=condition_ticks, width=width, yerr=mksp_96[1])
+plt.bar([1, 2, 3, 4], mksp_96[0], tick_label=condition_ticks, width=width, yerr=mksp_96[1], alpha=0.4, color='gray')
 plt.savefig("96cores.pdf")
 
 plt.clf()
@@ -52,7 +52,7 @@ plt.clf()
 mksp_25 = get_makespans(conditions_25_jobs)
 plt.ylabel("Makespan (s)")
 plt.ylim((0,5000))
-plt.bar([1, 2, 3, 4], mksp_25[0], tick_label=condition_ticks, width=width, yerr=mksp_25[1])
+plt.bar([1, 2, 3, 4], mksp_25[0], tick_label=condition_ticks, width=width, yerr=mksp_25[1], alpha=0.4, color='gray')
 plt.savefig("25cores.pdf")
 
 def time_to_secs(t):
@@ -113,8 +113,10 @@ plt.clf()
 res_96 = get_cpu_io_times(conditions_96_jobs)
 plt.ylabel("Total time (s)")
 plt.ylim(0, 400000)
-p1 = plt.bar([1, 2, 3, 4], res_96[2], tick_label=condition_ticks, width=width, yerr=res_96[3])
-p2 = plt.bar([1, 2, 3, 4], res_96[0], bottom=res_96[2], width=width, yerr=res_96[1])
+p1 = plt.bar([1, 2, 3, 4], res_96[2], tick_label=condition_ticks, width=width,
+        yerr=res_96[3], alpha=0.4, color='red', hatch='+++', edgecolor='green')
+p2 = plt.bar([1, 2, 3, 4], res_96[0], bottom=res_96[2], width=width,
+        yerr=res_96[1], alpha=0.4, color='blue')
 
 plt.legend((p1[0], p2[0]), ('I/O', 'CPU'))
 plt.savefig('96cores-sum.pdf')
@@ -125,7 +127,9 @@ plt.ylabel("Total time (s)")
 plt.ylim(0, 400000)
 res_25 = get_cpu_io_times(conditions_25_jobs)
 
[…]

### 2019-10-22 experiments/bigbrain-incrementation/figures/makespan.py

@@ -38,7 +38,7 @@ def configure_df(df):
 
 def makespan(df, task="task_duration", spark=False):
 
-    labels = ["tmpfs", "Optane", "local disk", "Isilon"]
+    labels = ["DRAM", "Optane", "local disk", "Isilon"]
     if not spark:
         dfp = df[df["Task"] == "task_duration"].groupby(["filename"])
     else:
@@ -112,13 +112,13 @@ def makespan(df, task="task_duration", spark=False):
 
     fig, ax = plt.subplots()
     if not spark:
-        ind_mem = np.delete(ind, 1)
+        ind_mem = np.delete(ind, 0)
         pmem_read = ax.bar(
             ind_mem - width / 2,
             df_pres_mem["mean"],
             width,
             alpha=0.4,
-            color="b",
+            color="gray",
             yerr=df_pres_mem["std"],
             label="Memory mode",
         )
@@ -127,16 +127,16 @@ def makespan(df, task="task_duration", spark=False):
             df_pres_ad["mean"],
             width,
             alpha=0.4,
-            color="orange",
+            color="gray",
+            hatch="//",
             yerr=df_pres_ad["std"],
             label="App Direct mode",
         )
     else:
         width = 0.2
-        ind_mem_real = np.delete(ind, [1, 2])
-        ind_mem_em = np.delete(ind, 1)
-        ind_ad_real = np.delete(ind, [0, 2])
-        labels = ["tmpfs", "optane", "local", "isilon"]
+        labels_em =  ["DRAM", "Optane", "local disk", "Isilon"]
+        ind = np.arange(len(labels_em))
+        ind_mem_em = np.delete(ind, 0)
 
         #print(ind_mem_em)
         #print(df_pres_mem_em)
@@ -145,7 +145,7 @@ def makespan(df, task="task_duration", spark=False):
             df_pres_mem_em["mean"],
             width,
             alpha=0.4,
-            color="r",
+            color="gray",
             yerr=df_pres_mem_em["std"],
             label="Memory mode - GNU Parallel",
         )
@@ -155,7 +155,8 @@ def makespan(df, task="task_duration", spark=False):
             df_res_ad["mean"],
             width,
             alpha=0.4,
-            color="orange",
+            color="gray",
+            hatch="//",
             yerr=df_res_ad["std"],
             label="App Direct real",
         )'''
@@ -165,7 +166,8 @@ def makespan(df, task="task_duration", spark=False):
             df_sres_mem_em["mean"],
             width,
             alpha=0.4,
[…]

### 2019-10-22 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -51,9 +51,9 @@ def get_mean_std(df, r_task, i_task, w_task):
 
 def stacked_bar(df, spark=False):
 
-    labels = ["tmpfs", "Optane", "local disk", "Isilon"]
+    labels = ["DRAM", "Optane", "local disk", "Isilon"]
     ind = np.arange(len(labels))
-    ind_mem = np.delete(ind, 1)
+    ind_mem = np.delete(ind, 0)
 
     r_task = "read_file"
     i_task = "increment_file"
@@ -131,8 +131,6 @@ def stacked_bar(df, spark=False):
         df_mean_ad = prep_ad_df(df_mean)
         df_std_ad = prep_ad_df(df_std)
     else:
-        ind_mem_r = np.delete(ind_mem, 1)
-        ind_ad_r = np.delete(ind, [0, 2])
         df_mean_ad_r = prep_ad_df(df_mean, "real")
         df_mean_ad_em = prep_ad_df(df_mean, "em")
         df_std_ad_r = prep_ad_df(df_std, "real")
@@ -379,6 +377,11 @@ def stacked_bar(df, spark=False):
 
         ax.legend(handles=[r_mem_em, r_ad_em, r_smem_em, r_sad_em, r_read, r_inc, r_write])
         plt.savefig("stacked-emulated-{}.pdf".format(sys.argv[2]))
+
+        labels = ["Optane", "Isilon"]
+        ind = np.arange(len(labels))
+        ind_mem_r = ind
+        ind_ad_r = ind
 
         figure, ax = plt.subplots()
         #### REAL

### 2019-10-22 scripts/gantt.py

@@ -14,16 +14,16 @@ import numpy as np
 
 def gantt_increment(df, data_file, ncpus, spark=False):
     color = {
-        "read_file": "turquoise",
-        "increment_file": "crimson",
-        "write_file": "purple",
+        "read_file": "red",
+        "increment_file": "blue",
+        "write_file": "green",
     }
 
     if spark:
         color = {
-            "load_img": "turquoise",
-            "increment_data": "crimson",
-            "save_incremented": "purple",
+            "load_img": "red",
+            "increment_data": "blue",
+            "save_incremented": "green",
         }
 
     fig, ax = plt.subplots(figsize=(6, 3))
@@ -41,27 +41,9 @@ def gantt_increment(df, data_file, ncpus, spark=False):
         df1 = pd.merge(df,df.reset_index(),on=['tmp'])
         df = df.drop('tmp', axis=1)
         df1 = df1[(df1["Start_x"]<=df1["Start_y"])  & (df1["End_x"]> df1["Start_y"])]
-        #df1 = df1[((df1["Start_x"]>=df1["Start_y"]) & (df1["Start_x"] <= df1["End_y"]))]
         df['overlaps'] = df1.groupby('index').size()
         return (df)
 
-    '''
-    for idx, rw in df.iterrows():
-        overlaps = 0
-        for index, row in df.iterrows():
-            if not rw.equals(row):
-                start_idx_row = row['Start']
-                end_idx_row = row['Duration'] + start_idx_row
-                start_idx_rw = rw['Start']
-                end_idx_rw = rw['Duration'] + start_idx_rw
-
-                if ((start_idx_rw >= start_idx_row and start_idx_rw <= end_idx_row) or
-                    (end_idx_rw >= start_idx_row and end_idx_rw <= end_idx_row)):
-                    overlaps += 1
-
-        if overlaps > max_overlaps:
-            max_overlaps = overlaps
-    '''
     print('Avg number of overlaps:', count_overlaps(df)['overlaps'].mean())
 
     print('Number of threads:', len(df.groupby("ThreadId")))
@@ -69,11 +51,11 @@ def gantt_increment(df, data_file, ncpus, spark=False):
         labels.append(task[0])
         for r in task[1].groupby("Task"):
             data = r[1][["Start", "Duration"]]
-            ax.broken_barh(data.values, (i - 0.4, 0.8), color=color[r[0]])
[…]

### 2019-10-23 paper.tex
fixed discussion

@@
 not improve other but a few 100s on Isilon. Overall, the general trend is similar
-than with 25 cpus. As can be seen in Figure~\ref{fig:20mksp96}, tmpfs is the
-fastest storage device. Regardless, Optane remains significantly faster than the
-other storage devices (e.g. Isilon). However, emulated disk results appear to 
-imply that the speedup difference between Optane and local disk will be smaller.
+than with 25 cpus. As can be seen in Figure~\ref{fig:20mksp96}, Optane in Memory
+mode is the fastest storage device. Furthermore, Optane in App Direct mode 
+remains significantly faster than the other storage devices (e.g. Isilon).
+However, emulated disk results appear to imply that the speedup difference
+between Optane in App Direct mode and local disk will be smaller than at 25 cpus.
+
+The vast majority of processing time appears to be spent on writing to the 
+respective storage device with the exception of Optane in Memory Mode, which 
+spends most of its time incrementing (Figure~\ref{fig:20stackedp96}). App Direct
+mode executions appear to have reduced incrementation time coupled with increased
+write times when compared to their memory mode equivalents. Furthermore, read
+times appear to be increased for Isilon memory mode executions. These events are
+not reflected in the emulated executions.
+
+Emulated breakdowns demonstrate that local disk executions will remain slower 
+than Optane in both App Direct and Memory modes. Results show that both 
+incrementation and write times will be generally be longer on local disk than
+on Optane.
 
@@
 \subsection{Memory vs App Direct Mode}
+
+Memory mode was found to result in increased computation time. When processing 
+the 20~$\mu$m \bigbrain, Memory mode results indicate that more time was spent on
+the incrementation tasks than its App Direct counterparts. Unlike App Direct mode,
+which relies exclusively on DRAM for working memory, Memory mode relies on Optane
+with a DRAM cache. Optane, while faster than other persistent storage devices, 
+still lacks the bandwidth of DRAM. Should anonymous memory need to access Optane
+for computation, it is not surprising that it would be slower than a DRAM-only
+access. However, whereas Memory mode experienced increased computation during the
+processing of the 20~$\mu$m \bigbrain, this was not always found to be the case
+with the processing of the 40~$\mu$m \bigbrain. In fact, incrementation time was
+largely similar between the two modes, with the notable exception of Optane in
+Memory mode expending significantly less computation time than in App Direct when
+processing the 40~$\mu$m \bigbrain with both 25 and 96 processes. Optane in 
+Memory mode executing on local disk using 96 cpus, however, required significant
+computation time as compared to App Direct. 
+- Conclusion: memory mode should only be used when there's a requirement for additional memory
+-App Direct mode results in increased I/O time for Optane
+- conclusion: Optane might benefit from the DRAM cache
 

### 2019-10-23 experiments/bigbrain-incrementation/figures/makespan.py

@@ -9,10 +9,13 @@ import numpy as np
 
 
 def configure_df(df):
-    df_end = df.End.max()
-    df_start = df.Start.min()
+    df_end = df.End.apply(lambda x: x.max())
+    df_start = df.Start.apply(lambda x: x.min())
+
+    #print(df.End.apply(lambda x: x.max()))
 
     df_data = pd.concat([df_start, df_end], axis=1)
+    print(df_data.head(5))
     df_data["makespan"] = df_data.End - df_data.Start
     df_data["disk"] = df_data.index.map(
         lambda x: x.split("_")[0].split("-")[-1] + ("-em" if "_em" in x else "-real")
@@ -67,6 +70,7 @@ def makespan(df, task="task_duration", spark=False):
             },
             inplace=True,
         )
+        print(df_pres)
     else:
         df_sres = configure_df(dfs)
         df_peres = configure_df(dfpe)

### 2019-11-21 paper.tex
cleanup/changes in paper re results/modifications to figure generating scripts

@@
 \usepackage{xspace}
+\usepackage{multirow}
 \usepackage[binary-units]{siunitx}
@@
 
-The server used in our experiments contain $\SI{3}{\tera\byte}$ of Intel Optane Persistent memory, 768GB of DRAM,
-Isilon, local disk(SSD).
-CPU: Cascade lake
-OS: Red Hat 7.1
-writeback enabled for local disk (unsure about Isilon)
-\valerie{include storage bandwidth table here}
+The server used for the experiments consisted of 12 \SI{64}{\gibi\byte} DRAM devices,
+resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12 \SI{256}{\giga\byte} Intel Optane DC Memory
+devices, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC Memory. The theoretical maximum bandwidth
+of DRAM is estimated to be around \SI{275}{\gibi\byte}~/s whereas for Optane, it is
+estimated to be around \SI{250}{\gibi\byte}~/s.
+Other storage devices included a \SI{240}{\gibi\byte} Micron SATA SSD, of which
+only \SI{149}{\gibi\byte} where available for the experiments, and a \SI{628}{\tebi\byte} shared Dell EMC Isilon cluster. The local disk
+was the only storage device set up as a writeback device. Both Isilon and Optane in
+App Direct mode were configured as write-through devices, meaning they did not leverage
+DRAM as a cache for writes. This was achieved in Optane by configuring the xfs filesystem with
+DAX. All storage devices were benchmarked using the following \href{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}{script}.
+The result of the benchmarks can be seen in Table \ref{table:bandwidths}.
+
+For processing,
+2 Intel(R) Xeon(R) Platinum 8260M CPU @ 2.40GHz where installed, enabling use of up to 96 threads.
+The server was running Red Hat Enterprise Linux 7.6 release (Maipo).
+
+\begin{table}
+\begin{center}
+ \begin{tabular}{ |c|c|c| } 
+     \cline{2-3}
+     \multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Bandwidth (MB/s)} \\\hline
+  Device & Read & Write \\
+ \hline
+ DRAM & 5304.3 & 3338.2 \\  
+ Optane & 3379.2 & 2396.2 \\   
+ Local Disk & 518.6 & 240.4 \\
+ Isilon & 117.0 & 111.8 \\
+ \hline
+\end{tabular}\caption{Perceived read and write bandwidth of each storage device}\label{table:bandwidths}
+\end{center}
+\end{table}
 
@@
 
-Memory Mode allows Intel Optane DC persistent memory to extend the system's available
-memory. In this mode, Optane uses DRAM as a cache and Optane can be accessed as
+Memory Mode leverages Intel Optane DC persistent memory (Optane) to extend the system's available
+memory. In this mode, Optane uses DRAM as a cache and is accessible as
 volatile addressable main memory. By extending main memory, Memory Mode enables the
@@
 
-As Memory Mode would affect the processing time regardless of the device
-selected, we evaluated Memory Mode on tmpfs, local disk and Isilon. tmpfs
-consisted of $\SI{298}{\giga\byte}$ and was located in \texttt{/run/user/61218}.
-Local disk comprised of a $\SI{240}{\giga\byte}$ Micron 5100 PRO SSD, of which
-$\SI{150}{\giga\byte}$ were mounted on \texttt{/home}. The $\SI{628}{\tera\byte}$
-Isilon was mounted as NFS4 to \texttt{/home/users}
+While enabling the access to larger amount of memory than typically accessible otherwise,
+it is anticipated that Memory Mode will be slower than AD mode for all memory accessed that
+are not cached in DRAM. We evaluated all available devices (Optane through tmpfs, local disk and Isilon)
+in Memory Mode.
 
-Optane was not evaluated separately as it is indistinct from DRAM memory in
-Memory Mode
 \subsubsection{App Direct Mode}
@@
 App Direct Mode enables Optane to be accessed as a high-performance storage device.
-Unlike Memory Mode, the OS is able to differentiate between main memory and Optane
-storage, treating them as two distinct entities. As Optane is located on
+Unlike Memory Mode, the OS is able to differentiate between DRAM and Optane,
+treating them as two distinct entities. As Optane is located on
 the DIMMs, it exhibits lower latencies than traditional SSDs, while being able
-to be accessed like an SSD. Optane does not use the DRAM as cache in App Direct Mode.
+to be accessed like an SSD. Optane does not use the DRAM as cache in App Direct Mode
+if configured with DAX.
 
-The same storage devices were used to evaluate App Direct Mode, with the inclusion
-of Intel Optane DC Persistent Memory, of which $\SI{1.5}{\tera\byte}$ of storage
-were mounted on \texttt{/pvme-1}.
+Similarly to Memory Mode, our experiments evaluated all available filesystems in App Direct
+Mode. In other words, we evaluated DRAM performance through tmpfs, Optane performance through
+a \SI{1.5}{\tebi\byte} Intel Optane DC Persistent Memory storage device and local disk
+and Isilon using the same storage mounts available in Memory Mode.
 
@@
 
-For the purposes of these experiments, we use the inequality introduced in
-\cite{paperinmem} to characterize data-intensive applications. That is,
+We characterize the performance of our experiments using the following model:
+
 \begin{equation}
-\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
+    M \geq \frac{D}{R} + \frac{D}{W} \label{eq:makespan}
 \end{equation}
-where:
+
+where,
 \begin{itemize}
-    \item $D$ is the total amount of data written by the application
-    \item $C$ is the total execution time of the application
-    \item $\delta$ is the disk bandwidth
-    \item $\gamma$ is the max number of concurrent processes on a node
+        \item $M$ is the application makespan
+        \item $D$ is the total amount of data processed by the application
+        \item $R$ is the device read bandwidth
+        \item $W$ is the device write bandwidth
 \end{itemize}
 
-Using this inequality, we anticipate that whenever the application is within the limits,
-the storage selection will make little difference, as storage devices with writeback enabled 
-will be able to leverage the memory cache. It is expected that Optane in App Direct Mode
-will be most useful when this inequality is not satisfied as storage devices will
-not be able to use the writeback cache. As Memory Mode will inevitably
-increase the amount of data that can be cached by the storage, it is expected that the overall
-performance of all storage devices improve, even when the inequality is not satisfied.
+For applications which have negligible CPU time, as is the case with our experiments, 
+it is expected that the I/O duration can estimate the total makespan. However, there
+are certain instances where the makespan may be below the I/O duration. This is expected
+to occur with storage devices that scale well with increased parallelism, as should be the
+case for DRAM, Optane and Isilon, which all consist of multiple devices. Should the
+scalable storage devices predict the makespan accurately, it is believed that other overheads
+would be at play making the device unfavourable for parallel processing.
 
@@
         \State read $chunk$ from $fs$
-        \For{$i \in [1, n]$}
         \State $chunk\gets chunk+1$
-            \State sleep $x$
-        \EndFor
         \State save $chunk$ to $fs$
@@
 Due to the size and uniqueness of the \bigbrain, standardized processing pipelines
-have yet to be developed for it. In order to quantify the effects of data location
+have yet to be developed. In order to quantify the effects of storage devices
 on such a dataset, we have implemented a naive synthetic application that takes 
@@
 was parallelized us in two different ways: using both GNU Parallel and Apache Spark 2.4.3 (PySpark). 
-All task code was implemented in Python 3.6 \tristan{In a subsequent 
-version you could give more details about implementations, including containerization, etc}. While the GNU Parallel application
-is currently limited to 1 incrementation step, the Apache Spark implementation allows us to increase
-the number of incrementation steps by a variable amount, thereby increasing the amount of intermediate
-data generated.
+All code was implemented in Python 3.6.
 
 Time is measured within the application using Python's \texttt{time} module 
-called before and after any reads, increments or writes.
+called before and after any read and writes.
 
-We have executed this pipeline on both the $\SI{76}{\gibi\byte}$ 40~$\mu$m 
-\bigbrain split into 125 $\SI{617}{\mega\byte}$ blocks and the $\SI{603}{\gibi\byte}$
+We have executed this pipeline on both the $\SI{75}{\gibi\byte}$ 40~$\mu$m 
+\bigbrain split into 125 $\SI{614}{\mebi\byte}$ blocks and the $\SI{603}{\gibi\byte}$
 \bigbrain at 20~$\mu$m split into 1000 $\SI{617}{\mebi\byte}$ blocks.
@@
 using 25 and 96 processes on the 125 40$\mu$m \bigbrain blocks. Data was read and written to either
-tmpfs, Optane (App Direct only), local disk and Isilon. In all configurations (25 and 96 processes)
-only 1 iteration was performed. At 25 processes, a sleep delay of 120 seconds was added at the
-incrementation step, which would emulate tasks of longer duration and ensure that we satisfy Equation~\ref{eq:page-cache-inequality} for all devices.
-At 96 processes with 0 sleep delay, we estimate being above the inequality for local disk and Isilon.
-All experiments were repeated 5x in each environment (App Direct and Memory Mode).
+DRAM (App Direct only), Optane, local disk and Isilon.
 
@@
 using both Apache Spark and GNU Parallel. The configuration was also generally the same, having 
-experiments using both 25 and 96 cpus. As the previous experiments would have already shown how
-page cache availability for the dirty writes can compare with Optane, we did not include
-a task duration delay. Rather, we just measure the performance differences between the various
-storage devices, in addition to comparing how each device scales by modifying the cluster size.
+experiments using both 25 and 96 cpus. Repeating the experiments using a larger dataset would enable
+us to determine the effects when Optane would have to be partially relied upon due to insufficient DRAM
+space. Moreover, with such large datasets, it is likely that Big Data frameworks would be used rather than
+more traditional parallelism frameworks. Using Spark, we can evaluate how Big Data Frameworks perform on
+different storage devices.
 
-Since storage was limited on tmpfs and local disk and was not large enough to process the entire
-20~$\mu$m \bigbrain, we had to emulate the reading and writing to these devices. We achieved this
-by copying only 70 of the 1000 blocks onto the storage devices, but used 1000 tasks to read, increment and
-write these 70 blocks.
+Since storage was limited on DRAM and local disk and was not large enough to process the entire
+20~$\mu$m \bigbrain, these devices were omitted in the processed of this dataset.
 
@@
 compliant application. It runs a standard neuroimaging brain extraction application (FSL BET) 
-on all the anatomical (T1W) images of a dataset containing numerous subjects. This step is 
-referred to as Participant analysis, within the BIDS App Example. An optional
+on all the anatomical (T1W) images of dataset containing numerous subjects. This step is 
+referred to as Participant analysis within the BIDS App Example. An optional
 step of the BIDS App example, referred to as Group analysis, computes the average brain
@@
 
-For our experiments, we used the entire CoRR dataset, made available on DataLad and
-applied Participant analysis on it. As in the BigBrain Incrementation, the experiment parallelized 
+For our experiments, we used the entire CoRR dataset, made available on \href{}{DataLad} and
+applied Participant analysis to it. The BIDS App Example was executed using a Singularity container stored on
+Isilon. As in the BigBrain Incrementation, the experiment parallelized 
 using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode
-using Isilon, local disk, Optane and tmpfs as storage. As participant analysis operates on only a subset of the dataset,
-the inequality is satified for all conditions. Each experiment was repeated 3x.
+using Isilon, local disk and Optane as storage. Each experiment was repeated 3x.
 
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-25cpus-120delay-1it.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpu.pdf}
     \captionsetup{width=\columnwidth}
-    \caption{Makespan of \bigbrain Incrementation using 25 processes with a sleep delay of 120s on 125 blocks on the 40~$\mu$m image given different storage configurations.
-5 repetitions were performed.}\label{fig:makespan-25cpus}
+    \caption{Makespan of the \bigbrain Incrementation application using 25 
+             processes on the 125 blocks on the 40~$\mu$m image given different
+             storage configurations. Anticipated read and write duration was 
+             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
+             to Eq.~\ref{eq:makespan}. 3 
+             repetitions were performed. *Local SSD did not complete the
+    writing of the last few blocks due to storage limitations.}\label{fig:makespan-25cpus}
 \end{figure}
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-25cpus-120delay-1it.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/violin-read-20bb_25cpu.pdf}
     \captionsetup{width=\columnwidth}
-    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 processes with a sleep delay of 120s on 125 blocks on the 40~$\mu$m image given different storage configurations.
-5 repetitions were performed.}\label{fig:stacked-25cpus}
+    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 2. 3 repetitions were performed.}\label{fig:stacked-25cpus}
 \end{figure}
@@
 
-As shown in Figure~\ref{fig:makespan-25cpus}, despite being under the inequality, the
-performance still varies according to selected storage device. As expected, Optane performs
-slightly better than local disk and significantly better than Isilon. Tmpfs outperforms all
-devices, with only slightly better performance than Optane and local disk, in both App Direct Mode and Memory Mode.
-There is no apparent difference between the two modes on the given devices.
+As can be seen in Figure~\ref{fig:makespan-25cpus}, DRAM (App Direct Mode) is the 
+most efficient, with a makespan of 7.6s. Optane in Memory mode is a close second
+with a makespan of approximately 10s. Optane in App Direct mode is approximately 
+12x slower than Optane in Memory Mode. Isilon in both Memory Mode and App Direct 
+Mode is more than 120x slower than Optane in Memory Mode. Other than for Optane, 
+there is no significant difference between Memory Mode and App Direct executions.
+
+When comparing the makespan to the estimated read and write duration, local SSD
+in both modes and Optane in App Direct Mode have longer makespans than the anticipated.
+Optane was approximately 2.2x slower than expected, whereas local SSD was approximately
+1.4x slower than expected, in both modes. For those where the I/O time was overestimated,
+DRAM was almost 5x faster in reality compared to what was estimated. Optane in Memory Mode
+had a smaller difference, with it being 3.8x faster than expected. Isilon in App Direct
+mode had a larger difference than Memory mode, with it being 1.1x faster than expected.
 
@@
 Memory mode executing on local disk using 96 cpus, however, required significant
-computation time as compared to App Direct. 
+computation time as compared to App Direct mode. For all GNU Parallel runs, the 
+read only consisted of loading metadata into
 - Conclusion: memory mode should only be used when there's a requirement for additional memory

### 2019-11-21 experiments/bigbrain-incrementation/figures/makespan.py

@@ -3,9 +3,20 @@
 import sys
 import pandas as pd
 import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
 from glob import glob
 from os import path as op
 import numpy as np
+
+
+ad_bench = '../../system_conf/ad_diskbench.out'
+mem_bench = '../../system_conf/mem_diskbench.out'
+data_size = 614 * 125
+total_1000 = 617 * 1000
+bb20 = '20bb' in sys.argv[1]
+
+if bb20:
+    data_size = total_1000
 
 
 def configure_df(df):
@@ -41,7 +52,7 @@ def configure_df(df):
 
 def makespan(df, task="task_duration", spark=False):
 
-    labels = ["DRAM", "Optane", "local disk", "Isilon"]
+    labels = ["DRAM", "Optane", "local SSD*", "Isilon"]
     if not spark:
         dfp = df[df["Task"] == "task_duration"].groupby(["filename"])
     else:
@@ -102,6 +113,12 @@ def makespan(df, task="task_duration", spark=False):
         ]
         df_sres_mem_em = df_sres_mem_em.sort_index(ascending=False)
 
+        df_pres_ad = df_pres[
+            df_pres.index.map(lambda x: 'AD' not in x)
+        ]
+        df_pres_ad = df_pres_ad.sort_index(ascending=False)
+        df_pres_ad.rename(index={'tmpfsAD':'tmpfs', 'optaneAD': 'optane', 'localAD': 'local', 'isilonAD':'isilon'}, inplace=True)
+
         df_sres_ad = df_sres[df_sres.index.map(lambda x: 'AD' in x)]
         df_sres_ad = df_sres_ad.sort_index(ascending=False)
         df_sres_ad.rename(index={'tmpfsAD':'tmpfs', 'optaneAD': 'optane', 'localAD': 'local', 'isilonAD':'isilon'}, inplace=True)
@@ -110,15 +127,58 @@ def makespan(df, task="task_duration", spark=False):
         df_sres_ad_em = df_seres[df_seres.index.map(lambda x: 'AD' in x and 'em' in x)]
         df_sres_ad_em = df_sres_ad_em.sort_index(ascending=False)
         df_sres_ad_em.rename(index={'tmpfsAD':'tmpfs', 'optaneAD': 'optane', 'localAD': 'local', 'isilonAD':'isilon'}, inplace=True)
-        
 
-    width = 0.35
+
[…]

### 2019-11-21 experiments/bigbrain-incrementation/figures/pythonrunner.py

@@ -8,6 +8,11 @@ import json
 
 
 sl_fldr = '../home/val/spark-logs/'
+res_fldr = '../results/20bb_redo/'
+ad_bench = '../../system_conf/ad_diskbench.out'
+mem_bench = '../../system_conf/mem_diskbench.out'
+total_1000 = 617*1000
+total_125 = 614*125
 
 def get_json(log):
     data = []
@@ -53,6 +58,21 @@ for fn, sl in slogs:
     else:
         df_sl = pd.concat([df_sl, pd.DataFrame(data)])
 
+df_res = pd.concat(
+    (
+        pd.read_csv(
+            f,
+            delim_whitespace='spark' not in f,
+            names=["Task", "Start", "End", "File", "ThreadId"],
+        ).assign(
+            filename="{0}-{1}".format(i, op.basename(f)),
+            disk=op.basename(f).split("_")[0].split("-")[-1],
+        )
+    )
+    for i, f in enumerate(glob(op.join(res_fldr, 'spark*' + sys.argv[2] + '*')))
+)
+
+
 df_mksp = pd.DataFrame(smakespans)
 df_mksp_mem = df_mksp[~df_mksp['filename'].str.contains("AD")].groupby(['storage'])
 df_mksp_mem_mean = df_mksp_mem.mean()
@@ -94,6 +114,15 @@ df_ad['storage'] = df_ad['filename'].apply(lambda x: x.split('_')[0].split('-')[
 df_ad_mean = df_ad.groupby(['storage']).mean()
 df_ad_std = df_ad.groupby(['storage']).std()
 
+
+
+df_res['Duration'] = df_res['End'] - df_res['Start'] 
+
+df_res_mem_mean = df_res[~df_res["filename"].str.contains("AD")].groupby(['filename', 'disk'])['Duration'].sum().reset_index().groupby(['disk']).mean()
+df_res_ad_mean = df_res[df_res["filename"].str.contains("AD")].groupby(['filename', 'disk'])['Duration'].sum().reset_index().groupby(['disk']).mean()
+
+print(df_res_mem_mean)
+print(df_res_ad_mean)
 
 print('***Python Runner***\n\n*Memory Mode*\n\nMean:')
 print(df_mem_mean)
@@ -243,12 +272,29 @@ ax.set_xticklabels(['Isilon', 'Optane'])
 ax.legend()
 plt.savefig('sparklogs-{}.pdf'.format(sys.argv[2]))
 
+df_adb = pd.read_csv(ad_bench)
[…]

### 2019-11-21 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -9,6 +9,13 @@ from os import path as op
 import numpy as np
 
 
+ad_bench = '../../system_conf/ad_diskbench.out'
+mem_bench = '../../system_conf/mem_diskbench.out'
+data_size = 614 * 125
+total_1000 = 617 * 1000
+bb20 = '20bb' in sys.argv[1]
+
+
 def add_disk_col(df):
     df["disk"] = df.index.map(
         lambda x: x.split("_")[0].split("-")[-1] + ("-em" if "_em" in x else "-real")
@@ -147,15 +154,68 @@ def stacked_bar(df, spark=False):
 
     figure, ax = plt.subplots()
 
+    df_adb = pd.read_csv(ad_bench)
+    convert_gb = lambda x: float(x.strip(' GB/s')) * 1024 if 'GB/s' in x else float(x.strip(' MB/s'))
+    df_adb['Write'] = df_adb['Write'].apply(convert_gb)
+    df_adb['Read'] = df_adb['Read'].apply(convert_gb)
+
+    if bb20:
+        df_adb = df_adb[df_adb['Device'].str.contains('isilon') | df_adb['Device'].str.contains('optane')]
+
+    df_adb['WriteTime'] = data_size / df_adb['Write']
+    df_adb['ReadTime'] = data_size / df_adb['Read']
+    df_adb_mean = df_adb.groupby(['Device']).mean().sort_index(ascending=False)
+    df_adb_std = df_adb.groupby(['Device']).std().sort_index(ascending=False)
+
+    df_memb = pd.read_csv(mem_bench)
+    convert_gb = lambda x: float(x.strip(' GB/s')) * 1024 if 'GB/s' in x else float(x.strip(' MB/s'))
+    df_memb['Write'] = df_memb['Write'].apply(convert_gb)
+    df_memb['Read'] = df_memb['Read'].apply(convert_gb)
+
+    if bb20:
+        df_memb = df_memb[df_memb['Device'].str.contains('isilon') | df_memb['Device'].str.contains('tmpfs')]
+
+    df_memb['WriteTime'] = data_size / df_memb['Write']
+    df_memb['ReadTime'] = data_size / df_memb['Read']
+    df_memb_mean = df_memb.groupby(['Device']).mean().sort_index(ascending=False)
+    df_memb_std = df_memb.groupby(['Device']).std().sort_index(ascending=False)
+
+    if not bb20:
[…]

### 2019-12-10 paper.tex
added updated scripts and text

@@
     dataset using 25 and 96 parallel
-    processes in both cases. Our results show that performance of all available storage
-    devices (e.g. tmpfs, local disk and Isilon) is somewhat deteriorated by memory mode
-    when compared to App Direct Mode. In contrast, Optane in App Direct Mode was
-    found to be slower than local disk, when the load on the bus was high due to sheer
-    amount of concurrent processes and short task duration. It is believed that
-    Optane would likely be useful in the accessing of typically larger-than-memory
-    data in a limited-parallelism manner, such as with visualization applications.
-    Further experiments with larger datasets remain to be performed.
-
+    processes in both cases. Our results show that performance of applications executing on
+    Optane is superior to that of other storage devices, with the exception of DRAM. This is
+    the case in both Memory and App Direct Mode and irrespective of the amount of data and 
+    parallelism. Furthermore, Optane in App Direct Mode is believed to benefit from the use
+    of memory as a cache for writing in cases where output data is significantly smaller than
+    available memory. We believe the use of Optane will be beneficial to both neuroimaging applications
+    running on HPC or visualization of large, high-resolution images.
 \end{abstract}
@@
 Other storage devices included a \SI{240}{\gibi\byte} Micron SATA SSD, of which
-only \SI{149}{\gibi\byte} where available for the experiments, and a \SI{628}{\tebi\byte} shared Dell EMC Isilon cluster. The local disk
+only \SI{149}{\gibi\byte} where available for the experiments, and a
+\SI{720}{\tebi\byte} shared Dell EMC Isilon cluster. The Isilon cluster is an nsf4 mounted
+cluster (10Gb network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
 was the only storage device set up as a writeback device. Both Isilon and Optane in
@@
 
+There are some notable differences between the GNU Parallel and Apache Spark implementations.
+For instance, in the GNU Parallel implementation, all operations (i.e. reading, incrementing and writing)
+are done within the same task. The same task is applied to all the \bigbrain blocks and GNU Parallel is 
+provided with all the different command-line calls, processing a subset of them at a time,
+depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
+is achieved using the popular neuroimaging I/O: Nibabel. When provided with a filename, as is the case
+for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memmap the data
+using Numpy. This step will be referred to as ``load header''. It is only when the data is actually required
+(i.e. during incrementation), that the data will be loaded into memory.
+
+The Apache Spark implementation differs from that of GNU Parallel in that scheduling decisions
+were entirely differed to Spark. That is, rather than simply parallelizing the filenames, to be
+loaded in a similar fashion to that of GNU Parallel, we chose to load the using Spark's built-in
+\texttt{BinaryFiles}, which loads whole files, in binary format, into Spark partitions. As the
+data would have been pre-loaded by Spark, ``loading header'' only measures the time to
+convert the binary images into Nibabel objects, and ``incrementation'' consists purely
+of the time it took to increment the data. Therefore, only the write time would be
+comparable between the two implementations, as it is achieved using the same method.
+Furthermore, the Spark implementation also differs in that read, increment and write
+were separated into three map tasks, which can enable shuffling between the tasks should 
+it be determined as necessary by the Spark scheduler.
+
 Time is measured within the application using Python's \texttt{time} module 
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpu.pdf}
-    \captionsetup{width=\columnwidth}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpus.pdf}
+    \captionsetup{belowskip=-10pt}
     \caption{Makespan of the \bigbrain Incrementation application using 25 
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/violin-read-20bb_25cpu.pdf}
-    \captionsetup{width=\columnwidth}
-    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 2. 3 repetitions were performed.}\label{fig:stacked-25cpus}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_25cpus.pdf}
+    \captionsetup{belowskip=0pt}
+    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
+             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-25cpus}
 \end{figure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567709282-tmpfsAD_1it_25cpus_120delay_withpc-2.pdf}
-    \caption{tmpfs}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572380543-tmpfsAD_1it_25cpus_40bb-1.pdf}
+    \caption{DRAM}
 \end{subfigure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567705458-optaneAD_1it_25cpus_120delay_withpc-2.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382482-optaneAD_1it_25cpus_40bb-1.pdf}
     \caption{Optane}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567704324-localAD_1it_25cpus_120delay_withpc-2.pdf}
-    \caption{local disk}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382794-localAD_1it_25cpus_40bb-1.pdf}
+    \caption{local SSD*}
 \end{subfigure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567707833-isilonAD_1it_25cpus_120delay_withpc-2.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572388118-isilonAD_1it_25cpus_40bb-2.pdf}
     \caption{Isilon}\label{fig:gantt25isilon}
 \end{subfigure}
-\caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes and a sleep delay of 120s}\label{fig:gantt25}
+    \captionsetup{belowskip=-10pt}
+\caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes.
+*Some local SSD writes did not complete due to storage limitations.}\label{fig:gantt25}
 \end{figure*}
@@
 
-When observing the total task breakdown (Figure~\ref{fig:stacked-25cpus}), it is
-apparent that the differences between the makespans arise from difference in incrementation
-time between all devices except Isilon. In addition, while tmpfs and local disk have comparable write times, Optane
-has a more significant write time difference. Whereas, Isilon has more significant read and
-write time compared to all other devices, its incrementation time is comparable to that
-of local disk. Further inspection of each storage device's Gantt chart (see Figure~\ref{fig:gantt25})
-shows that increased duration of tasks stagger the execution of each individual tasks. This, in turn,
-results in better performance towards the end of the execution, where tasks are significantly more staggered.
-Only tmpfs appears to be able to execute each task efficiently with little staggering.
+The total task duration breakdowns for each device (Figure~\ref{fig:stacked-25cpus})
+shows that, as expected, I/O times vary greatly between devices, with DRAM having the
+best total read and write speeds. Optane in Memory Mode is very close to DRAM in speed,
+whereas Optane in App Direct mode 7x slower in terms of reading and 23x slower in terms
+of writing. Local SSD, on the other hand, exhibits similar write times as DRAM, but is
+approximately 96x that DRAM with respect to reads. Header loading for local SSD is also 90x longer
+than that of DRAM, however, remains negligible compared to other tasks. Isilon, on the 
+other hand, has non negligible header load times, with it being more than 5000 times
+slower than that of DRAM. Interestingly enough, the read/increment times on Isilon are
+slower than that of the local SSD, although Isilon makes up for its speedier reads with 
+writes that are around 184x slower than DRAM.
+
+The App Direct mode Gantt charts (Figure~\ref{fig:gantt25}) also reflect what is observed in Figure~\ref{fig:stacked-25cpus}.
+DRAM is significantly faster than all other storage, leading it to barely appear within the Gantt chart.
+DRAM was measured to have an average of 17.2 parallel tasks throughout the execution. Optane takes
+significantly longer than DRAM and appears to spend a significant amount of time writing. The average
+parallelism measured for Optane was 24. Unlike Optane, local SDD spends the vast majority of its time
+reading, while spending a negligible amount of time writing. Like Optane, it averages 24 parallel tasks
+throughout its execution. Unlike the other three storages, Isilon shows spacing between the loading of the header,
+read and increment and write. Isilon was measured to have an average parallelism of 14. Overall, each storage device
+execution displayed that some tasks took longer than others.
+
 \subsubsection{96 Processes}
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-96cpus-0delay-1it.pdf}
-    \captionsetup{width=\columnwidth}
-    \caption{Makespan of \bigbrain Incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations.
-5 repetitions were performed.}\label{fig:makespan-96cpus}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_96cpus.pdf}
+    \captionsetup{belowskip=-10pt}
+    \caption{Makespan of the \bigbrain Incrementation application using 96 
+             processes on the 125 blocks on the 40~$\mu$m image given different
+             storage configurations. Anticipated read and write duration was 
+             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
+             to Eq.~\ref{eq:makespan}. 3 
+             repetitions were performed. *Local SSD did not complete the
+    writing of the last few blocks due to storage limitations.}\label{fig:makespan-96cpus}
 \end{figure}
 \begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-96cpus-0delay-1it.pdf}
-    \captionsetup{width=\columnwidth}
-    \caption{total read/increment/write breakdown of \bigbrain incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations.
-5 repetitions were performed.}\label{fig:stacked-96cpus}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_96cpus.pdf}
+    \captionsetup{belowskip=-10pt}
+    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
+             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-96cpus}
 \end{figure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1568150082-local_1it_96cpus_0delay_npc-3.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572306855-tmpfs_1it_96cpus_40bb-2.pdf}
     \caption{Memory Mode}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567714252-localAD_1it_96cpus_0delay_npc-3.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572379747-optaneAD_1it_96cpus_40bb-1.pdf}
     \caption{App Direct Mode}
 \end{subfigure}
-\caption{Gantt charts for local disk processing 125 blocks of the 40$\mu$m BigBrain using 96 processes with no delay}\label{fig:gantt96}
+    \captionsetup{belowskip=-10pt}
+\caption{Gantt charts for Optane processing 125 blocks of the 40$\mu$m BigBrain using 96 processes}\label{fig:gantt96}
 \end{figure*}
 
-At 96 processes, it was expected that the application would not satisfy the inequality.
-In Figure~\ref{fig:makespan-96cpus} we see that there is now a greater disparity between the
-makespans of experiments executed in Memory and App Direct Mode, given the same storage. Local disk
-in App Direct Mode is significantly faster than its Memory Mode counterpart. Local disk in App Direct
-mode is even faster than Optane, by a factor of 1.7. Tmpfs was also found to perform better in App Direct Mode, although
-the performance difference is not significant. Only Isilon's performance did not vary between the
-two execution modes.
+Despite having increased parallelism by a factor of 3.84, we see no visible reduction
+in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 cpus, DRAM and Optane in Memory Mode are the only storage
+devices that did better than the anticipated I/O, with a makespan duration of
+5.7 and 9s, respectively (total read and write estimate for both was measured to be around 39s). While
+Optane in App Direct Mode and local SSD both performed worse than the anticipated I/O estimates, Isilon
+performed as expected. Variance was found to be high on the local SSD.
 
-As with 25 processes, the main cause of performance variability appears to be the incrementation
-time (Figure~\ref{fig:stacked-96cpus}). Local disk running in App Direct Mode has an average
-incrementation time that is nearly 15~x faster than in Memory Mode. The incrementation time
-for local disk is, on average, 6~x faster than on Optane.
+The total task duration breakdowns (Figure~\ref{fig:stacked-96cpus}) show that total task durations
+have nearly quadrupled. Whereas most devices decreased in performance in both reading and writing,
+Isilon only appeared to display a performance decrease with respect to writes. Furthermore, while both
+read and writes decreased for Optane in App Direct Mode, read duration was nearly 6x slower with more threads,
+while write duration was only about 2x slower. For local SSD and Isilon, there does not appear to 
+be any significant difference between App Direct and Memory mode.
 
-Isilon in App Direct generally performed the same as in memory mode, however, the write
-was almost 4000~seconds slower, on average, in App Direct Mode.
+When analyzing the average parallelism during the workload, only Optane in App Direct mode
+came close to 96 parallel tasks, having an average parallelism of 88 tasks. Optane in Memory mode,
+DRAM and local SSD all averaged between 61-66 tasks. Isilon performed the worst with 38 tasks in Memory 
+mode and 45 tasks in App Direct mode.
 
-A further look at the Gantt charts between Memory Mode and App Direct Mode (Figure~\ref{fig:gantt96})
-shows that the sole difference between the two modes is the duration of the incrementation step.
-While both modes have an increased incrementation time, the incrementation duration in
-Memory Mode is much more important than that of App Direct Mode.
+A further look at the Gantt charts between Optane in Memory Mode and App Direct Mode (Figure~\ref{fig:gantt96})
+show that both read and write time of tasks are worse in App Direct mode as compare with Memory mode.
+While some reads tasks in App Direct mode appear to be of similar duration to the average read task time
+of Memory mode, the vast majority of read tasks are significantly slower. Futhermore, not write tasks in App
+Direct mode is capable of reaching the performance of equivalent tasks in Memory mode.
 
@@
 
-Although tmpfs offers the best performance in terms of makespan duration, Optane
-is a close contender. While tmpfs is approximately 5x faster than Optane, Optane can be
-as much as 12x faster than on Isilon (Figure~\ref{fig:20mksp25}). By using the results of 
-our emulation, we can speculate that Optane will also be significantly faster than 
-local disk in a real scenario. The real Optane execution is almost 3x faster than
-the fastest local disk emulation configuration (i.e. Memory mode - Spark). As the
-emulation results have been found to be faster than the real results, potentially as a side-effect
-of increased accessed to cached data, it is expected
-that the true makespan of the application running on local disk would be significantly
-greater.
+The anticipated I/O estimates did not correctly predict the makespan for 25 parallel
+processes (Figure~\ref{fig:20mksp25}). Both Optane in Memory Mode and Isilon appear 
+to be under the estimates: Optane in Memory Mode is approximately twice as fast as the
+estimates whereas Isilon is around 1.1x faster. Optane was 1.8x slower than anticipated.
+Once again, Memory and App Direct Mode appear to only affect performance for Optane.
 
-The read/increment/write breakdown (Figure~\ref{fig:20stackedp25}) shows that tmpfs
-spends the least amount of time executing all the tasks, however, it spends more
-time than Optane incrementing the data. Otherwise, tmpfs is nearly 470x faster with
-reading and nearly 11x faster at writing than Optane. Despite not reaching memory
-speeds, Optane is significantly faster than other persistent storage devices.
-For instance, it can be almost up to 10x faster in reads than Isilon and almost 
-9x faster in writes. It is important, however, to note that total read times in App
-Direct mode for the Spark execution on Isilon are almost twice as fast as their 
-equivalent Optane counterpart. On the other hand, this decrease read time is offset
-by the write time which are 9x slower.
+Total task duration breakdowns (Figure~\ref{fig:20total25}) show that for Optane in
+App Direct Mode spends significantly more time writing than its Memory Mode counterpart.
+In fact, it spends almost 12x more time writing. Isilon also appears to vary slightly
+between Memory and App Direct Mode, spending around 1.4x more time reading and incrementing
+in App Direct Mode. While the read and write duration appeared to be more-or-less balanced in
+Isilon (as was anticipated), Optane in App Direct Mode spent significantly more time writing, whereas the 
+inverse was apparent in Memory Mode
 
-Emulated breakdown results also agree that the bulk of the differences in makespans
-can be attributed to write time, with local disk writes being notably slower than
-Optane. In fact, local disk write speeds nearly match that of Isilon. Isilon results
-in a longer makespan due to increase read and incrementation time.
+Unlike the GNU Parallel executions, the Spark executions appear to be in better accordance with
+the I/O estimates (Figure~\ref{fig:20mkspspark25}. Only Optane in App Direct Mode appears
+to have significantly exceeded the estimates, by a factor of 2. Total write duration all appear
+to be longer in the Spark implementation than with GNU Parallel (Figure~\ref{fig:20totalspark:25}).
+Spark's writes are 1.4-1.6x longer with Optane and 1.1x longer with Isilon. Due to differences in
+how data is loaded in Spark, it is not possible to comment on how read times are affected. However, it
+is possible to note that data conversion (binary string to Numpy array) takes almost twice as 
+much time in Optane than it does with Isilon. Furthermore, the act of incrementation can take
+up to twice as much time with Isilon.
+
 \begin{figure*}
@@
     \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_25cpus.pdf}
-    \caption{real}
+    \caption{Makespan}\label{fig:20mksp25}
 \end{subfigure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-em-20bb_25cpus.pdf}
-    \caption{emulated}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_25cpus.pdf}
+    \caption{Total read/increment/write breakdown}\label{fig:20total25}
 \end{subfigure}
-\caption{Makespan of the incrementation algorithm processing the 20~$\mu$m BigBrain using
-25 processes}\label{fig:20mksp25}
+\captionsetup{belowskip=-10pt}
+\caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
+25 processes. 3 repetitions were performed.}\label{fig:2025}
 \end{figure*}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_25cpus.pdf}
-    \caption{real}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-spark_20bb_25cpus.pdf}
+    \caption{Makespan}\label{fig:20mkspspark25}
 \end{subfigure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-emulated-20bb_25cpus.pdf}
-    \caption{emulated}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark-20bb_25cpus.pdf}
+    \caption{Total convert/increment/write breadown}\label{fig:20totalspark:25}
 \end{subfigure}
-\caption{Total read/increment/write breakdown of the  incrementation algorithm processing the 20~$\mu$m BigBrain using
-25 processes}\label{fig:20stackedp25}
+\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 25 processes. 3 repetitions wereperformed}\label{fig:20spark25}
 \end{figure*}
@@
 
-With nearly a four-fold increase in parallelization, the makespan generally did
-not improve other but a few 100s on Isilon. Overall, the general trend is similar
-than with 25 cpus. As can be seen in Figure~\ref{fig:20mksp96}, Optane in Memory
-mode is the fastest storage device. Furthermore, Optane in App Direct mode 
-remains significantly faster than the other storage devices (e.g. Isilon).
-However, emulated disk results appear to imply that the speedup difference
-between Optane in App Direct mode and local disk will be smaller than at 25 cpus.
+A very slight performance improvement (1-1.3x faster) can be witnessed in Optane. In 
+contrast, Isilon 
+experienced a slight performance decrease (1.1x slower). Whereas Isilon makespan appears to match
+estimated I/O times, Optane once again has a superior Memory Mode makespan and an inferior App Direct
+Mode makespan (Figure~\ref{fig:20mksp96}). As previously observed with 25 processes,
+the bulk of the processing time is taken up by writing for Optane in App Direct Mode and 
+Isilon in both modes (Figure~\ref{fig:20total96}). Reading takes more time in the case of Optane in Memory mode.
 
-The vast majority of processing time appears to be spent on writing to the 
-respective storage device with the exception of Optane in Memory Mode, which 
-spends most of its time incrementing (Figure~\ref{fig:20stackedp96}). App Direct
-mode executions appear to have reduced incrementation time coupled with increased
-write times when compared to their memory mode equivalents. Furthermore, read
-times appear to be increased for Isilon memory mode executions. These events are
-not reflected in the emulated executions.
-
-Emulated breakdowns demonstrate that local disk executions will remain slower 
-than Optane in both App Direct and Memory modes. Results show that both 
-incrementation and write times will be generally be longer on local disk than
-on Optane.
+A notable difference in the Spark execution is that Optane in Memory and App Direct Mode
+exhibit the same performance (Figure~\ref{20mkspnspark96}). Both of which also performed
+significantly worse than the estimated I/O duration. Performance on Isilon, however, did
+not really differ. The total task breakdown shows that, as expected, write times are more
+significant on Isilon than on Optane and Optane in App Direct Mode spends more time writing
+than in Memory Mode (Figure~\ref{20fig:20totalspark96}). When comparing GNU Parallel and 
+Spark implementations, Spark spent nearly half as much time writing as GNU Parallel. However,
+it was found that Optane spent more time converting the data to Nibabel than GNU Parallel spent
+in loading the header.
 
@@
     \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_96cpus.pdf}
-    \caption{real}
+    \caption{Makespan}\label{fig:20mksp96}
 \end{subfigure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-em-20bb_96cpus.pdf}
-    \caption{emulated}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_96cpus.pdf}
+    \caption{Total read/increment/write breakdown}\label{fig:20total96}
 \end{subfigure}
-\caption{Makespan of the incrementation algorithm processing the 20~$\mu$m BigBrain using
-96 processes}\label{fig:20mksp96}
+\caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
+96 processes. 3 repetitions were performed.}\label{fig:2096}
 \end{figure*}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_96cpus.pdf}
-    \caption{real}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-spark-20bb_96cpus.pdf}
+    \caption{Makespan}\label{fig:20mkspspark96}
 \end{subfigure}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-emulated-20bb_96cpus.pdf}
-    \caption{emulated}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark-20bb_96cpus.pdf}
+    \caption{Total convert/increment/write breakdown}\label{fig:20totalspark96}
 \end{subfigure}
-\caption{Total read/increment/write breakdown of the  incrementation algorithm processing the 20~$\mu$m BigBrain using 96 processes}\label{fig:20stackedp96}
+\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 96 processes. 3 repetitions were performed}\label{fig:20stackedp96}
 \end{figure*}
@@
 \end{figure*}
-Unlike the BigBrain incrementation at 25 Processes, the makespan of the BIDS App Example
+Unlike the BigBrain incrementation, the makespan of the BIDS App Example
 executing on Optane is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
-to the BigBrain Incrementation, tmpfs has a slight lead over local disk and Isilon performs the
-worst. The same pattern can be found with 96 process (Figure~\ref{fig:bm96}). When comparing
+to the BigBrain Incrementation, Optane in Memory Mode performs better than the other storage devices
+while Isilon has the longest makespan. The same pattern can be found with 96 process (Figure~\ref{fig:bm96}). When comparing
 the executions of 25 and 96 processes, the performance improved by a similar factor on all
@@
 
-Memory mode was found to result in increased computation time. When processing 
-the 20~$\mu$m \bigbrain, Memory mode results indicate that more time was spent on
-the incrementation tasks than its App Direct counterparts. Unlike App Direct mode,
-which relies exclusively on DRAM for working memory, Memory mode relies on Optane
-with a DRAM cache. Optane, while faster than other persistent storage devices, 
-still lacks the bandwidth of DRAM. Should anonymous memory need to access Optane
-for computation, it is not surprising that it would be slower than a DRAM-only
-access. However, whereas Memory mode experienced increased computation during the
-processing of the 20~$\mu$m \bigbrain, this was not always found to be the case
-with the processing of the 40~$\mu$m \bigbrain. In fact, incrementation time was
-largely similar between the two modes, with the notable exception of Optane in
-Memory mode expending significantly less computation time than in App Direct when
-processing the 40~$\mu$m \bigbrain with both 25 and 96 processes. Optane in 
-Memory mode executing on local disk using 96 cpus, however, required significant
-computation time as compared to App Direct mode. For all GNU Parallel runs, the 
-read only consisted of loading metadata into
-- Conclusion: memory mode should only be used when there's a requirement for additional memory
--App Direct mode results in increased I/O time for Optane
-- conclusion: Optane might benefit from the DRAM cache
+In general, the selection of storage mode did not affect overall performance. This can be
+due to several reasons. For instance, the amount of available DRAM was abundant and exceeded 
+all our dataset sizes. While total memory used by the application may have exceeded DRAM, as is the
+case with the 20~$\mu$ \bigbrain, the amount of memory required by the application at any 
+given moment did not exceed the amount of available DRAM. Although GNU Parallel would have
+been able to load all the blocks in memory, had we increased the block size, it would have been unlikely
+that total available DRAM would have been exceeded, though it would have occurred in Memory Mode
+for the devices that leverage memory as a write cache. In our case, only the local SSD was set up to do so.
 
-While Memory Mode provides a cost-effective solution to increasing the total memory
-size, it is slower than App Direct Mode for embarrassingly parallel applications such
-as \bigbrain incrementation. This is likely due to the Page Cache being composed of a
-combination of both DRAM and Optane in Memory Mode. Since the data produced by the application should be
-able to fit within DRAM, writing to a slower bandwidth device such as Optane inevitably
-deteriorates performance. As App Direct separates usage of DRAM and Optane, devices only utilize DRAM in this mode.
-The only device that performed better in Memory Mode was 
-Isilon, however the difference is minimal.
-%This may be explained by a larger Page Cache in Memory Mode, enabling Isilon to
-%store more data in memory prior to initiating synchronous writes to Isilon.
+Optane was the only device affected by the choice of storage mode. In all cases, with the 
+exception of the Spark incrementation of the 20~$\mu$ \bigbrain, Memory Mode was superior to 
+App Direct Mode. This is a result of Optane using DRAM as cache in Memory Mode, whereas Optane
+was configured as a writethrough device in App Direct Mode. It is believed that had Optane been
+configured to use DRAM as a writeback cache in App Direct Mode, we  would have observed a similar
+performance between the two. However, for the 40~$\mu$ \bigbrain executions, read time on Optane in
+App Direct Mode appeared to be slower than Memory Mode, potentially indicating that the Optane was able
+to keep the input data cached in DRAM whereas in App Direct Mode, the data had to be loaded from Optane.
+Therefore, it is expected the Optane in App Direct mode may always have longer read times than Memory Mode
+if the data is already cached in DRAM in Memory Mode.
 
-I/O-less task duration also seem to be affected by the execution mode, with those tasks
-also performing significantly better in App Direct. A potential explanation for this may be 
-that all memory accesses to Optane are slower due to the device's bandwidth, meaning that
-tasks that do not perform I/O, but access data loaded into memory are slower in Memory Mode. Another explanation might be that the kernel is throttling the processes
-as they are producing significant amounts of output, thereby increasing the
-incrementation time.
-This can be seen in Figure~\ref{fig:gantt96} where for local disk, incrementation in App Direct
-mode is significantly shorter. Only when the memory bus is overwhelmed does incrementation take
-longer for App Direct, however, it is nothing compared to the average duration of the incrementation
-tasks in Memory Mode.
-
-Once again, Isilon is the only device which does not experience an increase in duration for
-I/O-less tasks. In fact, the incrementation duration is less than that of local disk in Memory Mode
-and Optane in App Direct Mode. It may be that since even the reads are staggered
-(Figure~\ref{fig:gantt25isilon}), the memory bus is less strained by all the parallel processes,
-therefore incurring less of an overall delay for all memory accesses.
+The Apache Spark and GNU Parallel implementations differed vastly in makespan 
+for Optane in Memory Mode, particularly at 96 processes. After taking a closer look at
+the captured Spark metrics, it was for that what really varied between the Spark runs on
+the different storage was garbage collection, with garbage collection taking longer on Optane 
+than on Isilon. Furthermore, garbage collection was longer in Optane Memory Mode than in 
+Optane App Direct Mode. It is suspected that, in these cases, task duration was so short
+that garbage collection could not keep up. Moreover, Optane in Memory Mode may have longer
+garbage collection if old anonymous data cannot be located in the DRAM cache due to overall memory 
+requirements, therefore requiring garbage collection to occur on slower memory.
 
@@
 
-As long as all the written data does not exceed the dirty ratio, there is little 
-benefit brought by the use of Optane. This is the case with both
-our applications writing less than $\SI{1}{\gibi\byte}$ to storage on a system with 
-$\SI{768}{\gibi\byte}$ of DRAM and a dirty ratio of 20. However, it is expected 
-that as the memory footprints of applications increase to the point where the kernel
-would be forced to perform synchronous writes to disk if it were to rely on DRAM alone,
-the benefit of using Optane in Memory Mode will become more apparent.
+The use of Optane in Memory Mode would enable more dirty data to be written to available memory
+rather than directly to slower storage devices. In our 40~$\mu$ \bigbrain experiments, it was found
+that read times were less important in Isilon than the local SSD (Figures~\ref{fig:stacked-25cpus}
+and~\ref{fig:stacked-96cpus}). Should Isilon have been given the opportunity to benefit from a 
+writeback cache like the SSD, it is believed that performance on Isilon could have been superior 
+to that of the SSD. Having Optane as a node-local burst buffer to shared network storage 
+may prove to be very beneficial in the case of Big Data neuroimaging pipelines running on
+High Performance Computing clusters.
 
-However, depending on the neuroimaging application, the amount of data that could be
-loaded in parallel would be limited by the amount of parallelism possible on the system.
-For instance, typical fMRI tasks process many small images (100s of MB in size) in parallel.
-While the datasets themselves can reach PB in size, the amount of data that can be processed in
-parallel can be $\SI{10}{\giga\byte}$ if each take processes $\SI{100}{\mega\byte}$ and the 
-maximum amount of parallelism is 100.
+For App Direct Mode applications (although also applicable to Memory Mode), the 
+choice of using the writeback cache depends on total amount of available memory, application
+memory usage and I/O. Should the application be able to use the page cache without ever being
+I/O-throttled, using a write-back cache will significantly improve performance. This can 
+be seen in Figure~\ref{fig:stacked-25cpus}, for instance, where the local SSD is writing
+entirely to memory, reducing significantly the total write time. However, should write operations
+become throttled, this may significantly slow down the write task duration. In such cases,
+it may be preferable to set them up as write-through devices. As Big Data neuroimaging applications
+tend to generate temporary data files that are larger than input size, it is likely that maintaining the
+devices as write-through is preferable. Furthermore, despite that Optane in App Direct Mode does
+not leverage the writeback cache like the local SSD, is it still found to be superior in performance.
+Therefore, the while the performance of Optane may not be enhanced by DRAM, it is still superior to that of
+SSDs using the writeback cache and scalable network storage.
 
-On the other hand, Memory Mode can be very useful for applications operating on
-ultra-high resolution images, like the \bigbrain. For instance, visualization applications
-many need to load significant amounts of data in memory. Given the size of these datasets, 
-extending the amount of available memory using Optane could be quite beneficial and allow
-for faster visualization of larger datasets.
+\subsection{Device scalability}
+The number of storage devices attached to a particular filesystem mountpoint varied between
+each storage device. For instance, DRAM consisted of 12 devices, whereas although there were 
+12 Optane devices, only 6 were accessible by the mountpoint selected in App Direct Mode. Furthermore,
+local SSD only consisted of a single device, whereas the Optane server was made up of 180 SSDs. The differences
+in number of storage devices would account for the attainable amount of parallel writes. For instance, 
+Isilon would have been the most scalable, followed by DRAM, Optane and the local SSD, which would
+only be able to process I/O sequentially. This explains why the anticipated I/O time, is sometimes greater
+than the real makespan of the pipeline.
+
+For instance, taking into consideration the total number of DRAM devices, the expected makespan of the
+40~$\mu$ \bigbrain is estimated to take around 3s, whereas sequentially it was measured to take
+around 37.5s. At 25 processes, we found the DRAM makespan to be around 7s, which is much closer to the 3s
+estimate than the 37.5s. Reasons for why it took longer that estimated could be due to application
+overheads, or even that some DRAM devices were occupied, and therefore, the maximum amount of parallelism 
+was not achievable.
+
+Despite the fact that only sequential I/O was considered for the anticipated I/O times, 
+Optane in App Direct Mode always took more time than its sequential estimates. The cause
+of the longer-than-expected makespan seems to be due to write times. While there may be application overheads,
+they should, in theory, affect Optane in App Direct Mode as much DRAM. While it is unknown
+why Optane writes are performing so poorly, it could be a result of filesystem configuration or 
+inaccurate benchmarking of the disk. Otherwise, Optane does not appear to scale very well with respect to
+writes. Using the storage benchmarks found in Table~\ref{table:bandwidths}, it is found that with an equivalent
+number of storage devices, processing the 40~$\mu$ \bigbrain using 25 processes would be faster using
+local disk when compared to the real times obtained on Optane. As there are overheads with both Optane and local
+SSD, it is uncertain that it would, in fact, be the case.
+
+While Isilon is, in theory, the most scalable device, its performance is expected. The Isilon
+server is made up entirely of Hard Disks Drives, which are the slowest type of device compared to
+the others, as seen in Table~\ref{table:bandwidths}. While Isilon displays a sufficient amount of
+scalability to read and write the data in a few seconds, it is a network-backed device. As a result,
+performance of this device is limited by the network speed of 10Gbps. Consequently, Isilon performs at around
+the same rate as sequential writes.
+
+Another interesting aspect of Isilon is the spacing that was observed between the read and writes
+(Figure~\ref{fig:gantt25isilon}). What occurred between the read/increment and
+write operations was simply writing the previous task's benchmarks to a unique
+file. This did not appear to affect any other storage device. This appears to imply
+that the latency alone of writing to Isilon was quite significant.
 
@@
 
-The added value of Optane as a persistent storage device depends on the amount of parallelism
-of the application. For instance, Optane performs better than local disk at 25 processes,
-but worse than local disk at 96 processes. This can be explained by the fact that it performs
-I/O on the memory bus. While this may reduce overall latency to Optane, the bandwidth of 
-writing to Optane is still significantly less than DRAM. Despite Optane having a similar bandwidth
-to local disk, local disk uses DRAM as cache whereas Optane does not. This fact is 
-particularly evident with the amount of data that is written in parallel at 96 processes. With 25 processes
-and at least a 2 min
-delay between each write, the load on the memory bus is low enough to be efficiently handled.
+Our results show that Optane has superior performance to that of other storage devices, with the
+exception of DRAM. Optane as a persistent memory storage device is expected to bring significant
+performance improvements to the processing of neuroimaging data. Input datasets to neuroimaging pipelines
+are increasing rapidly in size. Storing such datasets directly on Optane would significantly reduce the
+impacts of I/O on the processing. If Optane is located as a persistent memory storage device in HPC clusters,
+it could also be used as a burst buffer to network-attached storage devices. This would vastly
+improve the performance of standard neuroimaging pipelines, as they produce temporary data files that 
+are larger than the input dataset. Writing these temporary files to slow, network-attached storage, can have
+significant impacts on the performance of a pipeline.
 
-fMRI application will generally attempt to leverage as much parallelism as possible. As this
-is the case, it is not believed that App Direct Mode would bring a significant performance advantage
-unless the server's CPUs are limited, in which case there is less concurrent accesses on the device.
-The same would apply to ultra-high resolution application unless they are sequential, as may be the case with
-visualization.
+In the case of high-resolution images, Optane can be leveraged to enable the users to rapidly
+extract their regions of interest with minimal I/O costs. Futhermore, Optane would improve the
+speed and fluidity of visualization applications on these datasets.
+
 \section{Conclusion}
 
-The benefits of Optane for neuroimaging applications are limited to large memory
-accesses with limited parallelism, as would be the case for visualization. This is
-the case for both Memory Mode and App Direct Mode.
+Optane has been found to drastically reduce the processing time of neuroimaging applications and
+can bring the performance close to DRAM speeds. Extending available memory using Optane is also 
+expected to help reduce I/O times of write-back devices by potentially extending available cache space.
+It is unclear, however, if traditional SSDs can perform just as well as Optane with additional storage devices
+attached.
 
-While Memory Mode leads to deterioration in performance of non-Optane storage devices,
-they act as expected in App Direct Mode. It is currently unknown by how much performance
-of a non-Optane storage device would be affected by a large memory access.
-
-To be able to fully determine the benefits of Optane for neuroimaging application,
-it is necessary to process larger datasets. We plan to execute incrementation pipeline on the 
-20~$\mu$m BigBrain using Apache Spark. 
+It is believe Optane can be useful for a variety of neuroimaging applications. For
+instance, typically applications executing in HPC environments can benefit from 
+the speedups provided by Optane. Furthermore, image visualization servers will be able
+to display higher resolution images at significantly greater speeds.
 \section{Acknowledgement}
+We are thankful to Intel and Dell
+for the 90-day trial of the Intel Optane DC Server and high-quality technical support.
 \bibliographystyle{IEEEtran} 

### 2019-12-10 experiments/bids-app-example/plot.py

@@ -6,7 +6,7 @@ from statistics import mean, stdev
 
 conditions_96_jobs = ["in-mem-1", "pmem-1", "disk-1", "isilon-1"]
 conditions_25_jobs = ["in-mem-4", "pmem-4", "disk-4", "isilon-4"]
-condition_ticks = ["tmpfs", "Optane", "disk", "Isilon"]
+condition_ticks = ["Optane MM", "Optane ADM", "local SSD MM", "Isilon MM"]
 width = 0.35
 
 def get_makespans(conditions):
@@ -44,7 +44,7 @@ def get_makespans(conditions):
 mksp_96 = get_makespans(conditions_96_jobs)
 plt.ylabel("Makespan (s)")
 plt.ylim((0,5000))
-plt.bar([1, 2, 3, 4], mksp_96[0], tick_label=condition_ticks, width=width, yerr=mksp_96[1], alpha=0.4, color='gray')
+plt.bar([1, 2, 3, 4], mksp_96[0], tick_label=condition_ticks, width=width, yerr=mksp_96[1], alpha=0.4, color='purple')
 plt.savefig("96cores.pdf")
 
 plt.clf()
@@ -52,7 +52,7 @@ plt.clf()
 mksp_25 = get_makespans(conditions_25_jobs)
 plt.ylabel("Makespan (s)")
 plt.ylim((0,5000))
-plt.bar([1, 2, 3, 4], mksp_25[0], tick_label=condition_ticks, width=width, yerr=mksp_25[1], alpha=0.4, color='gray')
+plt.bar([1, 2, 3, 4], mksp_25[0], tick_label=condition_ticks, width=width, yerr=mksp_25[1], alpha=0.4, color='purple')
 plt.savefig("25cores.pdf")
 
 def time_to_secs(t):

### 2019-12-10 experiments/bigbrain-incrementation/figures/makespan.py

@@ -14,6 +14,7 @@ mem_bench = '../../system_conf/mem_diskbench.out'
 data_size = 614 * 125
 total_1000 = 617 * 1000
 bb20 = '20bb' in sys.argv[1]
+disks = ["tmpfs", "tmpfsAD", "optaneAD", "local", "localAD", "isilon", "isilonAD"]
 
 if bb20:
     data_size = total_1000
@@ -57,9 +58,7 @@ def makespan(df, task="task_duration", spark=False):
         dfp = df[df["Task"] == "task_duration"].groupby(["filename"])
     else:
         dfp = df[(~df["filename"].str.contains('spark')) & (df["Task"] == "task_duration") & ~(df["filename"].str.contains('em'))].groupby(["filename"])
-        dfpe = df[(~df["filename"].str.contains('spark')) & (df["Task"] == "task_duration") & (df["filename"].str.contains('em'))].groupby(["filename"])
         dfs = df[df["filename"].str.contains('spark') & ~(df["filename"].str.contains('em'))].groupby(["filename"])
-        dfse = df[df["filename"].str.contains('spark') & (df["filename"].str.contains('em'))].groupby(["filename"])
 
     ind = np.arange(len(labels))
 
@@ -84,13 +83,9 @@ def makespan(df, task="task_duration", spark=False):
         print(df_pres)
     else:
         df_sres = configure_df(dfs)
-        df_peres = configure_df(dfpe)
-        df_seres = configure_df(dfse)
 
         print(df_pres)
         print(df_sres)
-        print(df_peres)
-        print(df_seres)
 
 
         df_pres_mem = df_pres[
@@ -102,16 +97,6 @@ def makespan(df, task="task_duration", spark=False):
             df_sres.index.map(lambda x: 'AD' not in x)
         ]
         df_sres_mem = df_sres_mem.sort_index(ascending=False)
-        
-        df_pres_mem_em = df_peres[
-            df_peres.index.map(lambda x: 'AD' not in x)
-        ]
-        df_pres_mem_em = df_pres_mem_em.sort_index(ascending=False)
-
-        df_sres_mem_em = df_seres[
-            df_seres.index.map(lambda x: 'AD' not in x)
-        ]
-        df_sres_mem_em = df_sres_mem_em.sort_index(ascending=False)
 
         df_pres_ad = df_pres[
             df_pres.index.map(lambda x: 'AD' not in x)
@@ -122,11 +107,6 @@ def makespan(df, task="task_duration", spark=False):
[…]

### 2019-12-10 experiments/bigbrain-incrementation/figures/pythonrunner.py

@@ -65,11 +65,11 @@ df_res = pd.concat(
             delim_whitespace='spark' not in f,
             names=["Task", "Start", "End", "File", "ThreadId"],
         ).assign(
-            filename="{0}-{1}".format(i, op.basename(f)),
+            filename="{0}".format(op.basename(f)),
             disk=op.basename(f).split("_")[0].split("-")[-1],
         )
     )
-    for i, f in enumerate(glob(op.join(res_fldr, 'spark*' + sys.argv[2] + '*')))
+    for f in glob(op.join(res_fldr, 'spark*' + sys.argv[2] + '*'))
 )
 
 
@@ -142,210 +142,286 @@ print('\nSTD:')
 print(dfsl_ad_std)
 
 
-figure, ax = plt.subplots()
-width = 0.4
-ind = np.asarray([2, 1])
+def plot_figs():
 
-p_total = ax.bar(
-    ind - width/2,
-    df_mem_mean['total'],
-    width,
-    color="g",
-    label="total",
-    yerr=df_mem_std['total']
-)
-p_total_ad = ax.bar(
-    ind + width/2,
-    df_ad_mean['total'],
-    width,
-    color="g",
-    label="total AD",
-    alpha=0.4,
-    yerr=df_ad_std['total']
-)
-p_init = ax.bar(
-    ind - width/2,
-    df_mem_mean['init'],
-    width,
-    color="r",
-    label="init",
-    yerr=df_mem_std['init']
-)
-p_init_ad = ax.bar(
-    ind + width/2,
-    df_ad_mean['init'],
-    width,
-    color="r",
-    label="init AD",
-    alpha=0.4,
-    yerr=df_ad_std['init']
-)
-p_finish = ax.bar(
-    ind - width/2,
-    df_mem_mean['finish'],
-    width,
-    bottom=df_mem_mean['init'],
-    color="b",
-    label="finish",
-    yerr=df_mem_std['finish']
-)
-p_finish_ad = ax.bar(
-    ind + width/2,
-    df_ad_mean['finish'],
-    width,
-    bottom=df_ad_mean['init'],
-    color="b",
-    label="finish AD",
-    alpha=0.4,
-    yerr=df_ad_std['finish']
-)
+    figure, ax = plt.subplots()
+    width = 0.4
+    ind = np.asarray([2, 1])
 
-ax.set_ylabel("Average total execution time (s)")
-ax.set_xlabel("Storage type")
-ax.set_xticks(ind)
[…]

### 2019-12-10 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -58,7 +58,9 @@ def get_mean_std(df, r_task, i_task, w_task):
 
 def stacked_bar(df, spark=False):
 
-    labels = ["DRAM", "Optane", "local disk", "Isilon"]
+    global data_size, total_1000
+
+    labels = ["DRAM", "Optane", "local SSD", "Isilon"]
     ind = np.arange(len(labels))
     ind_mem = np.delete(ind, 0)
 
@@ -67,6 +69,8 @@ def stacked_bar(df, spark=False):
     w_task = "write_file"
 
     dfp = df[~df["filename"].str.contains('spark')]
+
+    print('**GNU Parallel**\n\n')
     df_mean, df_std = get_mean_std(dfp, r_task, i_task, w_task)
 
     def prep_mem_df(df, exec_type=None):
@@ -86,17 +90,14 @@ def stacked_bar(df, spark=False):
         df_std_mem = prep_mem_df(df_std)
     else:
         dfs = df[df["filename"].str.contains('spark')]
+        print('**Spark**\n\n')
         df_smean, df_sstd = get_mean_std(dfs, "load_img", "increment_data", "save_incremented") 
 
         df_mean_mem_r = prep_mem_df(df_mean, "real")
-        df_mean_mem_em = prep_mem_df(df_mean, "em")
         df_std_mem_r = prep_mem_df(df_std, "real")
-        df_std_mem_em = prep_mem_df(df_std, "em")
 
         df_smean_mem_r = prep_mem_df(df_smean, "real")
-        df_smean_mem_em = prep_mem_df(df_smean, "em")
         df_sstd_mem_r = prep_mem_df(df_sstd, "real")
-        df_sstd_mem_em = prep_mem_df(df_sstd, "em")
 
     def prep_ad_df(df, exec_type=None):
         df = df[df.index.map(lambda x: "AD" in x and (exec_type is None or exec_type in x))]
@@ -139,14 +140,10 @@ def stacked_bar(df, spark=False):
         df_std_ad = prep_ad_df(df_std)
     else:
         df_mean_ad_r = prep_ad_df(df_mean, "real")
-        df_mean_ad_em = prep_ad_df(df_mean, "em")
         df_std_ad_r = prep_ad_df(df_std, "real")
-        df_std_ad_em = prep_ad_df(df_std, "em")
 
         df_smean_ad_r = prep_ad_df(df_smean, "real")
[…]

### 2019-12-10 scripts/gantt.py

@@ -15,8 +15,13 @@ import numpy as np
 def gantt_increment(df, data_file, ncpus, spark=False):
     color = {
         "read_file": "red",
-        "increment_file": "blue",
+        "increment_file": "red",
         "write_file": "green",
+    }
+    alpha = {
+        "read_file": 0.4,
+        "increment_file": 1,
+        "write_file": 0.4,
     }
 
     if spark:
@@ -51,24 +56,31 @@ def gantt_increment(df, data_file, ncpus, spark=False):
         print(df_tasks)
         return df_tasks['C'].sum()/df_tasks['Time'].max()
 
-    print('number of overlaps:', count_overlaps(df))
+    #print('number of overlaps:', count_overlaps(df))
 
-    print('Number of threads:', len(df.groupby("ThreadId")))
+    #print('Number of threads:', len(df.groupby("ThreadId")))
     for i, task in enumerate(df.groupby("ThreadId")):
         labels.append(task[0])
         for r in task[1].groupby("Task"):
             data = r[1][["Start", "Duration"]]
-            ax.broken_barh(data.values, (i - 0.4, 0.8), color=color[r[0]], alpha=0.4)
+            ax.broken_barh(data.values, (i - 0.4, 0.8), color=color[r[0]], alpha=alpha[r[0]], edgecolor='none')
 
-    r_read = mpatch.Patch(facecolor='red', alpha=0.4, label='Read')
-    r_inc = mpatch.Patch(facecolor='blue', alpha=0.4, label='Increment')
-    r_write = mpatch.Patch(facecolor='green', alpha=0.4, label='Write')
+    r_read = mpatch.Patch(facecolor='red', alpha=0.4, label='load header')
+    r_inc = mpatch.Patch(facecolor='red', alpha=1, label='read/increment')
+    r_write = mpatch.Patch(facecolor='green', alpha=0.4, label='write')
     ax.legend(handles=[r_read, r_inc, r_write], loc='upper center',
               bbox_to_anchor=(0.5, 1.07), ncol=3, fancybox=True,
               fontsize='x-small', framealpha=1)
-    ax.set_yticks(range(len(labels)))
-    ax.set_yticklabels([])
-    #ax.set_xlim([0,100])
+    #ax.set_yticks(range(len(labels)))
+    #ax.set_yticklabels([])
+    ax.set_yticks([], [])
+    ax.set_ylabel("thread")
+
+    if not spark:
[…]

### 2019-12-10 gen_figures.sh
fixes to figure generation script

@@ -4,14 +4,14 @@ curr_dir=`pwd`
 cd ./experiments/bids-app-example
 ./plot.py
 cd $curr_dir/experiments/bigbrain-incrementation/figures
-./makespan.py '../results/*25cpus*' 25cpus-120delay-1it
-./stacked_bar.py '../results/*25cpus*' 25cpus-120delay-1it
-./makespan.py '../results/*96cpus*' 96cpus-0delay-1it
-./stacked_bar.py '../results/*96cpus*' 96cpus-0delay-1it
-./makespan.py '../results/20bb_experiments/*25cpus*' 20bb_25cpus spark
-./stacked_bar.py '../results/20bb_experiments/*25cpus*' 20bb_25cpus spark
-./makespan.py '../results/20bb_experiments/*96cpus*' 20bb_96cpus spark
-./stacked_bar.py '../results/20bb_experiments/*96cpus*' 20bb_96cpus spark
+./makespan.py '../results/40bb_red0/*25cpus*' 25cpus-120delay-1it
+./stacked_bar.py '../results/40bb_redo/*25cpus*' 25cpus-120delay-1it
+./makespan.py '../results/40bb_redo/*96cpus*' 96cpus-0delay-1it
+./stacked_bar.py '../results/40bb_redo/*96cpus*' 96cpus-0delay-1it
+./makespan.py '../results/20bb_redo/*25cpus*' 20bb_25cpus spark
+./stacked_bar.py '../results/20bb_redo/*25cpus*' 20bb_25cpus spark
+./makespan.py '../results/20bb_redo/*96cpus*' 20bb_96cpus spark
+./stacked_bar.py '../results/20bb_redo/*96cpus*' 20bb_96cpus spark
 cd $curr_dir
 echo 'GNU Parallel: 75BB local App Direct 25 cpus - 1567704324-localAD_1it_25cpus_120delay_withpc-2.out:'
 ./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567704324-localAD_1it_25cpus_120delay_withpc-2.out --cpus 25

### 2019-12-16 gen_figures.sh
fixed figure generation script

@@ -4,26 +4,26 @@ curr_dir=`pwd`
 cd ./experiments/bids-app-example
 ./plot.py
 cd $curr_dir/experiments/bigbrain-incrementation/figures
-./makespan.py '../results/40bb_red0/*25cpus*' 25cpus-120delay-1it
-./stacked_bar.py '../results/40bb_redo/*25cpus*' 25cpus-120delay-1it
-./makespan.py '../results/40bb_redo/*96cpus*' 96cpus-0delay-1it
-./stacked_bar.py '../results/40bb_redo/*96cpus*' 96cpus-0delay-1it
+./makespan.py '../results/40bb_redo/*25cpus*' 40bb_25cpus
+./stacked_bar.py '../results/40bb_redo/*25cpus*' 40bb_25cpus
+./makespan.py '../results/40bb_redo/*96cpus*' 40bb_96cpus
+./stacked_bar.py '../results/40bb_redo/*96cpus*' 40bb_96cpus
 ./makespan.py '../results/20bb_redo/*25cpus*' 20bb_25cpus spark
 ./stacked_bar.py '../results/20bb_redo/*25cpus*' 20bb_25cpus spark
 ./makespan.py '../results/20bb_redo/*96cpus*' 20bb_96cpus spark
 ./stacked_bar.py '../results/20bb_redo/*96cpus*' 20bb_96cpus spark
 cd $curr_dir
-echo 'GNU Parallel: 75BB local App Direct 25 cpus - 1567704324-localAD_1it_25cpus_120delay_withpc-2.out:'
-./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567704324-localAD_1it_25cpus_120delay_withpc-2.out --cpus 25
-echo 'GNU Parallel: 75BB Optane App Direct 25 cpus - 1567705458-optaneAD_1it_25cpus_120delay_withpc-2.out:'
-./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567705458-optaneAD_1it_25cpus_120delay_withpc-2.out --cpus 25
-echo 'GNU Parallel: 75BB Isilon App Direct 25 cpus - 1567707833-isilonAD_1it_25cpus_120delay_withpc-2.out:'
-./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/1567707833-isilonAD_1it_25cpus_120delay_withpc-2.out --cpus 25
[…]

### 2019-12-17 paper.tex
requested changes

@@
 \usepackage{multirow}
+\usepackage{textcomp}
 \usepackage[binary-units]{siunitx}
@@
 
-\title{Performance benefits of Intel Optane DC storage for the parallel processing of large neuroimaging data}
+\title{Performance benefits of Intel\textsuperscript{\textregistered}~Optane\texttrademark~DC persistent memory for the parallel processing of large neuroimaging data}
 
@@
 \begin{abstract}
-    Open-access neuroimaging datasets have reached up to $\SI{}{\peta\byte}$ of data, and
-    continue to grow. The ability to leverage the entirety of these datasets are
-    limited to a restricted number of labs with both the capacity and infrastructure
-    to process the data. Whereas BigData engines have significantly reduced
-    application performance penalties with respect to data movement, their applied
-    strategies are not necessarily practical within neuroimaging workflows where
-    intermediary results may need to be materialized to shared storage for post-processing
-    analysis. In this paper we evaluate the performance advantage brought by Intel
-    Optane DC persistent memory storage for the processing of large neuroimaging 
-    datasets using the two available configurations modes: Memory Mode and App Direct Mode.
-    We employ a synthetic algorithm on the $\SI{76}{\gibi\byte}$ \bigbrain
-    and a real neuroimaging application on the Consortium for Reliability and Reproducibility (CoRR)
-    dataset using 25 and 96 parallel
-    processes in both cases. Our results show that performance of applications executing on
-    Optane is superior to that of other storage devices, with the exception of DRAM. This is
-    the case in both Memory and App Direct Mode and irrespective of the amount of data and 
-    parallelism. Furthermore, Optane in App Direct Mode is believed to benefit from the use
-    of memory as a cache for writing in cases where output data is significantly smaller than
-    available memory. We believe the use of Optane will be beneficial to both neuroimaging applications
-    running on HPC or visualization of large, high-resolution images.
+    Open-access neuroimaging datasets have reached petabyte scale, and continue to grow.
+The ability to leverage the entirety of these datasets is limited to a restricted
+number of labs with both the capacity and infrastructure to
+process the data. Whereas Big Data engines have significantly
+reduced application performance penalties with respect to data
+movement, their applied strategies are not necessarily practical
+within neuroimaging workflows where intermediary results may
+need to be materialized to shared storage for post-processing
+analysis. In this paper we evaluate the performance advantage
+brought by Intel\textsuperscript{textregistered}~Optane\texttrademark~DC persistent memory for the
+processing of large neuroimaging datasets using the two available
+configurations modes: Memory Mode and App Direct Mode. We
+employ a synthetic algorithm on the 76 GiB and 603 GiB BigBrain, as well as apply a
+standard neuroimaging application on the Consortium for Reliability
+and Reproducibility (CoRR) dataset using 25 and 96 parallel
+processes in both cases. Our results show that the performance of
+the applications leveraging persistent memory is superior to that of other
+storage devices, with the exception of DRAM. This is the case
+in both Memory and App Direct Mode and irrespective of the
+amount of data and parallelism. Furthermore, persistent memory in App
+Direct Mode is believed to benefit from the use of memory as
+a cache for writing in cases where output data is significantly
+smaller than available memory. We believe the use of persistent memory
+will be beneficial to both neuroimaging applications running on
+HPC or visualization of large, high-resolution images.
 \end{abstract}
@@
 studies in neuroimaging remain limited to labs with access to adequate 
-infrastructure.
+infrastructures.
 
@@
 To mitigate the effects of data writes, kernel-based
-strategies,such as the writeback cache, have been developed. The writeback cache
+strategies, such as the writeback cache, have been developed. The writeback cache
 allows processes to use the memory as a cache for writing. The cache size,
@@
 is configurable, but nevertheless limited. When writes approach the cache's capacity,
-processes performing writes start to be throttled. After capacity is reached,
-processes can no longer use the cache until all written data on cache is flushed
+processes performing writes start to be throttled. Once the cache capacity is reached,
+processes can no longer use the cache until all cached written data is flushed
 to the appropriate storage device.
@@
 data locality, and lazy evaluation), hardware has also adapted to the growing datasets.
-One such improvement is the concept of placing flash storage directly on the Dual
+One such improvement is the concept of placing persistent storage directly on the Dual
 In-line Memory Module (DIMM), thereby reducing the latency of accessing data on 
@@
 
-Intel Optane DC Persistent Memory storage~\cite{optanebrief} is a recently introduced high-performance
-storage technology that also sits on the DIMMs to reduce latency to the device.
-Unlike DRAM DIMMs, it can accommodate twice the volume ($\SI{512}{\giga\byte}$) of
-storage enabling it to be a more cost effective alternative to DRAM storage.
-There are two principle configuration modes, namely Memory Mode and App Direct Mode,
-that enable the storage to either be accessed as an extension of available memory 
-or as a non-volatile memory storage device.
+Intel Optane DC Persistent Memory~\cite{optanebrief} is a high-performance
+storage technology that resides on the DIMMs to reduce latency to the device.
+The first generation of Intel Optane DC persistent memory offers capabities of 128~GB,
+256~GB, and 512~GB, enabling it to be more cost effective than DRAM storage.
+There are two configuration modes, Memory Mode and App Direct Mode,
+that enable the storage to either be accessed as an extension of volatile main memory
+or as a non-volatile memory storage device\footnote{https://itpeernetwork.intel.com/intel-optane-dc-persistent-memory-operating-modes/}.
 
@@
 \begin{itemize}
-        \item Quantify the added value of Intel Optane DC Persistent Memory on 
+        \item Quantify the added value of Intel Optane DC persistent memory on 
             processing large neuroimaging data using representative pipelines; and
-        \item Determine when a given Intel Optane storage configuration (Memory 
+        \item Determine when a given Intel Optane DC persistent memory configuration (Memory 
             and App Direct Mode) is preferable.
@@
 
-The server used for the experiments consisted of 12 \SI{64}{\gibi\byte} DRAM devices,
-resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12 \SI{256}{\giga\byte} Intel Optane DC Memory
-devices, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC Memory. The theoretical maximum bandwidth
-of DRAM is estimated to be around \SI{275}{\gibi\byte}~/s whereas for Optane, it is
-estimated to be around \SI{250}{\gibi\byte}~/s.
+The server used for the experiments consisted of 12X~\SI{64}{\gibi\byte} DRAM devices,
+resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\giga\byte} Intel Optane DC persistent memory
+modules, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC persistent memory.
 Other storage devices included a \SI{240}{\gibi\byte} Micron SATA SSD, of which
@@
 cluster (10Gb network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
-was the only storage device set up as a writeback device. Both Isilon and Optane in
+was the only storage device set up as a writeback device. Both Isilon and Optane DC persistent memory in
 App Direct mode were configured as write-through devices, meaning they did not leverage
-DRAM as a cache for writes. This was achieved in Optane by configuring the xfs filesystem with
+DRAM as a cache for writes. For Optane DC persistent memory, this was achieved by configuring the xfs filesystem with
 DAX. All storage devices were benchmarked using the following \href{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}{script}.
@@
  \hline
-\end{tabular}\caption{Perceived read and write bandwidth of each storage device}\label{table:bandwidths}
+\end{tabular}\caption{Measured read and write bandwidth of each storage device}\label{table:bandwidths}
 \end{center}
@@
 
-Memory Mode leverages Intel Optane DC persistent memory (Optane) to extend the system's available
-memory. In this mode, Optane uses DRAM as a cache and is accessible as
-volatile addressable main memory. By extending main memory, Memory Mode enables the
-fast access of large volumes of data.
+Memory mode leverages Intel Optane DC persistent memory (DCPMM) to extend the system's available
+memory. In this mode, Optane DCPMM uses DRAM as a cache and is accessible as
+volatile addressable main memory. By extending main memory, Memory mode enables the
+fast access of large volumes of data. For instance, Memory mode on our server enabled use of
+3~TiB of main memory, whereas it would have only had 768~GiB of main memory if Optane DCPMM could not be
+leveraged. Although DRAM is used as cache by Optane DCPMM, it is not visible to the operating system.
 
-While enabling the access to larger amount of memory than typically accessible otherwise,
-it is anticipated that Memory Mode will be slower than AD mode for all memory accessed that
-are not cached in DRAM. We evaluated all available devices (Optane through tmpfs, local disk and Isilon)
-in Memory Mode.
+While enabling access to a larger amount of memory than typically accessible otherwise,
+it is anticipated that Memory mode will be slower than App Direct mode for all memory accessed that
+are not cached in DRAM. We evaluated all available devices (Optane DCPMM through tmpfs, local disk and Isilon)
+in Memory mode.
 
@@
 
-App Direct Mode enables Optane to be accessed as a high-performance storage device.
-Unlike Memory Mode, the OS is able to differentiate between DRAM and Optane,
-treating them as two distinct entities. As Optane is located on
-the DIMMs, it exhibits lower latencies than traditional SSDs, while being able
-to be accessed like an SSD. Optane does not use the DRAM as cache in App Direct Mode
-if configured with DAX.
+App Direct mode enables Optane DCPMM to be accessed as a high-performance storage device.
+Unlike Memory mode, the OS is able to differentiate between DRAM and Optane DCPMM,
+treating them as two distinct memory tiers. Optane DCPMM
+does not use the DRAM as cache in App Direct Mode if configured with DAX.
 
 Similarly to Memory Mode, our experiments evaluated all available filesystems in App Direct
-Mode. In other words, we evaluated DRAM performance through tmpfs, Optane performance through
-a \SI{1.5}{\tebi\byte} Intel Optane DC Persistent Memory storage device and local disk
-and Isilon using the same storage mounts available in Memory Mode.
+Mode. In other words, we evaluated DRAM (tmpfs mount), Optane DCPMM (\SI{1.5}{\tebi\byte} persistent
+memory mount), local SSD and Isilon.
 
@@
 to occur with storage devices that scale well with increased parallelism, as should be the
-case for DRAM, Optane and Isilon, which all consist of multiple devices. Should the
+case for DRAM, Optane DCPMM and Isilon, which all consist of multiple devices. Should the
 scalable storage devices predict the makespan accurately, it is believed that other overheads
@@
         \Desc{$C$}{a set of image chunks}
-        \Desc{$fs$}{filesystem to write to (tmpfs, Optane, local disk, Isilon)}
+        \Desc{$fs$}{filesystem to write to (tmpfs, Optane DCPMM, local disk, Isilon)}
     \EndInput
@@
 in-memory, by ensuring that read and written data are not the same. This application
-was parallelized us in two different ways: using both GNU Parallel and Apache Spark 2.4.3 (PySpark). 
+was parallelized in two different ways: using both GNU Parallel and Apache Spark 2.4.3 (PySpark). 
 All code was implemented in Python 3.6.
@@
 For instance, in the GNU Parallel implementation, all operations (i.e. reading, incrementing and writing)
-are done within the same task. The same task is applied to all the \bigbrain blocks and GNU Parallel is 
-provided with all the different command-line calls, processing a subset of them at a time,
+are done within a single task. This task is applied to all the \bigbrain blocks using GNU Parallel,
+processing a subset of them at a time,
 depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
-is achieved using the popular neuroimaging I/O: Nibabel. When provided with a filename, as is the case
-for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memmap the data
-using Numpy. This step will be referred to as ``load header''. It is only when the data is actually required
+is achieved using the popular neuroimaging I/O: \href{https://nipy.org/nibabel/}{NiBabel}. When provided with a filename, as is the case
+for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memory map the data
+using \href{https://numpy.org/}{NumPy}. This step will be referred to as ``load header''. It is only when the data is actually required
 (i.e. during incrementation), that the data will be loaded into memory.
@@
 The Apache Spark implementation differs from that of GNU Parallel in that scheduling decisions
-were entirely differed to Spark. That is, rather than simply parallelizing the filenames, to be
-loaded in a similar fashion to that of GNU Parallel, we chose to load the using Spark's built-in
-\texttt{BinaryFiles}, which loads whole files, in binary format, into Spark partitions. As the
+were entirely differed to Spark. That is, rather than simply parallelizing over the \bigbrain block filenames,
+we opted to read the data using Spark's built-in
+\texttt{BinaryFiles}, loading whole files, in binary format, into Spark partitions. As the
 data would have been pre-loaded by Spark, ``loading header'' only measures the time to
-convert the binary images into Nibabel objects, and ``incrementation'' consists purely
+convert the binary images into NiBabel objects, and ``incrementation'' consists solely
 of the time it took to increment the data. Therefore, only the write time would be
@@
 using 25 and 96 processes on the 125 40$\mu$m \bigbrain blocks. Data was read and written to either
-DRAM (App Direct only), Optane, local disk and Isilon.
+DRAM (App Direct only), Optane DCPMM, local disk and Isilon.
 
@@
 experiments using both 25 and 96 cpus. Repeating the experiments using a larger dataset would enable
-us to determine the effects when Optane would have to be partially relied upon due to insufficient DRAM
+us to determine the effects when Optane DCPMM would have to be partially relied upon due to insufficient DRAM
 space. Moreover, with such large datasets, it is likely that Big Data frameworks would be used rather than
@@
 Since storage was limited on DRAM and local disk and was not large enough to process the entire
-20~$\mu$m \bigbrain, these devices were omitted in the processed of this dataset.
+20~$\mu$m \bigbrain, these devices were omitted in the processing of this dataset.
 
@@
 on all the anatomical (T1W) images of dataset containing numerous subjects. This step is 
-referred to as Participant analysis within the BIDS App Example. An optional
+referred to as Participant analysis within the application. An optional
 step of the BIDS App example, referred to as Group analysis, computes the average brain
@@
 applied Participant analysis to it. The BIDS App Example was executed using a Singularity container stored on
-Isilon. As in the BigBrain Incrementation, the experiment parallelized 
-using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode
-using Isilon, local disk and Optane as storage. Each experiment was repeated 3x.
+Isilon. As in the BigBrain Incrementation, the experiment was parallelized 
+using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode,
+where Isilon and the local SSD were only evaluated in Memory Mode. Each experiment was repeated 3x.
 
@@
     \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382482-optaneAD_1it_25cpus_40bb-1.pdf}
-    \caption{Optane}
+    \caption{Optane DCPMM}
 \end{subfigure}
@@
 
-As can be seen in Figure~\ref{fig:makespan-25cpus}, DRAM (App Direct Mode) is the 
+As can be seen in Figure~\ref{fig:makespan-25cpus}, DRAM (App Direct mode) is the 
 most efficient, with a makespan of 7.6s. Optane in Memory mode is a close second
-with a makespan of approximately 10s. Optane in App Direct mode is approximately 
-12x slower than Optane in Memory Mode. Isilon in both Memory Mode and App Direct 
-Mode is more than 120x slower than Optane in Memory Mode. Other than for Optane, 
-there is no significant difference between Memory Mode and App Direct executions.
+with a makespan of approximately 10s. As caching was disabled in App Direct mode, 
+Optane DCPMM in this mode is approximately 
+12x slower than Optane in Memory Mode. Isilon in both Memory mode and App Direct 
+mode is more than 120x slower than Optane DCPMM in Memory mode. Other than for Optane DCPMM, 
+there is no significant difference between Memory mode and App Direct executions.
 
 When comparing the makespan to the estimated read and write duration, local SSD
-in both modes and Optane in App Direct Mode have longer makespans than the anticipated.
-Optane was approximately 2.2x slower than expected, whereas local SSD was approximately
+in both modes and Optane DCPMM in App Direct mode have longer makespans than anticipated.
+Optane DCPMM was approximately 2.2x slower than expected, whereas local SSD was approximately
 1.4x slower than expected, in both modes. For those where the I/O time was overestimated,
-DRAM was almost 5x faster in reality compared to what was estimated. Optane in Memory Mode
-had a smaller difference, with it being 3.8x faster than expected. Isilon in App Direct
-mode had a larger difference than Memory mode, with it being 1.1x faster than expected.
+DRAM was almost 5x faster in reality compared to what was estimated. Optane DCPMM in Memory Mode
+had a larger difference, with it being 3.8x faster than expected. Isilon in App Direct
+mode had a smaller difference than Memory mode, with it being 1.1x faster than expected.
 
@@
 shows that, as expected, I/O times vary greatly between devices, with DRAM having the
-best total read and write speeds. Optane in Memory Mode is very close to DRAM in speed,
-whereas Optane in App Direct mode 7x slower in terms of reading and 23x slower in terms
+best total read and write speeds. Optane DCPMM in Memory Mode is very close to DRAM in speed,
+whereas Optane DCPMM in App Direct mode was 7x slower in terms of reading and 23x slower in terms
 of writing. Local SSD, on the other hand, exhibits similar write times as DRAM, but is
-approximately 96x that DRAM with respect to reads. Header loading for local SSD is also 90x longer
-than that of DRAM, however, remains negligible compared to other tasks. Isilon, on the 
-other hand, has non negligible header load times, with it being more than 5000 times
+approximately 96x slower than DRAM with respect to reads. Header loading for local SSD is also 90x longer
+than that of DRAM, however, it remains negligible compared to other tasks. Isilon, on the 
+other hand, has non negligible header load times, with it being more than 5000x
 slower than that of DRAM. Interestingly enough, the read/increment times on Isilon are
-slower than that of the local SSD, although Isilon makes up for its speedier reads with 
+faster than that of the local SSD, although Isilon makes up for its speedier reads with 
 writes that are around 184x slower than DRAM.
@@
 DRAM is significantly faster than all other storage, leading it to barely appear within the Gantt chart.
-DRAM was measured to have an average of 17.2 parallel tasks throughout the execution. Optane takes
+DRAM was measured to have an average of 17.2 parallel tasks throughout the execution. Optane DCPMM takes
 significantly longer than DRAM and appears to spend a significant amount of time writing. The average
-parallelism measured for Optane was 24. Unlike Optane, local SDD spends the vast majority of its time
-reading, while spending a negligible amount of time writing. Like Optane, it averages 24 parallel tasks
-throughout its execution. Unlike the other three storages, Isilon shows spacing between the loading of the header,
+parallelism measured for Optane DCPMM was 24. Unlike Optane DCPMM, local SSD spends the vast majority of its time
+reading, while spending a negligible amount of time writing. Like Optane DCPMM, it averages 24 parallel tasks
+throughout its execution. Unlike the other three storage, Isilon shows spacing between the loading of the header,
 read and increment and write. Isilon was measured to have an average parallelism of 14. Overall, each storage device
@@
     \captionsetup{belowskip=-10pt}
-\caption{Gantt charts for Optane processing 125 blocks of the 40$\mu$m BigBrain using 96 processes}\label{fig:gantt96}
+\caption{Gantt charts for Optane DCPMM processing 125 blocks of the 40$\mu$m BigBrain using 96 processes}\label{fig:gantt96}
 \end{figure*}
@@
 Despite having increased parallelism by a factor of 3.84, we see no visible reduction
-in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 cpus, DRAM and Optane in Memory Mode are the only storage
-devices that did better than the anticipated I/O, with a makespan duration of
-5.7 and 9s, respectively (total read and write estimate for both was measured to be around 39s). While
-Optane in App Direct Mode and local SSD both performed worse than the anticipated I/O estimates, Isilon
+in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 cpus, DRAM and Optane DCPMM in Memory mode are the only devices that did better than the anticipated I/O, with a makespan duration of
+5.7s and 9s, respectively (total read and write estimate for both was measured to be around 39s). While
+Optane DCPMM in App Direct mode and local SSD both performed worse than the anticipated I/O estimates, Isilon
 performed as expected. Variance was found to be high on the local SSD.
@@
 Isilon only appeared to display a performance decrease with respect to writes. Furthermore, while both
-read and writes decreased for Optane in App Direct Mode, read duration was nearly 6x slower with more threads,
+read and writes decreased for Optane DCPMM in App Direct mode, read duration was nearly 6x slower with more threads,
 while write duration was only about 2x slower. For local SSD and Isilon, there does not appear to 
@@
 
-When analyzing the average parallelism during the workload, only Optane in App Direct mode
-came close to 96 parallel tasks, having an average parallelism of 88 tasks. Optane in Memory mode,
+When analyzing the average parallelism during the workload, only Optane DCPMM in App Direct mode
+came close to 96 parallel tasks, having an average parallelism of 88 tasks. Optane DCPMM in Memory mode,
 DRAM and local SSD all averaged between 61-66 tasks. Isilon performed the worst with 38 tasks in Memory 
@@
 
-A further look at the Gantt charts between Optane in Memory Mode and App Direct Mode (Figure~\ref{fig:gantt96})
+A further look at the Gantt charts between Optane DCPMM in Memory mode and App Direct mode (Figure~\ref{fig:gantt96})
 show that both read and write time of tasks are worse in App Direct mode as compare with Memory mode.
-While some reads tasks in App Direct mode appear to be of similar duration to the average read task time
-of Memory mode, the vast majority of read tasks are significantly slower. Futhermore, not write tasks in App
+While some read tasks in App Direct mode appear to be of similar duration to the average read task time
+of Memory mode, the vast majority of read tasks are significantly slower. Furthermore, no write tasks in App
 Direct mode is capable of reaching the performance of equivalent tasks in Memory mode.
@@
 The anticipated I/O estimates did not correctly predict the makespan for 25 parallel
-processes (Figure~\ref{fig:20mksp25}). Both Optane in Memory Mode and Isilon appear 
-to be under the estimates: Optane in Memory Mode is approximately twice as fast as the
+processes (Figure~\ref{fig:20mksp25}). Both Optane DCPMM in Memory Mode and Isilon appear 
+to be under the estimates: Optane DCPMM in Memory Mode is approximately twice as fast as the
 estimates whereas Isilon is around 1.1x faster. Optane was 1.8x slower than anticipated.
@@
 
-Total task duration breakdowns (Figure~\ref{fig:20total25}) show that for Optane in
+Total task duration breakdowns (Figure~\ref{fig:20total25}) show that Optane DCPMM in
 App Direct Mode spends significantly more time writing than its Memory Mode counterpart.
@@
 between Memory and App Direct Mode, spending around 1.4x more time reading and incrementing
-in App Direct Mode. While the read and write duration appeared to be more-or-less balanced in
-Isilon (as was anticipated), Optane in App Direct Mode spent significantly more time writing, whereas the 
-inverse was apparent in Memory Mode
+in App Direct Mode.
 
 Unlike the GNU Parallel executions, the Spark executions appear to be in better accordance with
-the I/O estimates (Figure~\ref{fig:20mkspspark25}. Only Optane in App Direct Mode appears
+the I/O estimates (Figure~\ref{fig:20mkspspark25}. Only Optane DCPMM in App Direct mode appears
 to have significantly exceeded the estimates, by a factor of 2. Total write duration all appear
 to be longer in the Spark implementation than with GNU Parallel (Figure~\ref{fig:20totalspark:25}).
-Spark's writes are 1.4-1.6x longer with Optane and 1.1x longer with Isilon. Due to differences in
+Spark's writes are 1.4-1.6x longer with Optane DCPMM and 1.1x longer with Isilon. Due to differences in
 how data is loaded in Spark, it is not possible to comment on how read times are affected. However, it
-is possible to note that data conversion (binary string to Numpy array) takes almost twice as 
-much time in Optane than it does with Isilon. Furthermore, the act of incrementation can take
-up to twice as much time with Isilon.
+is possible to note that data conversion (binary string to NumPy array) takes almost twice as 
+much time on Optane DCPMM than it does on Isilon. Furthermore, the act of incrementation can take
+up to twice as much time on Isilon.
 
@@
 
-A very slight performance improvement (1-1.3x faster) can be witnessed in Optane. In 
+A very slight performance improvement (1-1.3x faster) can be witnessed in Optane DCPMM. In 
 contrast, Isilon 
 experienced a slight performance decrease (1.1x slower). Whereas Isilon makespan appears to match
-estimated I/O times, Optane once again has a superior Memory Mode makespan and an inferior App Direct
-Mode makespan (Figure~\ref{fig:20mksp96}). As previously observed with 25 processes,
-the bulk of the processing time is taken up by writing for Optane in App Direct Mode and 
-Isilon in both modes (Figure~\ref{fig:20total96}). Reading takes more time in the case of Optane in Memory mode.
+estimated I/O times, Optane DCPMM once again has a superior Memory mode makespan and an inferior App Direct
+mode makespan (Figure~\ref{fig:20mksp96}). As previously observed with 25 processes,
+the bulk of the processing time is taken up by writing for Optane DCPMM in App Direct Mode and 
+Isilon in both modes (Figure~\ref{fig:20total96}). Reading takes more time in the case of Optane DCPMM 
+in Memory mode.
 
-A notable difference in the Spark execution is that Optane in Memory and App Direct Mode
-exhibit the same performance (Figure~\ref{20mkspnspark96}). Both of which also performed
+A notable difference in the Spark execution is that Optane DCPMM in Memory and App Direct mode
+exhibit the same performance (Figure~\ref{20mkspspark96}). Both of which also performed
 significantly worse than the estimated I/O duration. Performance on Isilon, however, did
 not really differ. The total task breakdown shows that, as expected, write times are more
-significant on Isilon than on Optane and Optane in App Direct Mode spends more time writing
-than in Memory Mode (Figure~\ref{20fig:20totalspark96}). When comparing GNU Parallel and 
+significant on Isilon than on Optane DCPMM and Optane DCPMM in App Direct mode spends more time writing
+than in Memory mode (Figure~\ref{fig:20totalspark96}). When comparing GNU Parallel and 
 Spark implementations, Spark spent nearly half as much time writing as GNU Parallel. However,
-it was found that Optane spent more time converting the data to Nibabel than GNU Parallel spent
+it was found that Optane DCPMM spent more time converting the data to NiBabel than GNU Parallel spent
 in loading the header.
@@
 Unlike the BigBrain incrementation, the makespan of the BIDS App Example
-executing on Optane is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
-to the BigBrain Incrementation, Optane in Memory Mode performs better than the other storage devices
+executing on Optane DCPMM is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
+to the BigBrain Incrementation, Optane DCPMM in Memory mode performs better than other storage devices
 while Isilon has the longest makespan. The same pattern can be found with 96 process (Figure~\ref{fig:bm96}). When comparing
@@
 given moment did not exceed the amount of available DRAM. Although GNU Parallel would have
-been able to load all the blocks in memory, had we increased the block size, it would have been unlikely
-that total available DRAM would have been exceeded, though it would have occurred in Memory Mode
-for the devices that leverage memory as a write cache. In our case, only the local SSD was set up to do so.
+been able to load all the blocks in memory had we increased the block size, it would have been unlikely
+that total available DRAM would have been exceeded. However, available DRAM could have been exceeded in
+Memory Mode when using devices that leverage memory as a write cache.
+In our case, only the local SSD was set up to do so.
 
-Optane was the only device affected by the choice of storage mode. In all cases, with the 
-exception of the Spark incrementation of the 20~$\mu$ \bigbrain, Memory Mode was superior to 
-App Direct Mode. This is a result of Optane using DRAM as cache in Memory Mode, whereas Optane
-was configured as a writethrough device in App Direct Mode. It is believed that had Optane been
-configured to use DRAM as a writeback cache in App Direct Mode, we  would have observed a similar
-performance between the two. However, for the 40~$\mu$ \bigbrain executions, read time on Optane in
-App Direct Mode appeared to be slower than Memory Mode, potentially indicating that the Optane was able
-to keep the input data cached in DRAM whereas in App Direct Mode, the data had to be loaded from Optane.
-Therefore, it is expected the Optane in App Direct mode may always have longer read times than Memory Mode
+Optane DCPMM was the only device affected by the choice of storage mode. In all cases, with the 
+exception of the Spark incrementation of the 20~$\mu$ \bigbrain, Memory mode was superior to 
+App Direct mode. This is a result of Optane DCPMM using DRAM as cache in Memory mode, whereas Optane
+DCPMM 
+was configured as a writethrough device in App Direct mode. It is believed that had Optane DCPMM been
+configured to use DRAM as a writeback cache in App Direct mode, we would have observed a similar
+performance between the two. However, for the 40~$\mu$ \bigbrain executions, read time on Optane DCPMM in
+App Direct mode appeared to be slower than Memory mode, potentially indicating that the Optane DCPMM was able
+to keep the input data cached in DRAM, whereas in App Direct Mode, the data had to be loaded from Optane DCPMM.
+Therefore, it is expected the Optane DCPMM in App Direct mode may always have longer read times than Memory Mode
 if the data is already cached in DRAM in Memory Mode.
@@
 The Apache Spark and GNU Parallel implementations differed vastly in makespan 
-for Optane in Memory Mode, particularly at 96 processes. After taking a closer look at
-the captured Spark metrics, it was for that what really varied between the Spark runs on
-the different storage was garbage collection, with garbage collection taking longer on Optane 
-than on Isilon. Furthermore, garbage collection was longer in Optane Memory Mode than in 
-Optane App Direct Mode. It is suspected that, in these cases, task duration was so short
-that garbage collection could not keep up. Moreover, Optane in Memory Mode may have longer
-garbage collection if old anonymous data cannot be located in the DRAM cache due to overall memory 
+for Optane DCPMM in Memory Mode, particularly at 96 processes. After taking a closer look at
+the captured Spark metrics, what significantly varied between the Spark runs on
+the different storage was garbage collection, with garbage collection taking longer on Optane DCPMM 
+than on Isilon. Furthermore, garbage collection was longer in Memory mode than in
+App Direct mode. It is suspected that, in these cases, task duration was so short
+that garbage collection could not keep up. Moreover, Optane DCPMM in Memory mode may have longer
+garbage collection if old anonymous data cannot be stored in the DRAM cache due to overall memory 
 requirements, therefore requiring garbage collection to occur on slower memory.
@@
 
-The use of Optane in Memory Mode would enable more dirty data to be written to available memory
+The use of Optane DCPMM in Memory mode would enable more data to be written to available memory
 rather than directly to slower storage devices. In our 40~$\mu$ \bigbrain experiments, it was found
-that read times were less important in Isilon than the local SSD (Figures~\ref{fig:stacked-25cpus}
+that read times were less important on Isilon than on the local SSD (Figures~\ref{fig:stacked-25cpus}
 and~\ref{fig:stacked-96cpus}). Should Isilon have been given the opportunity to benefit from a 
 writeback cache like the SSD, it is believed that performance on Isilon could have been superior 
-to that of the SSD. Having Optane as a node-local burst buffer to shared network storage 
+to that of the SSD. Having Optane DCPMM as a node-local burst buffer to shared network storage 
 may prove to be very beneficial in the case of Big Data neuroimaging pipelines running on
@@
 
-For App Direct Mode applications (although also applicable to Memory Mode), the 
+For App Direct mode applications (although also applicable to Memory mode), the 
 choice of using the writeback cache depends on total amount of available memory, application
@@
 be seen in Figure~\ref{fig:stacked-25cpus}, for instance, where the local SSD is writing
-entirely to memory, reducing significantly the total write time. However, should write operations
-become throttled, this may significantly slow down the write task duration. In such cases,
+entirely to memory, significantly reducing the total write time. However, should write operations
+become throttled, this may slow down the write task duration. In such cases,
 it may be preferable to set them up as write-through devices. As Big Data neuroimaging applications
 tend to generate temporary data files that are larger than input size, it is likely that maintaining the
-devices as write-through is preferable. Furthermore, despite that Optane in App Direct Mode does
+devices as write-through is preferable. Furthermore, despite the fact that Optane DCPMM in App Direct Mode does
 not leverage the writeback cache like the local SSD, is it still found to be superior in performance.
-Therefore, the while the performance of Optane may not be enhanced by DRAM, it is still superior to that of
-SSDs using the writeback cache and scalable network storage.
+Therefore, the while the performance of Optane DCPMM may not be enhanced by DRAM, it is still superior to that of
+an SSD using the writeback cache and scalable network storage.
 
@@
 each storage device. For instance, DRAM consisted of 12 devices, whereas although there were 
-12 Optane devices, only 6 were accessible by the mountpoint selected in App Direct Mode. Furthermore,
-local SSD only consisted of a single device, whereas the Optane server was made up of 180 SSDs. The differences
+12 Optane devices, only 6 were accessible by the mountpoint selected in App Direct mode. Furthermore,
+local SSD only consisted of a single device, whereas the Isilon server was made up of 180 SSDs. The differences
 in number of storage devices would account for the attainable amount of parallel writes. For instance, 
-Isilon would have been the most scalable, followed by DRAM, Optane and the local SSD, which would
+Isilon would have been the most scalable, followed by DRAM, Optane DCPMM and then the local SSD, which would
 only be able to process I/O sequentially. This explains why the anticipated I/O time, is sometimes greater
@@
 around 37.5s. At 25 processes, we found the DRAM makespan to be around 7s, which is much closer to the 3s
-estimate than the 37.5s. Reasons for why it took longer that estimated could be due to application
-overheads, or even that some DRAM devices were occupied, and therefore, the maximum amount of parallelism 
+estimate than the 37.5s. Reasons for why it took longer than estimated could be due to the fact that
+our applications were not optimized for non-uniform memory accesses (NUMA-aware), or even that some DRAM devices were occupied, and therefore, the maximum amount of parallelism 
 was not achievable.
@@
 Despite the fact that only sequential I/O was considered for the anticipated I/O times, 
-Optane in App Direct Mode always took more time than its sequential estimates. The cause
+Optane DCPMM in App Direct mode always took more time than its sequential estimates. The cause
 of the longer-than-expected makespan seems to be due to write times. While there may be application overheads,
-they should, in theory, affect Optane in App Direct Mode as much DRAM. While it is unknown
-why Optane writes are performing so poorly, it could be a result of filesystem configuration or 
-inaccurate benchmarking of the disk. Otherwise, Optane does not appear to scale very well with respect to
+they should, in theory, affect Optane DCPMM in App Direct Mode as much DRAM. While it is unknown
+why Optane DCPMM writes are performing so poorly, it could be a result of filesystem configuration or 
+inaccurate benchmarking of the disk. Otherwise, Optane DCPMM does not appear to scale very well with respect to
 writes. Using the storage benchmarks found in Table~\ref{table:bandwidths}, it is found that with an equivalent
-number of storage devices, processing the 40~$\mu$ \bigbrain using 25 processes would be faster using
-local disk when compared to the real times obtained on Optane. As there are overheads with both Optane and local
+number of storage devices, processing the 40~$\mu$m \bigbrain using 25 processes would be faster using
+local disk when compared to the real times obtained on Optane DCPMM. As there are overheads with both Optane DCPMM and local
 SSD, it is uncertain that it would, in fact, be the case.
@@
 
-Our results show that Optane has superior performance to that of other storage devices, with the
-exception of DRAM. Optane as a persistent memory storage device is expected to bring significant
+Our results show that Optane DCPMM has superior performance to that of other storage devices.
+Optane DCPMM as a persistent memory storage device is expected to bring significant
 performance improvements to the processing of neuroimaging data. Input datasets to neuroimaging pipelines
-are increasing rapidly in size. Storing such datasets directly on Optane would significantly reduce the
-impacts of I/O on the processing. If Optane is located as a persistent memory storage device in HPC clusters,
+are rapidly increasing in size. Storing such datasets directly on Optane DCPMM would significantly reduce the
+impacts of I/O on the processing. If Optane DCPMM is located as a persistent memory
+storage device in HPC clusters,
 it could also be used as a burst buffer to network-attached storage devices. This would vastly
@@
 
-In the case of high-resolution images, Optane can be leveraged to enable the users to rapidly
-extract their regions of interest with minimal I/O costs. Futhermore, Optane would improve the
+In the case of high-resolution images, Optane DCPMM can be leveraged to enable the users to rapidly
+extract their regions of interest with minimal I/O costs. Futhermore, Optane DCPMM would improve the
 speed and fluidity of visualization applications on these datasets.
@@
 
-Optane has been found to drastically reduce the processing time of neuroimaging applications and
-can bring the performance close to DRAM speeds. Extending available memory using Optane is also 
+Optane DCPMM has been found to drastically reduce the processing time of neuroimaging applications and
+can bring the performance close to DRAM speeds. Extending available memory using Optane DCPMM is also 
 expected to help reduce I/O times of write-back devices by potentially extending available cache space.
-It is unclear, however, if traditional SSDs can perform just as well as Optane with additional storage devices
+It is unclear, however, if traditional SSDs can perform just as well as Optane DCPMM with additional storage devices
 attached.
 
-It is believe Optane can be useful for a variety of neuroimaging applications. For
+It is believe Optane DCPMM can be useful for a variety of neuroimaging applications. For
 instance, typically applications executing in HPC environments can benefit from 
-the speedups provided by Optane. Furthermore, image visualization servers will be able
+the speedups provided by Optane DCPMM. Furthermore, image visualization servers will be able
 to display higher resolution images at significantly greater speeds.

### 2019-12-17 paper.tex
tristan modifs

@@
 reduced application performance penalties with respect to data
-movement, their applied strategies are not necessarily practical
+movement, their applied strategies (e.g. data locality, in-memory computing and lazy evaluation)
+are not necessarily practical
 within neuroimaging workflows where intermediary results may
@@
 processing of large neuroimaging datasets using the two available
-configurations modes: Memory Mode and App Direct Mode. We
+configurations modes: Memory mode and App Direct mode. We
 employ a synthetic algorithm on the 76 GiB and 603 GiB BigBrain, as well as apply a
@@
 storage devices, with the exception of DRAM. This is the case
-in both Memory and App Direct Mode and irrespective of the
+in both Memory and App Direct mode and irrespective of the
 amount of data and parallelism. Furthermore, persistent memory in App
-Direct Mode is believed to benefit from the use of memory as
-a cache for writing in cases where output data is significantly
+Direct mode is believed to benefit from the use of DRAM as
+a cache for writing when output data is significantly
 smaller than available memory. We believe the use of persistent memory
@@
 the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
-1,200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
+1,200 healthy adults, expected to exceed $\SI{2}{\peta\byte}$ in size; 
 and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
@@
 in higher-capacity, slower storage devices, such as hard disk drives, or external
-parallel file systems, such as Lustre. In such conditions, large-scale 
+parallel file systems, such as Lustre~\cite{lustre}. In such conditions, large-scale 
 studies in neuroimaging remain limited to labs with access to adequate 
@@
 
-To mitigate the effects of data writes, kernel-based
-strategies, such as the writeback cache, have been developed. The writeback cache
-allows processes to use the memory as a cache for writing. The cache size,
-however,
+To mitigate the effects of data writes, the Linux kernel has implemented
+strategies, such as the writeback cache. The writeback cache
+allows processes to use the memory as a cache for writing. This strategy contrasts writethrough, which 
+writes data directly to the device. The writeback cache size
 is configurable, but nevertheless limited. When writes approach the cache's capacity,
@@
 
-Intel Optane DC Persistent Memory~\cite{optanebrief} is a high-performance
+Intel Optane DC Persistent Memory Module~\cite{optanebrief} (DCPMM) is a high-performance
 storage technology that resides on the DIMMs to reduce latency to the device.
-The first generation of Intel Optane DC persistent memory offers capabities of 128~GB,
+The first generation of Intel Optane DCPMM offers capabities of 128~GB,
 256~GB, and 512~GB, enabling it to be more cost effective than DRAM storage.
-There are two configuration modes, Memory Mode and App Direct Mode,
+There are two configuration modes, Memory mode and App Direct mode,
 that enable the storage to either be accessed as an extension of volatile main memory
@@
         \item Determine when a given Intel Optane DC persistent memory configuration (Memory 
-            and App Direct Mode) is preferable.
+            and App Direct mode) is preferable.
 \end{itemize}
@@
 
-The server used for the experiments consisted of 12X~\SI{64}{\gibi\byte} DRAM devices,
-resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\giga\byte} Intel Optane DC persistent memory
+The server used consisted of 12X~\SI{64}{\gibi\byte} DRAM devices,
+resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\giga\byte} Optane DCPMM
 modules, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC persistent memory.
@@
 only \SI{149}{\gibi\byte} where available for the experiments, and a
-\SI{720}{\tebi\byte} shared Dell EMC Isilon cluster. The Isilon cluster is an nsf4 mounted
+\SI{720}{\tebi\byte} shared Dell EMC Isilon network-attached storage platform. Isilon is an nsf4 mounted
 cluster (10Gb network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
-was the only storage device set up as a writeback device. Both Isilon and Optane DC persistent memory in
+was the only storage device set up as a writeback device. Both Isilon and Optane DCPMM in
 App Direct mode were configured as write-through devices, meaning they did not leverage
 DRAM as a cache for writes. For Optane DC persistent memory, this was achieved by configuring the xfs filesystem with
-DAX. All storage devices were benchmarked using the following \href{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}{script}.
+direct access (DAX).
+DAX enables persistent memory to be byte addressable, bypassing the page cache.
+All storage devices were benchmarked using the following script: \url{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}.
 The result of the benchmarks can be seen in Table \ref{table:bandwidths}.
@@
 2 Intel(R) Xeon(R) Platinum 8260M CPU @ 2.40GHz where installed, enabling use of up to 96 threads.
-The server was running Red Hat Enterprise Linux 7.6 release (Maipo).
+The server was running Red Hat Enterprise Linux 7.6 (Maipo) kernel version 3.10.0-957.
 
@@
 
-\subsubsection{Memory Mode}
+\subsubsection{Memory mode}
 
@@
 While enabling access to a larger amount of memory than typically accessible otherwise,
-it is anticipated that Memory mode will be slower than App Direct mode for all memory accessed that
-are not cached in DRAM. We evaluated all available devices (Optane DCPMM through tmpfs, local disk and Isilon)
+it is anticipated that Memory mode will be slower than App Direct mode for all memory accesses that
+are not cached in DRAM, due to reduced storage bandwidth. We evaluated all available devices (Optane DCPMM through tmpfs, local SSD and Isilon)
 in Memory mode.
 
-\subsubsection{App Direct Mode}
+\subsubsection{App Direct mode}
 
@@
 treating them as two distinct memory tiers. Optane DCPMM
-does not use the DRAM as cache in App Direct Mode if configured with DAX.
+does not use the DRAM as cache in App Direct Mode when configured with DAX, as is our case.
 
@@
 
-We characterize the performance of our experiments using the following model:
+We characterize the performance of our data-intensive experiments using the following model:
 
@@
 are certain instances where the makespan may be below the I/O duration. This is expected
-to occur with storage devices that scale well with increased parallelism, as should be the
-case for DRAM, Optane DCPMM and Isilon, which all consist of multiple devices. Should the
+to occur with scalable devices, as should be the
+case for DRAM, Optane DCPMM and Isilon, which can support parallel I/O. Should the
 scalable storage devices predict the makespan accurately, it is believed that other overheads
-would be at play making the device unfavourable for parallel processing.
+would be at play, making the device unfavourable for parallel processing.
 
@@
 enables us to read and write to different storage devices and ensure that written
-data could not have been previously cache
+data could not have been previously cached
 in-memory, by ensuring that read and written data are not the same. This application
-was parallelized in two different ways: using both GNU Parallel and Apache Spark 2.4.3 (PySpark). 
+was parallelized in two different ways: using both GNU Parallel~\cite{gnuparallel} and Apache Spark 2.4.3 (PySpark)~\cite{spark}. 
 All code was implemented in Python 3.6.
@@
 depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
-is achieved using the popular neuroimaging I/O: \href{https://nipy.org/nibabel/}{NiBabel}. When provided with a filename, as is the case
+is achieved using the popular neuroimaging I/O library: \href{https://nipy.org/nibabel/}{NiBabel}. When provided with a filename, as is the case
 for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memory map the data
@@
 
-The Apache Spark implementation differs from that of GNU Parallel in that scheduling decisions
-were entirely differed to Spark. That is, rather than simply parallelizing over the \bigbrain block filenames,
-we opted to read the data using Spark's built-in
+The Apache Spark implementation differs from that of GNU Parallel. Whereas GNU Parallel will
+simply forks processes for each task and defers scheduling to the kernel, Spark is responsible for 
+task scheduling decisions. Furthermore, in our Spark implementation, we opted to read the data using Spark's built-in
 \texttt{BinaryFiles}, loading whole files, in binary format, into Spark partitions. As the
@@
 compliant application. It runs a standard neuroimaging brain extraction application (FSL BET) 
-on all the anatomical (T1W) images of dataset containing numerous subjects. This step is 
+on all the anatomical (T1W) images of datasets containing numerous subjects. This step is 
 referred to as Participant analysis within the application. An optional
@@
 
-For our experiments, we used the entire CoRR dataset, made available on \href{}{DataLad} and
+For our experiments, we used the entire CoRR dataset, made available on \href{https://www.datalad.org/}{DataLad} and
 applied Participant analysis to it. The BIDS App Example was executed using a Singularity container stored on
@@
 
-Time in this application was obtained using Linux's \texttt{time(1)} application.
+Time in this application was obtained using Linux's \texttt{time()} application.
 \texttt{Real + Sys} was used to measure CPU time,
@@
              measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
-             to Eq.~\ref{eq:makespan}. 3 
+             to Eq.~\ref{eq:makespan}. Three 
              repetitions were performed. *Local SSD did not complete the
-    writing of the last few blocks due to storage limitations.}\label{fig:makespan-25cpus}
+         writing of the last few blocks (approximately 5-10) due to storage limitations.}\label{fig:makespan-25cpus}
 \end{figure}
@@
     \captionsetup{belowskip=0pt}
-    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
-             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-25cpus}
+    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. Three repetitions were performed. *Local SSD did not
+    complete the writing of the last few blocks (approximately 5-10) due to storage limitations}\label{fig:stacked-25cpus}
 \end{figure}
@@
     \caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes.
-*Some local SSD writes did not complete due to storage limitations.}\label{fig:gantt25}
+*Some local SSD writes (approximately 5-10 writes) did not complete due to storage limitations.}\label{fig:gantt25}
 \end{figure*}
@@
 parallelism measured for Optane DCPMM was 24. Unlike Optane DCPMM, local SSD spends the vast majority of its time
-reading, while spending a negligible amount of time writing. Like Optane DCPMM, it averages 24 parallel tasks
+reading, while spending a negligible amount of time writing, due to writeback to DRAM. Like Optane DCPMM, it averages 24 parallel tasks
 throughout its execution. Unlike the other three storage, Isilon shows spacing between the loading of the header,
 read and increment and write. Isilon was measured to have an average parallelism of 14. Overall, each storage device
-execution displayed that some tasks took longer than others.
+execution displayed heterogeneous task durations.
 
@@
              measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
-             to Eq.~\ref{eq:makespan}. 3 
+             to Eq.~\ref{eq:makespan}. Three 
              repetitions were performed. *Local SSD did not complete the
-    writing of the last few blocks due to storage limitations.}\label{fig:makespan-96cpus}
+         writing of the last few blocks (approximately 5-10) due to storage limitations.}\label{fig:makespan-96cpus}
 \end{figure}
@@
     \captionsetup{belowskip=-10pt}
-    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
-             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-96cpus}
+    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. Three repetitions were performed. *Local SSD did not
+    complete the writing of the last few blocks (approximately 5-10) due to storage limitations}\label{fig:stacked-96cpus}
 \end{figure}
@@
     \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572306855-tmpfs_1it_96cpus_40bb-2.pdf}
-    \caption{Memory Mode}
+    \caption{Memory mode}
 \end{subfigure}
@@
     \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572379747-optaneAD_1it_96cpus_40bb-1.pdf}
-    \caption{App Direct Mode}
+    \caption{App Direct mode}
 \end{subfigure}
@@
 Despite having increased parallelism by a factor of 3.84, we see no visible reduction
-in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 cpus, DRAM and Optane DCPMM in Memory mode are the only devices that did better than the anticipated I/O, with a makespan duration of
+in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 processes, DRAM and Optane DCPMM in Memory mode are the only devices that did better than the anticipated I/O, with a makespan duration of
 5.7s and 9s, respectively (total read and write estimate for both was measured to be around 39s). While
@@
 A further look at the Gantt charts between Optane DCPMM in Memory mode and App Direct mode (Figure~\ref{fig:gantt96})
-show that both read and write time of tasks are worse in App Direct mode as compare with Memory mode.
+shows that both read and write time of tasks are worse in App Direct mode as compared to Memory mode.
 While some read tasks in App Direct mode appear to be of similar duration to the average read task time
@@
 Unlike the GNU Parallel executions, the Spark executions appear to be in better accordance with
-the I/O estimates (Figure~\ref{fig:20mkspspark25}. Only Optane DCPMM in App Direct mode appears
+the I/O estimates (Figure~\ref{fig:20mkspspark25}). Only Optane DCPMM in App Direct mode appears
 to have significantly exceeded the estimates, by a factor of 2. Total write duration all appear
@@
 \caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
-25 processes. 3 repetitions were performed.}\label{fig:2025}
+25 processes. Three repetitions were performed.}\label{fig:2025}
 \end{figure*}
@@
 \end{subfigure}
-\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 25 processes. 3 repetitions wereperformed}\label{fig:20spark25}
+\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 25 processes. Three repetitions wereperformed}\label{fig:20spark25}
 \end{figure*}
@@
 \caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
-96 processes. 3 repetitions were performed.}\label{fig:2096}
+96 processes. Three repetitions were performed.}\label{fig:2096}
 \end{figure*}
@@
 \end{subfigure}
-\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 96 processes. 3 repetitions were performed}\label{fig:20stackedp96}
+\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 96 processes. Three repetitions were performed}\label{fig:20stackedp96}
 \end{figure*}
@@
 \end{subfigure}
-\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}
+\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices. Three repetitions were performed}
 \end{figure*}
@@
 \end{subfigure}
-\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}\label{fig:bbd}
+\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices. Three repetitions were performed}\label{fig:bbd}
 \end{figure*}
 Unlike the BigBrain incrementation, the makespan of the BIDS App Example
-executing on Optane DCPMM is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
+executing on Optane DCPMM in App Direct mode is longer than local disk in Memory mode (Figure~\ref{fig:bm25}). However, similarly
 to the BigBrain Incrementation, Optane DCPMM in Memory mode performs better than other storage devices
@@
 configured to use DRAM as a writeback cache in App Direct mode, we would have observed a similar
-performance between the two. However, for the 40~$\mu$ \bigbrain executions, read time on Optane DCPMM in
+performance between the two. Furthermore, in the 40~$\mu$ \bigbrain executions, read time on Optane DCPMM in
 App Direct mode appeared to be slower than Memory mode, potentially indicating that the Optane DCPMM was able
@@
 not leverage the writeback cache like the local SSD, is it still found to be superior in performance.
-Therefore, the while the performance of Optane DCPMM may not be enhanced by DRAM, it is still superior to that of
+Therefore, while the performance of Optane DCPMM may not be enhanced by DRAM, it is still superior to that of
 an SSD using the writeback cache and scalable network storage.
@@
 of the longer-than-expected makespan seems to be due to write times. While there may be application overheads,
-they should, in theory, affect Optane DCPMM in App Direct Mode as much DRAM. While it is unknown
+they should, in theory, affect Optane DCPMM in App Direct Mode as much as DRAM. While it is unknown
 why Optane DCPMM writes are performing so poorly, it could be a result of filesystem configuration or 
@@
 
-It is believe Optane DCPMM can be useful for a variety of neuroimaging applications. For
+It is believed that Optane DCPMM can be useful for a variety of neuroimaging applications. For
 instance, typically applications executing in HPC environments can benefit from 

### 2019-12-17 paper.tex
added some additional comments

@@
 
+\subsection{Other comments}
+
+Storage performance when scaling the application from 25 to 96 cores could have been improved using a
+NUMA-aware application, since memory is tied to the CPUs. Should a task requiring memory
+stored on one CPU socket be sent to another CPU, this would incur some overheads, reducing overall application
+and storage performance. Furthermore, Intel optimized Python and NumPy libraries could have been used to improve 
+performance.
+
+While these experiments may not accurately reflect the maximum performance of the storage devices for the applications,
+they are meant to reflect the performance that would be obtained by the average neuroimaging researcher
+executing standardized tools on infrastructure made available to them.
+
 \section{Conclusion}

### 2019-12-17 paper.tex
some fixes

@@
 the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
-1,200 healthy adults, expected to exceed $\SI{2}{\peta\byte}$ in size; 
+1,200 healthy adults, expected to exceed $\SI{1}{\peta\byte}$ in size; 
 and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
 initiative which aggregates MR data from various centres around the world, 
-of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
+of which 32 are currently available and make up about $\SI{937}{\gibi\byte}$ of
 data in total.
@@
 storage technology that resides on the DIMMs to reduce latency to the device.
-The first generation of Intel Optane DCPMM offers capabities of 128~GB,
-256~GB, and 512~GB, enabling it to be more cost effective than DRAM storage.
+The first generation of Intel Optane DCPMM offers capabities of 128~GiB,
+256~GiB, and 512~GiB, enabling it to be more cost effective than DRAM storage.
 There are two configuration modes, Memory mode and App Direct mode,
@@
 The server used consisted of 12X~\SI{64}{\gibi\byte} DRAM devices,
-resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\giga\byte} Optane DCPMM
+resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\gibi\byte} Optane DCPMM
 modules, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC persistent memory.
@@
 App Direct mode were configured as write-through devices, meaning they did not leverage
-DRAM as a cache for writes. For Optane DC persistent memory, this was achieved by configuring the xfs filesystem with
+DRAM as a cache for writes. For Optane DCPMM, this was achieved by configuring the xfs filesystem with
 direct access (DAX).
@@
 
-Memory mode leverages Intel Optane DC persistent memory (DCPMM) to extend the system's available
+Memory mode leverages Optane DCPMM to extend the system's available
 memory. In this mode, Optane DCPMM uses DRAM as a cache and is accessible as
@@
 treating them as two distinct memory tiers. Optane DCPMM
-does not use the DRAM as cache in App Direct Mode when configured with DAX, as is our case.
+does not use the DRAM as cache in App Direct mode when configured with DAX, as is our case.
 
-Similarly to Memory Mode, our experiments evaluated all available filesystems in App Direct
+Similarly to Memory mode, our experiments evaluated all available filesystems in App Direct
 Mode. In other words, we evaluated DRAM (tmpfs mount), Optane DCPMM (\SI{1.5}{\tebi\byte} persistent
@@
 Isilon. As in the BigBrain Incrementation, the experiment was parallelized 
-using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory Mode,
-where Isilon and the local SSD were only evaluated in Memory Mode. Each experiment was repeated 3x.
+using GNU Parallel with 25 and 96 processes. The conditions were executed in both App Direct and Memory mode,
+where Isilon and the local SSD were only evaluated in Memory mode. Each experiment was repeated 3x.
 
@@
     \captionsetup{belowskip=-10pt}
-    \caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes.
+    \caption{Gantt charts for each storage device (App Direct mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes.
 *Some local SSD writes (approximately 5-10 writes) did not complete due to storage limitations.}\label{fig:gantt25}
@@
 Optane DCPMM in this mode is approximately 
-12x slower than Optane in Memory Mode. Isilon in both Memory mode and App Direct 
+12x slower than Optane in Memory mode. Isilon in both Memory mode and App Direct 
 mode is more than 120x slower than Optane DCPMM in Memory mode. Other than for Optane DCPMM, 
@@
 1.4x slower than expected, in both modes. For those where the I/O time was overestimated,
-DRAM was almost 5x faster in reality compared to what was estimated. Optane DCPMM in Memory Mode
+DRAM was almost 5x faster in reality compared to what was estimated. Optane DCPMM in Memory mode
 had a larger difference, with it being 3.8x faster than expected. Isilon in App Direct
@@
 shows that, as expected, I/O times vary greatly between devices, with DRAM having the
-best total read and write speeds. Optane DCPMM in Memory Mode is very close to DRAM in speed,
+best total read and write speeds. Optane DCPMM in Memory mode is very close to DRAM in speed,
 whereas Optane DCPMM in App Direct mode was 7x slower in terms of reading and 23x slower in terms
@@
 The anticipated I/O estimates did not correctly predict the makespan for 25 parallel
-processes (Figure~\ref{fig:20mksp25}). Both Optane DCPMM in Memory Mode and Isilon appear 
-to be under the estimates: Optane DCPMM in Memory Mode is approximately twice as fast as the
+processes (Figure~\ref{fig:20mksp25}). Both Optane DCPMM in Memory mode and Isilon appear 
+to be under the estimates: Optane DCPMM in Memory mode is approximately twice as fast as the
 estimates whereas Isilon is around 1.1x faster. Optane was 1.8x slower than anticipated.
-Once again, Memory and App Direct Mode appear to only affect performance for Optane.
+Once again, Memory and App Direct mode appear to only affect performance for Optane.
 
 Total task duration breakdowns (Figure~\ref{fig:20total25}) show that Optane DCPMM in
-App Direct Mode spends significantly more time writing than its Memory Mode counterpart.
+App Direct mode spends significantly more time writing than its Memory mode counterpart.
 In fact, it spends almost 12x more time writing. Isilon also appears to vary slightly
-between Memory and App Direct Mode, spending around 1.4x more time reading and incrementing
-in App Direct Mode.
+between Memory and App Direct mode, spending around 1.4x more time reading and incrementing
+in App Direct mode.
 
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark-20bb_25cpus.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark_20bb_25cpus.pdf}
     \caption{Total convert/increment/write breadown}\label{fig:20totalspark:25}
@@
 mode makespan (Figure~\ref{fig:20mksp96}). As previously observed with 25 processes,
-the bulk of the processing time is taken up by writing for Optane DCPMM in App Direct Mode and 
+the bulk of the processing time is taken up by writing for Optane DCPMM in App Direct mode and 
 Isilon in both modes (Figure~\ref{fig:20total96}). Reading takes more time in the case of Optane DCPMM 
@@
 A notable difference in the Spark execution is that Optane DCPMM in Memory and App Direct mode
-exhibit the same performance (Figure~\ref{20mkspspark96}). Both of which also performed
+exhibit the same performance (Figure~\ref{fig:20mkspspark96}). Both of which also performed
 significantly worse than the estimated I/O duration. Performance on Isilon, however, did
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-spark-20bb_96cpus.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-spark_20bb_96cpus.pdf}
     \caption{Makespan}\label{fig:20mkspspark96}
@@
         \centering
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark-20bb_96cpus.pdf}
+    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark_20bb_96cpus.pdf}
     \caption{Total convert/increment/write breakdown}\label{fig:20totalspark96}
@@
 \section{Discussion}
-\subsection{Memory vs App Direct Mode}
+\subsection{Memory vs App Direct mode}
 
@@
 that total available DRAM would have been exceeded. However, available DRAM could have been exceeded in
-Memory Mode when using devices that leverage memory as a write cache.
+Memory mode when using devices that leverage memory as a write cache.
 In our case, only the local SSD was set up to do so.
@@
 App Direct mode appeared to be slower than Memory mode, potentially indicating that the Optane DCPMM was able
-to keep the input data cached in DRAM, whereas in App Direct Mode, the data had to be loaded from Optane DCPMM.
-Therefore, it is expected the Optane DCPMM in App Direct mode may always have longer read times than Memory Mode
-if the data is already cached in DRAM in Memory Mode.
+to keep the input data cached in DRAM, whereas in App Direct mode, the data had to be loaded from Optane DCPMM.
+Therefore, it is expected the Optane DCPMM in App Direct mode may always have longer read times than Memory mode
+if the data is already cached in DRAM in Memory mode.
 
 The Apache Spark and GNU Parallel implementations differed vastly in makespan 
-for Optane DCPMM in Memory Mode, particularly at 96 processes. After taking a closer look at
+for Optane DCPMM in Memory mode, particularly at 96 processes. After taking a closer look at
 the captured Spark metrics, what significantly varied between the Spark runs on
@@
 tend to generate temporary data files that are larger than input size, it is likely that maintaining the
-devices as write-through is preferable. Furthermore, despite the fact that Optane DCPMM in App Direct Mode does
+devices as write-through is preferable. Furthermore, despite the fact that Optane DCPMM in App Direct mode does
 not leverage the writeback cache like the local SSD, is it still found to be superior in performance.
@@
 of the longer-than-expected makespan seems to be due to write times. While there may be application overheads,
-they should, in theory, affect Optane DCPMM in App Direct Mode as much as DRAM. While it is unknown
+they should, in theory, affect Optane DCPMM in App Direct mode as much as DRAM. While it is unknown
 why Optane DCPMM writes are performing so poorly, it could be a result of filesystem configuration or 

### 2019-12-17 experiments/bigbrain-incrementation/figures/makespan.py

@@ -53,7 +53,7 @@ def configure_df(df):
 
 def makespan(df, task="task_duration", spark=False):
 
-    labels = ["DRAM", "Optane", "local SSD*", "Isilon"]
+    labels = ["DRAM", "Optane DCPMM", "local SSD*", "Isilon"]
     if not spark:
         dfp = df[df["Task"] == "task_duration"].groupby(["filename"])
     else:
@@ -144,7 +144,7 @@ def makespan(df, task="task_duration", spark=False):
     if bb20:
         ind = np.asarray([0, 1])
         ind_mem = ind
-        labels = ['Optane', 'Isilon']
+        labels = ['Optane DCPMM', 'Isilon']
 
     print(df_memb_mean)
     print(df_memb_std)

### 2019-12-17 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -60,7 +60,7 @@ def stacked_bar(df, spark=False):
 
     global data_size, total_1000
 
-    labels = ["DRAM", "Optane", "local SSD", "Isilon"]
+    labels = ["DRAM", "Optane DCPMM", "local SSD", "Isilon"]
     ind = np.arange(len(labels))
     ind_mem = np.delete(ind, 0)
 
@@ -187,7 +187,7 @@ def stacked_bar(df, spark=False):
     if bb20:
         ind = np.asarray([0, 1])
         ind_mem = ind
-        labels = ['Optane', 'Isilon']
+        labels = ['Optane DCPMM', 'Isilon']
 
     print(data_size)
     print(df_memb_mean)
@@ -316,7 +316,7 @@ def stacked_bar(df, spark=False):
 
     else:
 
-        labels = ["Optane", "Isilon"]
+        labels = ["Optane DCPMM", "Isilon"]
         ind = np.arange(len(labels))
         ind_mem_r = ind
         ind_ad_r = ind
@@ -470,7 +470,7 @@ def stacked_bar(df, spark=False):
     r_inc = mpatch.Patch(facecolor="r", alpha=alpha+0.2, label="read/increment")
 
     if spark:
-        if not bb20:
+        if not bb20 or '25cpu' in sys.argv[1] :
             ax.set_ylim(0, 140000)
         else:
             ax.set_ylim(0, 500000)

### 2019-12-17 gen_figures.sh

@@ -8,10 +8,14 @@ cd $curr_dir/experiments/bigbrain-incrementation/figures
 ./stacked_bar.py '../results/40bb_redo/*25cpus*' 40bb_25cpus
 ./makespan.py '../results/40bb_redo/*96cpus*' 40bb_96cpus
 ./stacked_bar.py '../results/40bb_redo/*96cpus*' 40bb_96cpus
-./makespan.py '../results/20bb_redo/*25cpus*' 20bb_25cpus spark
-./stacked_bar.py '../results/20bb_redo/*25cpus*' 20bb_25cpus spark
-./makespan.py '../results/20bb_redo/*96cpus*' 20bb_96cpus spark
-./stacked_bar.py '../results/20bb_redo/*96cpus*' 20bb_96cpus spark
+./makespan.py '../results/20bb_redo/*25cpus*' 20bb_25cpus
+./makespan.py '../results/20bb_redo/*25cpus*' spark_20bb_25cpus spark
+./stacked_bar.py '../results/20bb_redo/*25cpus*' 20bb_25cpus
+./stacked_bar.py '../results/20bb_redo/*25cpus*' spark_20bb_25cpus spark
+./makespan.py '../results/20bb_redo/*96cpus*' 20bb_96cpus
+./makespan.py '../results/20bb_redo/*96cpus*' spark_20bb_96cpus spark
+./stacked_bar.py '../results/20bb_redo/*96cpus*' 20bb_96cpus
+./stacked_bar.py '../results/20bb_redo/*96cpus*' spark_20bb_96cpus spark
 cd $curr_dir
 echo 'GNU Parallel: 75BB local App Direct 25 cpus - 1572382794-localAD_1it_25cpus_40bb-1.out'
 ./scripts/gantt.py --incrementation_tf ./experiments/bigbrain-incrementation/results/40bb_redo/1572382794-localAD_1it_25cpus_40bb-1.out --cpus 25

### 2019-12-17 paper.tex
Minor fixes, made it fit in 10 pages

@@
 \usepackage{hyperref}
+\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
 \usepackage{xspace}
@@
 analysis. In this paper we evaluate the performance advantage
-brought by Intel\textsuperscript{textregistered}~Optane\texttrademark~DC persistent memory for the
+brought by Intel\textsuperscript{\textregistered}~Optane\texttrademark~DC persistent memory for the
 processing of large neuroimaging datasets using the two available
@@
 processes in both cases. Our results show that the performance of
-the applications leveraging persistent memory is superior to that of other
+applications leveraging persistent memory is superior to that of other
 storage devices, with the exception of DRAM. This is the case
@@
 that enable the storage to either be accessed as an extension of volatile main memory
-or as a non-volatile memory storage device\footnote{https://itpeernetwork.intel.com/intel-optane-dc-persistent-memory-operating-modes/}.
+or as a non-volatile memory storage device~\cite{memory-modes}.
 
@@
 to implement the methods described hereafter are all available at 
-\url{https://github.com/valhayot/paper-memory-storage} for 
+\url{https://github.com/big-data-lab-team/paper-memory-storage} for 
 further inspection and reproducibility.
@@
 
-The server used consisted of 12X~\SI{64}{\gibi\byte} DRAM devices,
-resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12X~\SI{256}{\gibi\byte} Optane DCPMM
+The server used consisted of 12$\times$\SI{64}{\gibi\byte} DRAM devices,
+resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12$\times$\SI{256}{\gibi\byte} Optane DCPMM
 modules, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC persistent memory.
@@
 only \SI{149}{\gibi\byte} where available for the experiments, and a
-\SI{720}{\tebi\byte} shared Dell EMC Isilon network-attached storage platform. Isilon is an nsf4 mounted
-cluster (10Gb network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
+\SI{720}{\tebi\byte} shared Dell EMC Isilon network-attached storage platform. Isilon is an NFS 4 mounted
+cluster (10Gbps network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
 was the only storage device set up as a writeback device. Both Isilon and Optane DCPMM in
 App Direct mode were configured as write-through devices, meaning they did not leverage
-DRAM as a cache for writes. For Optane DCPMM, this was achieved by configuring the xfs filesystem with
+DRAM as a cache for writes. For Optane DCPMM, this was achieved by configuring the XFS filesystem with
 direct access (DAX).
 DAX enables persistent memory to be byte addressable, bypassing the page cache.
-All storage devices were benchmarked using the following script: \url{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}.
+All storage devices were benchmarked using the script available at \url{https://github.com/big-data-lab-team/paper-memory-storage/blob/master/scripts/bench_disks.sh}.
 The result of the benchmarks can be seen in Table \ref{table:bandwidths}.
@@
 For processing,
-2 Intel(R) Xeon(R) Platinum 8260M CPU @ 2.40GHz where installed, enabling use of up to 96 threads.
+2$\times$Intel(R) Xeon(R) Platinum 8260M CPU @ 2.40GHz where installed, enabling use of up to 96 threads.
 The server was running Red Hat Enterprise Linux 7.6 (Maipo) kernel version 3.10.0-957.
@@
  \hline
-\end{tabular}\caption{Measured read and write bandwidth of each storage device}\label{table:bandwidths}
+\end{tabular}\caption{Measured read and write bandwidths}\label{table:bandwidths}
 \end{center}
@@
 have yet to be developed. In order to quantify the effects of storage devices
-on such a dataset, we have implemented a naive synthetic application that takes 
+on such a dataset, we have implemented a basic synthetic application that takes 
 the image, split into blocks, and increments all the voxels within each block, in parallel (Algorithm~\ref{alg:incrementation}).
@@
 depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
-is achieved using the popular neuroimaging I/O library: \href{https://nipy.org/nibabel/}{NiBabel}. When provided with a filename, as is the case
+is achieved using the popular neuroimaging I/O library NiBabel v2.5.0~\cite{nibabel}. When provided with a filename, as is the case
 for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memory map the data
@@
 We have executed this pipeline on both the $\SI{75}{\gibi\byte}$ 40~$\mu$m 
-\bigbrain split into 125 $\SI{614}{\mebi\byte}$ blocks and the $\SI{603}{\gibi\byte}$
-\bigbrain at 20~$\mu$m split into 1000 $\SI{617}{\mebi\byte}$ blocks.
+\bigbrain split into 125$\times$$\SI{614}{\mebi\byte}$ blocks and the $\SI{603}{\gibi\byte}$
+\bigbrain at 20~$\mu$m split into 1000$\times$$\SI{617}{\mebi\byte}$ blocks.
 
@@
 using both Apache Spark and GNU Parallel. The configuration was also generally the same, having 
-experiments using both 25 and 96 cpus. Repeating the experiments using a larger dataset would enable
+experiments using both 25 and 96 processes. Repeating the experiments using a larger dataset would enable
 us to determine the effects when Optane DCPMM would have to be partially relied upon due to insufficient DRAM
@@
 The BIDS App Example is a template example for creating a Brain Imaging Data Structure (BIDS)
-compliant application. It runs a standard neuroimaging brain extraction application (FSL BET) 
-on all the anatomical (T1W) images of datasets containing numerous subjects. This step is 
+compliant application. It runs a standard neuroimaging brain extraction application 
+on all the anatomical images of datasets containing numerous subjects. This step is 
 referred to as Participant analysis within the application. An optional
@@
 
-For our experiments, we used the entire CoRR dataset, made available on \href{https://www.datalad.org/}{DataLad} and
+We used the entire CoRR dataset, available on \href{https://www.datalad.org/}{DataLad} and
 applied Participant analysis to it. The BIDS App Example was executed using a Singularity container stored on
@@
 \subsubsection{25 Processes}
+
 \begin{figure}
+    \begin{subfigure}{\columnwidth}
+        \centering
         \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpus.pdf}
+        \caption{Makespan}\label{fig:makespan-25cpus}
+    \end{subfigure}
+    \begin{subfigure}{\columnwidth}
+        \centering
+        \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_25cpus.pdf}
+        \caption{Total read/increment/write breakdown}\label{fig:stacked-25cpus}
+    \end{subfigure}
     \captionsetup{belowskip=-10pt}
-    \caption{Makespan of the \bigbrain Incrementation application using 25 
-             processes on the 125 blocks on the 40~$\mu$m image given different
-             storage configurations. Anticipated read and write duration was 
+    \caption{GNU Parallel incrementation application processing the 40 $\mu$m BigBrain using 25 processes (three repetitions). Anticipated read and write duration was 
     measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
-             to Eq.~\ref{eq:makespan}. Three 
-             repetitions were performed. *Local SSD did not complete the
-         writing of the last few blocks (approximately 5-10) due to storage limitations.}\label{fig:makespan-25cpus}
+    to Eq.~\ref{eq:makespan}. *Local SSD did not complete the
+    writing of the last few blocks (approximately 5-10) due to storage limitations.}
 \end{figure}
-\begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_25cpus.pdf}
-    \captionsetup{belowskip=0pt}
-    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. Three repetitions were performed. *Local SSD did not
-    complete the writing of the last few blocks (approximately 5-10) due to storage limitations}\label{fig:stacked-25cpus}
-\end{figure}
+
 \begin{figure*}
@@
 Optane DCPMM was approximately 2.2x slower than expected, whereas local SSD was approximately
-1.4x slower than expected, in both modes. For those where the I/O time was overestimated,
+1.4x slower than expected, in both modes. Conversely,
 DRAM was almost 5x faster in reality compared to what was estimated. Optane DCPMM in Memory mode
@@
 \subsubsection{96 Processes}
+
 \begin{figure}
+    \begin{subfigure}{\columnwidth}
+        \centering
         \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_96cpus.pdf}
+        \caption{Makespan}\label{fig:makespan-96cpus}
+    \end{subfigure}
+    \begin{subfigure}{\columnwidth}
+        \centering
+        \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_96cpus.pdf}
+        \caption{Total read/increment/write breakdown}\label{fig:stacked-96cpus}
+    \end{subfigure}
     \captionsetup{belowskip=-10pt}
-    \caption{Makespan of the \bigbrain Incrementation application using 96 
-             processes on the 125 blocks on the 40~$\mu$m image given different
-             storage configurations. Anticipated read and write duration was 
-             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
-             to Eq.~\ref{eq:makespan}. Three 
-             repetitions were performed. *Local SSD did not complete the
-         writing of the last few blocks (approximately 5-10) due to storage limitations.}\label{fig:makespan-96cpus}
-\end{figure}
-\begin{figure}
-    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_96cpus.pdf}
-    \captionsetup{belowskip=-10pt}
-    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. Three repetitions were performed. *Local SSD did not
-    complete the writing of the last few blocks (approximately 5-10) due to storage limitations}\label{fig:stacked-96cpus}
+    \caption{GNU Parallel incrementation application processing the 40 $\mu$m BigBrain using 96 processes (three repetitions). *Local SSD did not complete the
+    writing of the last few blocks (approximately 5-10) due to storage limitations.}
 \end{figure}
@@
 
-When analyzing the average parallelism during the workload, only Optane DCPMM in App Direct mode
-came close to 96 parallel tasks, having an average parallelism of 88 tasks. Optane DCPMM in Memory mode,
+When analyzing the average parallelism, only Optane DCPMM in App Direct mode
+came close to 96 parallel tasks, with an average parallelism of 88 tasks. Optane DCPMM in Memory mode,
 DRAM and local SSD all averaged between 61-66 tasks. Isilon performed the worst with 38 tasks in Memory 
@@
 
-\subsection{20~$\mu$ \bigbrain incrementation}
+\subsection{20~$\mu$m \bigbrain incrementation}
 \subsubsection{25 Processes}
@@
 Total task duration breakdowns (Figure~\ref{fig:20total25}) show that Optane DCPMM in
-App Direct mode spends significantly more time writing than its Memory mode counterpart.
+App Direct mode spends significantly more time writing than in Memory mode.
 In fact, it spends almost 12x more time writing. Isilon also appears to vary slightly
@@
 \caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
-25 processes. Three repetitions were performed.}\label{fig:2025}
+25 processes (three repetitions).}\label{fig:2025}
 \end{figure*}
@@
 \end{subfigure}
-\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 25 processes. Three repetitions wereperformed}\label{fig:20spark25}
+\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 25 processes (three repetitions).}\label{fig:20spark25}
+\captionsetup{belowskip=-10pt}
 \end{figure*}
@@
 
-A very slight performance improvement (1-1.3x faster) can be witnessed in Optane DCPMM. In 
+A very slight performance improvement (1-1.3x faster) can be witnessed in Optane DCPMM (Figure~\ref{fig:20mksp96}). In 
 contrast, Isilon 
@@
 \caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
-96 processes. Three repetitions were performed.}\label{fig:2096}
+96 processes (three repetitions).}\label{fig:2096}
+\captionsetup{belowskip=-10pt}
 \end{figure*}
@@
 \end{subfigure}
-\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 96 processes. Three repetitions were performed}\label{fig:20stackedp96}
+\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 96 processes (three repetitions).}\label{fig:20stackedp96}
+\captionsetup{belowskip=-10pt}
 \end{figure*}
@@
 \end{subfigure}
-\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices. Three repetitions were performed}
+\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices (three repetitions).}
+\captionsetup{belowskip=-10pt}
 \end{figure*}
@@
 \end{subfigure}
-\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices. Three repetitions were performed}\label{fig:bbd}
+\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices (three repetitions).}\label{fig:bbd}
+\captionsetup{belowskip=-10pt}
 \end{figure*}
@@
 the executions of 25 and 96 processes, the performance improved by a similar factor on all
-storage devices.
+devices.
 
 The CPU and I/O breakdowns (Figure~\ref{fig:bbd}) show that there is a significant increase
-in the time spent on I/O at 96 processes when compared to 25 processes. Furthermore, there appears
+in the I/O time at 96 processes when compared to 25 processes. Furthermore, there appears
 to also be an increase in the amount of time spent on CPU processing. In both cases,
-tmpfs spends the least amount of time on I/O and CPU, followed by local disk, Optane, and finally,
-Isilon.
+tmpfs spends the least amount of time on I/O and CPU, followed by local disk, Optane, and Isilon.
 \section{Discussion}
@@
 all our dataset sizes. While total memory used by the application may have exceeded DRAM, as is the
-case with the 20~$\mu$ \bigbrain, the amount of memory required by the application at any 
+case with the 20~$\mu$m \bigbrain, the amount of memory required by the application at any 
 given moment did not exceed the amount of available DRAM. Although GNU Parallel would have
@@
 Optane DCPMM was the only device affected by the choice of storage mode. In all cases, with the 
-exception of the Spark incrementation of the 20~$\mu$ \bigbrain, Memory mode was superior to 
+exception of the Spark incrementation of the 20~$\mu$m \bigbrain, Memory mode was superior to 
 App Direct mode. This is a result of Optane DCPMM using DRAM as cache in Memory mode, whereas Optane
@@
 configured to use DRAM as a writeback cache in App Direct mode, we would have observed a similar
-performance between the two. Furthermore, in the 40~$\mu$ \bigbrain executions, read time on Optane DCPMM in
+performance between the two. Furthermore, in the 40~$\mu$m \bigbrain executions, read time on Optane DCPMM in
 App Direct mode appeared to be slower than Memory mode, potentially indicating that the Optane DCPMM was able
@@
 The use of Optane DCPMM in Memory mode would enable more data to be written to available memory
-rather than directly to slower storage devices. In our 40~$\mu$ \bigbrain experiments, it was found
+rather than directly to slower storage devices. In our 40~$\mu$m \bigbrain experiments, it was found
 that read times were less important on Isilon than on the local SSD (Figures~\ref{fig:stacked-25cpus}
-and~\ref{fig:stacked-96cpus}). Should Isilon have been given the opportunity to benefit from a 
+and~\ref{fig:stacked-96cpus}). Should Isilon have benefitted from a 
 writeback cache like the SSD, it is believed that performance on Isilon could have been superior 
@@
 For instance, taking into consideration the total number of DRAM devices, the expected makespan of the
-40~$\mu$ \bigbrain is estimated to take around 3s, whereas sequentially it was measured to take
+40~$\mu$m \bigbrain is estimated to take around 3s, whereas sequentially it was measured to take
 around 37.5s. At 25 processes, we found the DRAM makespan to be around 7s, which is much closer to the 3s
-estimate than the 37.5s. Reasons for why it took longer than estimated could be due to the fact that
+estimate than the 37.5s. Reasons for why it took longer than estimated could be that
 our applications were not optimized for non-uniform memory accesses (NUMA-aware), or even that some DRAM devices were occupied, and therefore, the maximum amount of parallelism 
@@
 
-Another interesting aspect of Isilon is the spacing that was observed between the read and writes
+Another interesting aspect of Isilon is the spacing observed between the read and writes
 (Figure~\ref{fig:gantt25isilon}). What occurred between the read/increment and
 write operations was simply writing the previous task's benchmarks to a unique
-file. This did not appear to affect any other storage device. This appears to imply
+file. This did not appear to affect any other storage device. This implies
 that the latency alone of writing to Isilon was quite significant.
@@
 can bring the performance close to DRAM speeds. Extending available memory using Optane DCPMM is also 
-expected to help reduce I/O times of write-back devices by potentially extending available cache space.
-It is unclear, however, if traditional SSDs can perform just as well as Optane DCPMM with additional storage devices
+expected to reduce I/O times of write-back devices by extending available cache space.
+It is unclear, however, if SSDs can perform just as well as Optane DCPMM with additional storage devices
 attached.
@@
 \section{Acknowledgement}
-We are thankful to Intel and Dell
-for the 90-day trial of the Intel Optane DC Server and high-quality technical support.
+We are very grateful to the Intel and Dell-EMC teams for giving us the
+opportunity to experiment with the Intel Optane DC Server, for providing
+high-quality technical support, and for their thorough feedback on the
+manuscript. 
 \bibliographystyle{IEEEtran} 

### 2019-12-17 experiments/bigbrain-incrementation/figures/stacked_bar.py

@@ -60,7 +60,7 @@ def stacked_bar(df, spark=False):
 
     global data_size, total_1000
 
-    labels = ["DRAM", "Optane DCPMM", "local SSD", "Isilon"]
+    labels = ["DRAM", "Optane DCPMM", "local SSD*", "Isilon"]
     ind = np.arange(len(labels))
     ind_mem = np.delete(ind, 0)
 

### 2019-12-17 scripts/gantt.py

@@ -74,7 +74,7 @@ def gantt_increment(df, data_file, ncpus, spark=False):
     #ax.set_yticks(range(len(labels)))
     #ax.set_yticklabels([])
     ax.set_yticks([], [])
-    ax.set_ylabel("thread")
+    ax.set_ylabel("process")
 
     if not spark:
         if '25cpu' in data_file:

### 2019-12-17 paper.tex
Final fixes

@@
 In-line Memory Module (DIMM), thereby reducing the latency of accessing data on 
-flash storage devices. While the latency of these devices is improved, the bandwidth
+storage devices. While the latency of these devices is improved, the bandwidth
 remains the same. However, as noted by \cite{nvdimms}, a severe performance degradation
@@
 using both Apache Spark and GNU Parallel. The configuration was also generally the same, having 
-experiments using both 25 and 96 processes. Repeating the experiments using a larger dataset would enable
+experiments using both 25 and 96 processes. Repeating the experiments with the 20~$\mu$m BigBrain would enable
 us to determine the effects when Optane DCPMM would have to be partially relied upon due to insufficient DRAM
@@
 As can be seen in Figure~\ref{fig:makespan-25cpus}, DRAM (App Direct mode) is the 
-most efficient, with a makespan of 7.6s. Optane in Memory mode is a close second
-with a makespan of approximately 10s. As caching was disabled in App Direct mode, 
+most efficient, with a makespan of 7.6s (all makespan values are available \href{https://github.com/big-data-lab-team/paper-memory-storage/blob/master/experiments/bigbrain-incrementation/results/makespan.csv}{here}).
+Optane DCPMM in Memory mode is a close second
+with a makespan of 10.2s. As caching was disabled in App Direct mode, 
 Optane DCPMM in this mode is approximately 
-12x slower than Optane in Memory mode. Isilon in both Memory mode and App Direct 
+12x slower than Optane DCPMM in Memory mode. Isilon in both Memory mode and App Direct 
 mode is more than 120x slower than Optane DCPMM in Memory mode. Other than for Optane DCPMM, 
@@
 become throttled, this may slow down the write task duration. In such cases,
-it may be preferable to set them up as write-through devices. As Big Data neuroimaging applications
-tend to generate temporary data files that are larger than input size, it is likely that maintaining the
-devices as write-through is preferable. Furthermore, despite the fact that Optane DCPMM in App Direct mode does
+it may be preferable to set them up as write-through devices. In any case,
+ despite the fact that Optane DCPMM in App Direct mode does
 not leverage the writeback cache like the local SSD, is it still found to be superior in performance.
@@
 
-Storage performance when scaling the application from 25 to 96 cores could have been improved using a
+Storage performance when scaling the application from 25 to 96 CPU threads could have been improved using a
 NUMA-aware application, since memory is tied to the CPUs. Should a task requiring memory

### 2019-12-17 scripts/gantt.py

@@ -74,7 +74,7 @@ def gantt_increment(df, data_file, ncpus, spark=False):
     #ax.set_yticks(range(len(labels)))
     #ax.set_yticklabels([])
     ax.set_yticks([], [])
-    ax.set_ylabel("process")
+    ax.set_ylabel("CPU threads")
 
     if not spark:
         if '25cpu' in data_file:

### 2019-12-17 experiments/bigbrain-incrementation/figures/makespan.py
added makespan csv

@@ -30,7 +30,7 @@ def configure_df(df):
     print(df_data.head(5))
     df_data["makespan"] = df_data.End - df_data.Start
     df_data["disk"] = df_data.index.map(
-        lambda x: x.split("_")[0].split("-")[-1] + ("-em" if "_em" in x else "-real")
+        lambda x: x.split("_")[0].split("-")[-1]
     )
     s_mean = (
         df_data.rename(columns={"makespan": "mean"})
@@ -66,20 +66,20 @@ def makespan(df, task="task_duration", spark=False):
 
     if not spark:
         df_pres_mem = df_pres[
-            df_pres.index.map(lambda x: x in ["tmpfs-real", "local-real", "isilon-real"])
+            df_pres.index.map(lambda x: x in ["tmpfs", "local", "isilon"])
         ]
         df_pres_mem = df_pres_mem.sort_index(ascending=False)
         df_pres_ad = df_pres[df_pres.index.map(lambda x: "AD" in x)]
         df_pres_ad = df_pres_ad.sort_index(ascending=False)
-        df_pres_ad.rename(
-            index={
-                "tmpfsAD-real": "tmpfs",
-                "optaneAD-real": "optane",
-                "localAD-real": "local",
-                "isilonAD-real": "isilon",
-            },
-            inplace=True,
-        )
+        #df_pres_ad.rename(
+        #    index={
+        #        "tmpfsAD-real": "tmpfs",
+        #        "optaneAD-real": "optane",
+        #        "localAD-real": "local",
+        #        "isilonAD-real": "isilon",
+        #    },
+        #    inplace=True,
+        #)
         print(df_pres)
     else:
         df_sres = configure_df(dfs)
@@ -150,7 +150,7 @@ def makespan(df, task="task_duration", spark=False):
     print(df_memb_std)
     print(df_adb_mean)
     print(df_adb_std)
-
+    
     width = 0.15
 
     fig, ax = plt.subplots()
@@ -160,6 +160,26 @@ def makespan(df, task="task_duration", spark=False):
         df_pres_mem = df_sres_mem
         df_pres_ad = df_sres_ad
 
+
+    df_pres_mem['Mode'] = 'Memory mode'
+    df_pres_ad['Mode'] = 'App Direct mode'
+
[…]

### 2019-12-17 gen_figures.sh

@@ -1,5 +1,6 @@
 #!/bin/bash
 
+rm ./experiments/bigbrain-incrementation/makespan.csv
 curr_dir=`pwd`
 cd ./experiments/bids-app-example
 ./plot.py

### 2020-03-21 paper.tex
some text fixes

@@
       #1 \color{black}}
+\captionsetup{belowskip=-1pt}
 
@@
 \begin{abstract}
-    Open-access neuroimaging datasets have reached petabyte scale, and continue to grow.
+    Open-access neuroimaging datasets have reached petabyte scale and continue to grow.
 The ability to leverage the entirety of these datasets is limited to a restricted
-number of labs with both the capacity and infrastructure to
+number of research laboratories with both the capacity and infrastructure to
 process the data. Whereas Big Data engines have significantly
@@
 storage devices, with the exception of DRAM. This is the case
-in both Memory and App Direct mode and irrespective of the
+in both Memory and App Direct mode, irrespective of the
 amount of data and parallelism. Furthermore, persistent memory in App
@@
 
-Furthermore, intermediary data is often required for post-processing analysis. This limits
+Intermediary data is often required for post-processing analysis. This limits
 any performance benefits that can arise from volatile in-memory computing as intermediary
@@
 is configurable, but nevertheless limited. When writes approach the cache's capacity,
-processes performing writes start to be throttled. Once the cache capacity is reached,
+processes performing writes begin to be throttled. Once the cache capacity is reached,
 processes can no longer use the cache until all cached written data is flushed
@@
 
-Whereas Operating Systems and popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
+Whereas operating systems (OS) and popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
 incorporated software solutions to limit data transfers (e.g. writeback cache, in-memory computing,
@@
 have yet to be developed. In order to quantify the effects of storage devices
-on such a dataset, we have implemented a basic synthetic application that takes 
-the image, split into blocks, and increments all the voxels within each block, in parallel (Algorithm~\ref{alg:incrementation}).
+on such a dataset, we have implemented a basic synthetic application that splits an image into blocks and increments all the voxels within each block, in parallel (Algorithm~\ref{alg:incrementation}).
 This application
@@
 are done within a single task. This task is applied to all the \bigbrain blocks using GNU Parallel,
-processing a subset of them at a time,
+which then processes a subset of them at a time
 depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
@@
 opportunity to experiment with the Intel Optane DC Server, for providing
-high-quality technical support, and for their thorough feedback on the
-manuscript. 
+high-quality technical support, and, in  particular, Steve Scargall, for
+providing thorough technical support on the manuscript.
 \bibliographystyle{IEEEtran} 
